### 2022-06-06 src/tex/ms.tex
Added smooth color distribution plot.

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}

\newcommand{\comment}[1]{}

% note command for inline notes
\newcommand{\note}[1]{\textsf{\textcolor{red}{#1}}}
% command to note a missing citation
\newcommand{\needscite}{\note{[needs citation]}}

% some commonly used math expressions
\newcommand{\px}{p^{}_{X}}
\newcommand{\pu}{p^{}_{U}}
\newcommand{\R}{\mathbb{R}}

\shorttitle{Modeling Galaxy Catalogues with Normalizing Flows}
\shortauthors{Crenshaw et al.}

% Begin!
\begin{document}

% Title
\title{Statistical Modeling of Galaxy Catalogues with Normalizing Flows}

% Author list
\correspondingauthor{John Franklin Crenshaw}
\email{jfc20@uw.edu}

\author[0000-0002-2495-3514]{John Franklin Crenshaw}
\affiliation{DIRAC Institute and Department of Physics, University of Washington, Seattle, WA 98195, USA}

\author{J. Bryce Kalmbach}

\author{Alexander Gagliano}

\author{Alex Malz}

\author{Sam Schmidt}

\author{Francois Lanusse}

\collaboration{99}{(The LSST Dark Energy Science Collaboration)}

% Abstract
\begin{abstract}
    Normalizing flows are useful for modeling high-dimensional probability distributions.
    We give some background and introduce PZFlow, a Python package for statistical modeling of tabular data.
    We demonstrate via modeling photometry and redshifts for a galaxy catalog, and estimate photo-z posteriors.
    \end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Astronomical data represent realizations of complex probability distributions.
Need to write more here.

While direct access to the probability distributions that characterize our data is usually impossible, machine learning provides powerful tools to learn these distributions from samples thereof.
Generative Adversarial Networks (GANs; \needscite) and Variational Autoencoders (VAEs; \needscite) are two popular methods that provide generative models trained on samples.
However, neither of these methods allow exact evaluation of probability densities.

Normalizing flows are a deep learning tool for density estimation and generative modeling that allow efficient sampling and exact likelihood calculation.
The ability to generate samples from a probability distribution with respect to which you can calculate exact likelihoods is very powerful for generative modeling, as you can provide a notion of ``true'' posteriors over sample properties when simulating new data.
For example, when simulating galaxy catalogs for photometric redshift (photo-z) estimation, you can calculate the ``true'' redshift posterior for each galaxy.
This is useful for ...

To facilitate the statistical modeling of galaxy catalogs and other astronomical data sets, we have developed PZFlow, a normalizing flow package for python.
With relatively little tuning required by the user, PZFlow can provide a generative model for any tabular data, including continuous and discrete variables, and variables with Euclidean or periodic topology.
In addition to generative modeling, PZFlow can calculate posteriors over any columns in your data set, and can convolve errors and marginalize over missing columns while training the model or calculating posteriors for samples.

\note{[need to work in citations to Francois' paper on Deep Generative Models for Galaxy Image Sims, and any other examples of normalizing flows in the astronomy literature.]}

Describe outline of the paper\dots

Mention PZFlow and note that all of the NFs in this paper were created using it.

Maybe mention ShowYourWork and how the scripts and data are all available at the links in the paper.

\section{Normalizing Flows}
\label{sec:nf}

\begin{figure*}[t!]
    \script{plot_two_moons.py}
    \begin{centering}
        \includegraphics{figures/twomoons_example.pdf}
        \caption{
            A normalizing flow demonstrated on the two moons data set from \texttt{scikit-learn}.
            The two moons data on the left is mapped onto a two dimensional uniform distribution by the bijection $f$.
            The data are colored by quadrant to visualize their image in the latent space.
            You can sample the data distribution by sampling from the uniform distribution, and using $f^{-1}$ to map the back to the data space.
        }
        \label{fig:two-moons}
    \end{centering}
\end{figure*}

Normalizing flows model complex, high-dimensional probability distributions by learning a mapping from the data distribution to a tractable latent distribution\footnote{Some of the machine learning literature defines the mapping in the opposite direction.}.
Often the latent distribution is a standard Normal distribution, and so the mapping ``normalizes'' the data, hence the name ``noramlizing flow''.
This mapping allows us to sample and evaluate densities using the latent distribution, rather than the unknown data distribution.

Assume we have a differentiable function $f$ that maps samples $x$ from the data distribution $\px$ onto samples $u$ from the latent distribution $\pu$.
Using the change of variables formula, we can evaluate the probability density of the data:
\begin{align}
    \px(x) = \pu(u=f(x)) \, |\det \nabla f(x)|,
    \label{eq:px}
\end{align}
where $\nabla f(x)$ is the Jacobian of $f$ evaluated at $x$.
In words, computing the density $\px(x)$ is accomplished by mapping $x$ to the latent distribution, calculating its density there, and multiplying by the associated Jacobian determinant, which accounts for how the function $f$ distorts volume elements of the space.

If we further assume that is invertible, we can sample from the data distribution by appling $f^{-1}$ to samples from the latent distribution:
\begin{align}
    x = f^{-1}(u) \quad \text{where} \quad u \sim \pu.
\end{align}

Figure \ref{fig:two-moons} shows an example of a normalizing flow that transforms the \texttt{scikit-learn} two moons distribution into a uniform distribution.
The data points are colored by quadrant to visualize their image under the bijection $f$.

The following sections discuss how to build a normalizing flow to model data with various features.
Section \ref{sec:bijections} discusses the bijection $f$ and introduces the building blocks from which our bijections will be built;
Section \ref{sec:latent} discusses how to choose an appropriate latent distribution for your data;
Section \ref{sec:conditional} describe how to condition a normalizing flow on some conditional variables;
Section \ref{sec:periodic} explains how to model data with periodic topology;
finally Section \ref{sec:discrete} explains how to model data with discrete variables.

\subsection{Designing a bijection}
\label{sec:bijections}

The bijection of a normalizing flow must be powerful enough to model complex relationships in data, while simultaneously possessing an efficiently computable Jacobian determinant.
This latter constraint is the primary difficulty in designing a normalizing flow.
The most popular strategy for achieving these requirements is to exploit the fact that a composition of bijections is also bijective.
By chaining together multiple less-expressive bijections whose Jacobians are efficiently computable, a composite bijections can be constructed that meets our requirements:
\begin{align}
    f &= \dots \circ f_3 \circ f_2 \circ f_1.
\end{align}
The overall Jacobian determinant can be efficiently calculated using the chain rule.

There is an extensive literature on constructing these sub-bijections (see \citealt{kobyzev2020} for a review).
Some bijections are specialized to be particularly efficient at either density estimation or sampling, but for many science cases, we wish to do both.
For this reason, we will focus on Rational-Quadratic Rolling Spline Couplings (RQ-RSCs), which achieve state-of-the-art performance, while being efficient with both tasks \citep{durkan2019}.

\subsubsection{Rational-Quadratic Rolling Spline Couplings}
\label{sec:rq-rsc}

\begin{figure}
    \centering
    \begin{tikzpicture}[thick, outer sep=0]
        \node [draw, rectangle, minimum height=2cm, minimum width=1.2cm, fill=blue!20] (1) {$x^{}_{1:d}$};
        \node [draw, rectangle, minimum height=2cm, minimum width=1.2cm, fill=red!20, below = 0 of 1] (2) {$x^{}_{d+1:D}$};
        \node [draw, circle, minimum size=8mm, fill=gray!20, right = 2cm of 1] (3) {$=$};
        \node [draw, circle, minimum size=8mm, fill=gray!20, right = 2cm of 2] (4) {$g$};
        \node [draw, rectangle, minimum height=2cm, minimum width=1.2cm, fill=blue!20, right = 1.5cm of 3] (5) {$y^{}_{1:d}$};
        \node [draw, rectangle, minimum height=2cm, minimum width=1.2cm, fill=red!20, below = 0 of 5] (6) {$y^{}_{d+1:D}$};
        \draw[->] (1) -- (3);
        \draw[->] (2) -- (4);
        \draw[->] (3) -- (5);
        \draw[->] (4) -- (6);
        \draw[->] (1) -- (4) node [midway, draw, circle, fill=white, minimum size=8mm, fill=gray!20] {$m$};
    \end{tikzpicture}
    \caption{
    Diagram of a coupling layer.
    The first partition, $x^{}_{1:d}$, is passed through the layer unchanged.
    The second partition, $x^{}_{d+1:D}$, is transformed by the coupling law $g$, which is parameterized by the coupling function $m$ applied to the first partition.
    }
    \label{fig:coupling}
\end{figure}

RQ-RSCs are based on coupling layers \citep{dinh2015, dinh2017}.
A coupling layer partitions the data, $x \in \R^D$, into two sets, $x_{1:d}$ and $x_{d+1:D}$.
The first set is then used to transform the second set:
\begin{align}
    \begin{split}
    y^{}_{1:d} &= x^{}_{1:d} \\
    y^{}_{d+1:D} &= g(x^{}_{d+1:D}; m(x^{}_{1:d})),
    \end{split}
\end{align}
where $g : \R^{D-d} \times \R^d \to \R^{D-d}$ is an invertible \emph{coupling law}, and $m$ is a \emph{coupling function} defined on $\R^d$.
This is illustrated in Figure \ref{fig:coupling}.
The advantage of this structure is that the Jacobian is triangular,
\begin{align}
    \frac{\partial y}{\partial x} =
    \begin{pmatrix}
         I_d & 0 \\
         \frac{\partial y^{}_{d+1:D}}{\partial x^{}_{1:d}}
         & \frac{\partial y^{}_{d+1:D}}{\partial x^{}_{d+1:D}}
    \end{pmatrix},
\end{align}
and in particular, the Jacobian determinant is
\begin{align}
    \det \frac{\partial y}{\partial x} = \det \frac{\partial y^{}_{d+1:D}}{\partial x^{}_{d+1:D}}.
\end{align}
Furthermore, the inverse can be calculated as
\begin{align}
    \begin{split}
    x^{}_{1:d} &= y^{}_{1:d} \\
    x^{}_{d+1:D} &= g^{-1}(y^{}_{d+1:D}; m(x^{}_{1:d})),
    \end{split}
\end{align}
Notice that neither inverting a coupling layer nor calculating the Jacobian determinant requires inverting or taking derivatives of the coupling function $m$, which can thus be arbitrarily complex.

The obvious limitation of a coupling later is that only a subset of the data variables are transformed.
This is overcome by stacking multiple coupling layers in succession, and switching which variables belong to which partition.
In practice, this is achieved by interspersing coupling layers with bijections that shuffle the dimensions of $x$.
These shuffling bijections are trivially inverted and have a Jacobian determinant of one.

There are a variety of different coupling laws one can use.
One particularly expressive choice is a Rational-Quadratic Neural Spline Coupling (RQ-NSC; \citealt{durkan2019}).
As the name suggests, the \emph{coupling law} for RQ-NSCs is a set of rational-quadratic splines and the \emph{coupling function} is a neural network.
RQ-NSCs define a spline $g_i: [-B, B] \subset \R \to [-B, B]$ for each dimension $i$ of $x_{d+1:D}$, where $g_i$ is a piecewise combination of $K$ segments, and each segment is a rational-quadratic function.
The positions and derivatives of the knots that parameterize the splines are calculated from $x_{1:d}$ by a neural network.

RQ-NSCs are state-of-the-art for efficient sampling and density estimation \citep{kobyzev2020}, and are flexible enough to model complex distributions with multiple discontinuities and hundreds of modes.
In addition, they are easily adaptable for flows with periodic topology (Section \ref{sec:periodic}).
For more details, see \citet{durkan2019}.

In this work, we introduce a simple extension of RQ-NSCs, which we name Rational-Quadratic Rolling Spline Couplings (RQ-RSCs).
An RQ-RSC models $D$ dimensional data by stacking $D$ RQ-NSC layers (with $d=D-1$), each followed by a Rolling layer.
A Rolling layer shifts the dimensions of $x$ by one place:
\begin{align}
    \mathrm{Roll}: [x_1, \dots , x_{D-1}, x_D] \to [x_D, x_1, \dots , x_{D-1}].
\end{align}
In other words, RQ-RSCs individually transform each of the $D$ dimensions of $x$ as a function of the other $D-1$ dimensions.
In the limit of high spline resolution (i.e. $K \to \infty$), RQ-RSCs can model any differentiable, monotonic function on $[-B, B]^D$ and can thus model arbitrarily complex distributions in this region.
In practice, we find very good performance for diverse data sets with much smaller $K$.

Note that you can also specify a different value of $K$ for each of the $D$ spline layers in order to individually control the resolution of each dimension.

\subsubsection{Data processing bijections}
\label{sec:data-processing}

While RQ-RSCs perform the heavy lifting of mapping the data distribution $\px$ onto the latent distribution $\pu$, it is also convenient to define other bijections that perform useful operations such as pre- and post-processing.
We name these \emph{data processing bijections}.

For example, RQ-RSCs (and the RQ-NSCs on which they are based) are defined on the domain [-B, B], and thus will not transform samples outside this range.
It is therefore useful to define a \emph{Shift Bounds} bijection, which shifts the original range of each dimension to match the domain of the splines.
Note that this shift must be set at training time, with the assumption that future test data will lie within the same bounds\footnote{This bijection may be poorly suited to data with large outliers. In that case, more care should be taken in designing a bijection that maps the data onto the range of the splines.}.
You can choose a range wider than that covered by the training set if you wish to allow the flow to sample outside the range of the training set.

\note{Discuss the Inverse Softplus Bijection here! If I use it below. Not sure if I will...}

For an example of building an application-specific data processing bijection, see the \emph{Color Transform} bijection defined in Section \ref{sec:fwd-model}, which maps galaxy magnitudes to galaxy colors.
See section \ref{sec:discrete} for data processing bijections, which enable the flow to model discrete data.

Instead of using these data processing bijections, you can of course manually pre-process the data before evaluating densities and post-process samples drawn from the normalizing flow.
However, by building pre- and post-processing directly into the bijection, you remove these extra steps from the workflow.
This reduces the complexity of working with the normalizing flow and ensures that the flow always ``remembers'' how to correctly perform these pre- and post-processing steps.

\subsection{Choosing a latent distribution}

\note{
Discuss latent distributions here.
Note that most applications use a Normal Distribution, but since the splines have a compact domain, we need to choose a latent distribution with compact support, or else we might generate strange outliers when sampling from the flow.
Some text for this might already be in the creation module draft.
}

\subsection{Conditional flows}
\label{sec:conditional}

The bijectors and latent distributions discussed above can be easily adapted to directly learn conditional probability distributions:
you only need to make the replacement $f(x) \to f(x;y)$, where $y$ is a vector of conditions \citep{winkler2019}.
In practice, as the bijection $f$ is parameterized by a neural network, this can be simply achieved by appending the vector $y$ to the neural network's input.

While $p(x|y)$ is technically encoded within $p(x,y)$, which can be learned with a regular normalizing flow, directly modeling $p(x|y)$ with a conditional flow has a few benefits.
Training is typically faster, since the latent distribution has a smaller number of dimensions.
You can also draw samples of $x$ at fixed values of the conditions $y$, and you can calculate $p(x|y)$ without having to numerically calculate and divide by $p(y)$, which can be computationally expensive.

\subsection{Flows with periodic topology}
\label{sec:periodic}

The flows we have considered so far model data that live in $\R^n$.
This assumption is insufficient for modeling variables from spaces with non-Euclidean topology, e.g. positions on the sky.
While progress has been made on building flows for general topologies (e.g. \citealt{gemici2016} and \citealt{falorsi2019}), we will focus on building flows on the sphere, $S^2$, as this is the case most relevant in astronomy.
We will see that by carefully choosing the latent space, we can construct flows with periodic topology with minimal additional effort \citep{rezende2020}.

Positions on the sphere are specified by two angles\footnote{
We use the physicists' convention where $\theta$ and $\phi$ are the zenith and azimuthal angles, respectively.
},
$\theta$ and $\phi$, the latter of which is periodic.
By mapping $\theta$ to $\cos\theta$, we map the sphere to a cylinder\footnote{
This map can be explicitly constructed via an embedding in $\R^3$.
Technically, the map is not defined for $\theta \in \{0, \pi\}$, however as this set has zero measure, it can be safely ignored.}:
$S^2 \to [-1,1] \times S^1$ (i.e. the Cartesian product of an interval and a circle).
In other words, we can transform $\cos\theta$ with a Euclidean flow, as long as we ensure that the flow bounds samples to the range $[-1, 1]$.
However, the $S^1$ piece, $\phi$, has a periodic topology and must be handled more carefully.

First, we will address transformations of $\cos\theta$.
The only constraint we must impose is that samples of $\cos\theta$ must lie in the range $[-1, 1]$.
Fortunately, RQ-NSCs are bounded, mapping a range in $u$ to the same range in $x$.
Thus, if we pick a latent distribution with compact support in $[-1, 1]$, samples of $\cos\theta$ are guaranteed to lie in the same range, as long as we set the range of the RQ-NSC $B$ = 1.

Next we will address transformations of $\phi$.
For $f$ to be a valid diffeomorphism on the circle, $S^1$, it is sufficient that $f$ obey the following constraints:
\begin{align}
    f(0) &= 0 \\
    f(2\pi) &= 2\pi \\
    \nabla f(0) &= \nabla f(2\pi) \label{eq:df=df} \\
    \nabla f(\phi) &> 0.
\end{align}
The first two constraints ensure continuity of $f$ by designating $\phi=0$ as a fixed point, and the third constraint ensures continuity of $\nabla f$ at that fixed point.
While the designation of $\phi=0$ as a fixed point is an unnecessary restriction on $f$, any diffeomorphism on the circle has at least one fixed point up to a phase change, and so this restriction does not actually restrict the expressiveness of $f$.
The fourth restriction ensures monotonicity, which guarantees invertibility.

If we make the phase change $\phi \to \phi - \pi$ so that our angles $\phi \in [-\pi, \pi]$, a RQ-NSC with $B=\pi$ automatically fulfills all four constraints.
In fact, regular RQ-NSC's impose the further condition
\begin{align}
    \nabla f(-\pi) = \nabla f(\pi) = 1 \label{eq:df=1}
\end{align}
to match an identity transform for inputs outside of the range $[-\pi, \pi]$.
By choosing a latent distribution with compact support in the range $[-\pi, \pi]$, we ensure that no samples will lie outside the range of the splines, and so we can relax the boundary condition of Equation \ref{eq:df=1} in favor of the boundary condition in Equation \ref{eq:df=df}.
Spline transforms with this relaxed boundary condition are named \emph{Circular Splines} by \citet{rezende2020}.

See Figure \ref{fig:sphere} for an example of circular splines used to model the population density of the Earth \note{cite the data set here}.
The angles are shifted so that $\phi=\pm \pi$ corresponds to the Prime Meridian.
You can see that the circular splines have no problem with smoothly modeling the data around this discontinuity in $\phi$.

The circular spline construction above is easily generalized to n-spheres and n-tori: $S^n \to [-1, 1]^{n-1} \times S^1$ and $T^n \to (S^1)^n$ (see \citealt{rezende2020} for more details).
We can model the joint distribution of periodic and non-periodic variables with RQ-RSCs simply by choosing appropriate bounds $B$ for each dimension, and by swapping boundary condition \ref{eq:df=1} for condition \ref{eq:df=df} for any periodic dimensions.

\subsection{Modeling discrete variables}
\label{sec:discrete}

In addition to the continuous variables described above, normalizing flows can also be used to model discrete variables.
This can be achieved by ``dequantizing'' the discrete dimensions of the data, which can then be mapped onto continuous latent distributions using regular continuous bijectors.
Dequantization consists of adding some kind of continuous noise to the discrete dimensions, transforming them into continuous dimensions.
When sampling from the flow, you simply do the opposite, and ``quantize'' the discrete dimensions after applying all of the regular bijections, mapping the noisy, continuous variables onto their discrete counterparts.

A common method for dequantization is uniform dequantization, in which random uniform noise in the range (0, 1) is added to the discrete dimensions.
The corresponding quantization applied while sampling from the flow consists of applying the floor function to the dequantized dimensions, mapping these samples onto the nearest integer less than the sampled value.
More sophisticated dequantization schemes use variational inference or even another normalizing flow to determine the noise distributions, which improves results by smoothing the discontinuities between neighboring discrete values.
See \citet{ho2019} \citet{hoogeboom2020} for more details.

While the dequantizers are not technically bijections, they can be treated as data processing bijections and be chained together with the other bijectors in your normalizing flow.

\note{Try to get the discrete example working again! Add a figure here.}

\section{PZFlow}
\label{sec:pzflow}

PZFlow is a Python package for building normalizing flows, with a focus on easy high-performance modeling of high-dimensional tabular data.
Data is handled in Pandas DataFrames \citep{pandas}, while the normalizing flows are implemented in Jax \citep{jax}, which allows for efficient, parallelizable, GPU-enabled calculations for very large data sets.
The code is easily installable for the Python Package Index\footnote{\url{https://pypi.org/project/pzflow/}} (PyPI) and is hosted on Github,\footnote{\url{https://github.com/jfcrenshaw/pzflow}}.
The documentation\footnote{\note{URL}} includes tutorial notebooks covering all of the features demonstrated in this paper.

PZFlow is a package for building normalizing flows in Python, with a focus on high-performance out-of-the-box modeling of high dimensional tabular data.
Data is handled in Pandas DataFrames \citep{pandas} while the normalizing flows are implemented in Jax \citep{jax}, which allows efficient, parallelizable, GPU-enabled calculations for very large data sets.
The code is easily installable from the Python Package Index\footnote{\url{https://pypi.org/project/pzflow/}} (PyPI) and is hosted on Github,\footnote{\url{https://github.com/jfcrenshaw/pzflow}} including several tutorial notebooks.

The rest of this paper will demonstrate using PZFlow for the statistical modeling of galaxy catalogs.
Section \ref{sec:galaxy-catalog} uses PZFlow to forward model a photometric catalog, including spectroscopic redshifts, Rubin photometry, ``true'' photo-z posteriors, ellipticities, and sizes.
Section \ref{sec:photo-z} uses PZFlow for photo-z estimation, demonstrating the power of PZFlow as a density estimator, including numerous useful features for photo-z estimation.

In addition to the examples in this paper, PZFlow has already being used in various other projects:
\begin{itemize}
    \item \needscite used PZFlow to build a metric for observing strategy optimization based on information theory;
    \item \needscite used PZFlow to forward model galaxy data with true redshift posteriors in order to evaluate the impact of survey incompleteness and spec-z errors on photo-z estimation;
    \item \needscite used PZFlow to smooth high-redshift artifacts in simulations of supernova host galaxy simulations.
\end{itemize}

\section{Forward Modeling a Galaxy Catalog}
\label{sec:galaxy-catalog}

In this section, we apply PZFlow to forward model a photometric galaxy catalog for the Rubin Observatory.
The advantage of using a catalog generated from a normalizing flow is that we have direct access to the exact distribution from which the data is drawn, enabling us to calculate truth values for derived statistical products, such as the true photo-z redshift posterior for each galaxy.

In Section \ref{sec:fwd-model} we construct a normalizing flow to model the galaxy redshifts and photometry.
In Section \note{LABEL}, we calculate posteriors?
In Section \ref{sec:fwd-model-conditional} we build a conditional flow to add additional properties to the catalog.

\subsection{Forward modeling redshifts and photometry}
\label{sec:fwd-model}

\begin{figure*}[t!]
    \script{plot_main_galaxy_corner.py}
    \begin{centering}
        \includegraphics{figures/main_galaxy_corner.pdf}
        \caption{
            Distribution of the CosmoDC2 test set compared to the distribution learned by PZFlow.
            The close overlap of every pair-wise distribution demonstrates that PZFlow was able to learn the structure present in CosmoDC2 with high fidelity.
        }
        \label{fig:main-corner}
    \end{centering}
\end{figure*}

\begin{figure*}[t!]
    \script{plot_smooth_color_distribution.py}
    \begin{centering}
        \includegraphics{figures/smooth_color_distribution.pdf}
        \caption{
            Comparing the $r-i$ vs redshift distribution for galaxy samples from CosmoDC2 (left) and from the normalizing flow (right).
            The high-redshift galaxies in CosmoDC2 lie along discrete tracks in color space due to the discrete number of galaxy SED templates used in the simulation.
            PZFlow smooths over these discrete tracks, resulting in a color distribution that is smooth to high redshifts.
        }
        \label{fig:smooth-color-dist}
    \end{centering}
\end{figure*}

To create a generative model of galaxy redshifts and photometry, we use the true redshifts and $ugrizy$ magnitudes from the DESC CosmoDC2 simulation (\needscite).
We selected all galaxies from CosmoDC2 with at most one band beyond the LSST 10-year 5-sigma point source depths.
Of these, we selected $10^6$ galaxies and split them intro training and test sets consisting of 80\% and 20\% of the galaxies, respectively.

For the latent distribution we use a 7 dimensional Uniform distribution over the range $[-5, 5]$.
To map the data onto the latent distribution, we use the following bijection:
\begin{align}
    f = \text{RQ-RSC} \circ \text{Shift Bounds} \circ \text{Color Transform}.
\end{align}
\note{I might add an InverseSoftPlus}
We will explain each layer of the bijection in the order they are applied to the input data.

The first layer of the bijection is a Color Transform.
This bijection converts galaxy magnitudes to colors, but keeps the $i$ band magnitude as a proxy for the apparent luminosity:
\begin{multline}
    \text{Color Transform} : (\text{redshift},\, u,\, g,\, r,\, i,\, z,\, y) \to \\
    (\text{redshift},\, i,\, u-g,\, g-r,\, r-i,\, i-z,\, z-y).
\end{multline}
This layer is useful as galaxy redshifts correlate more directly with galaxy colors than galaxy magnitudes.

The next layer, Shift Bounds, is the data processing bijection defined in Section \ref{sec:data-processing}, which maps the range of the data onto the support of the RQ-RSC.
Note that since Shift Bounds is on the ``other side'' of the Color Transform, we need to map the ranges of the colors $u-g$, $g-r$, etc. onto the support of the splines, instead of the original magnitudes.

The final layer is an RQ-RSC, described in detail in Section \ref{sec:rq-rsc}.
This layer performs the heavy lifting of transforming the data distribution into the uniform latent distribution.
We use $D=7$ layers to transform all 7 dimensions of our data, and set $B=5$ to match the support of the latent distribution.
We use the coupling function (with two hidden layers of 128 neurons) described in \citet{durkan2019}.
We use $K=16$ spline knots.

We trained the normalizing flow for 150 epochs using the Adam optimizer \citep{adam} (with a learning rate of $10^{-3}$, that decreased by a factor of 10 every 50 epochs) to minimize the negative log-likelihood,
\begin{align}
    \mathcal{L} = - \, \mathbb{E}[ \, \log p(x) \, ],
\end{align}
where the expectation is performed over galaxies in the training set and $p(x)$ is defined in Equation \ref{eq:px}.
Training took 7 minutes on a Tesla P100 12GB GPU.

We assess the results by drawing $10^4$ galaxies from the trained flow, and plotting their distribution against  $10^4$ galaxies from the test set (Figure \ref{fig:main-corner}).
We see that the normalizing flow has done an excellent job of reproducing the distribution of DESC DC2, without any unusual artifacts or outliers.
In addition, Figure \ref{fig:smooth-color-dist} compares the distribution of galaxy $r-i$ vs redshift.
The CosmoDC2 simulation is known to exhibit discrete tracks in this space at high redshift, due to the discrete number of SED templates used during simulation.
These tracks are visible in the left panel.
The right panel shows that PZFlow smooths over this discreteness, resulting in a color distribution that is smooth up to high redshifts.

We note that these results were obtained without any extensive hyperparameter search, and that very similar (slightly worse results) are obtained without the \texttt{ColorTransform} bijection, demonstrating the flexibility of the method to adapt to unseen data sets.

With this normalizing flow, we have an efficient CosmoDC2 emulator that produces a smooth distribution of realistic galaxies up to high-redshifts.

To simulate photometric errors, we use the LSST error model of \citet{ivezic2019}, except without the high-SNR assumptions made there (see Appendix B for details).

\subsection{Calculating true posteriors}
\label{sec:true-posteriors}

Since we have direct access to the probability distribution from which the photometry and redshifts are drawn, we can calculate the true redshift posterior for each galaxy: $p(z|m)$ where $m$ is the vector of galaxy magnitudes.
We note that this is not an estimate, like what would be returned by a photo-z estimator, but rather the truth, obtained from the model that generated the photometry and redshifts in the first place.

In addition to calculating the true posterior, $p(z|m)$, we can calculate a true posterior that is consistent with the photometric errors:
\begin{align}
    p(z|m, \sigma_m) = \int p(z|\hat{m}) p(\hat{m}|m, \sigma_m) d\hat{m},
    \label{eq:err-conv}
\end{align}
where $\sigma_m$ is the vector of photometric errors returned by the error model, and $\hat{m}$ are possible noisy observations of the true magnitudes, $m$.
We can evaluate this integral numerically by sampling $\hat{m} \sim p(\hat{m}|m, \sigma_m)$, evaluating $p(z, \hat{m})$ on a redshift grid, and normalizing to obtain $p(z|\hat{m})$.
Averaging $p(z|\hat{m})$ over the samples $\hat{m}$ yields $p(z|m, \sigma_m)$.

We can also marginalize over any missing bands, $n$:
\begin{align}
    p(z|\hat{m}) = \frac{1}{p(\hat{m})} \int p(z, n, \hat{m}) dn,
\end{align}
which can be calculated by evaluating $p(z, n, \hat{m})$ on a grid of $z$ and $n$, summing over $n$ to yield $p(z, \hat{m})$, and normalizing with respect to redshift to yield $p(z|\hat{m})$.
You can once again average over $\hat{m}$ samples to convolve the photometric errors.
You may wish to marginalize over all values of $n$ if the galaxy was not observed in that band.
This might occur, for example, if you include photometry from Euclid, which will not have complete coverage of the LSST catalog \needscite.
You may also wish to marginalize over all values beyond a limiting magnitude if the galaxy was observed, but not detected in that band.
This might occur, for example, in the low wavelength bands of Lyman dropout galaxies observed by LSST.

\note{Figure of the posteriors.}

Figure \ref{fig:posteriors} shows a number of redshift posteriors to visualize the effects of error convolution and band marginalization.
The black posterior is calculated using the true galaxy magnitudes, while the blue posteriors is calculated after adding photometric errors.
Calculating the posterior using the noisy photometry results in a biased posterior.
The orange posterior has had the photometric errors convolved as in Equation \ref{eq:err-conv}.
Convolving the errors de-biases the posterior, but the resultant posterior is broader, reflecting the increase uncertainty due to the photometric errors.
Finally, the green posterior has had the $u$ band marginalized.
This posterior is even broader, with a significant low-redshift tail, demonstrating how the loss of information in the $u$ band translates to a loss of information about the redshift.
\note{WILL NEED TO UPDATE THIS AFTER MAKING THE NEW FIGURE.}

Calculating these posteriors enables direct comparison of true redshift posteriors (consistent with the photometric errors and missing bands) with the redshift posteriors estimated by photo-z estimators.
This allows you to test photo-z estimation on a posterior-per-galaxy basis, which will alleviate the photo-z validation problems discovered by \citet{schmidt2020}.

\note{Point towards the photo-z section for a more comprehensive discussion of posterior calculation.}

\subsection{Additional properties with conditional flows}
\label{sec:fwd-model-condtional}

\begin{figure*}[t!]
    \script{plot_conditional_galaxy_corner.py}
    \begin{centering}
        \includegraphics{figures/conditional_galaxy_corner.pdf}
        \caption{
            Distribution of the ellipticities and sizes of the galaxies in the CosmoDC2 test set compared to the distribution learned by PZFlow.
            The close overlap of every pair-wise distribution demonstrates that PZFlow was able to learn the structure present in CosmoDC2 with high fidelity.
        }
        \label{fig:conditional-corner}
    \end{centering}
\end{figure*}

In addition to the galaxy magnitudes and redshifts modeled above, we wish to include other galaxy properties in the catalog, such as galaxy size and ellipticity.
In principle, we could have included these variables in the original normalizing flow.
However, we did not want the true redshift posteriors to be conditioned on these variables, as most photo-z estimators only use galaxy photometry.
Therefore, we will build a second flow that models these additional values conditioned on the galaxy redshift and magnitudes.
Note that while we have only chosen to model these additional two properties, any other values you desire can be similarly modeled.

For the latent distribution, we once more use a Uniform distribution over the range $[-5, 5]$.
For the bijection, we use
\begin{align}
    \begin{split}
        f =& ~ \text{RQ-RSC} \circ \text{Shift Bounds}. \\
    \end{split}
\end{align}
The RQ-RSC acts on the two dimensional space of size and ellipticity, but also takes the galaxy redshift and magnitudes as inputs.
The redshifts and magnitudes are transformed to have zero mean and unit variance before being input to the neural network\footnote{These variables are standard scaled instead of mapped onto the domain [-5, 5], because the neural network that parameterizes the splines has no limit on inputs, unlike the splines themselves, which are limited to the range [-5, 5].} that parameterizes the splines.
Aside from the change in inputs, the RQ-RSC has the same settings as listed for the previous normalizing flow.

The conditional flow is trained for \note{N} epochs using the same loss function and optimizer as before.
\note{is the LR the same?}
We sample a size and ellipticity for each galaxy in the PZFlow catalog created in the previous section (using the true magnitudes), and plot the distribution of these features against the distribution in the test set (Figure \note{N}).
Once again, we see that the normalizing flow does a good job of emulating the CosmoDC2 distribution.

The final simulated catalog consists of $10^4$ galaxies, each with a redshift, $ugrizy$ magnitudes including photometric errors, a true photo-z posterior consistent with the photometric errors, a size, and an ellipticity.
This small catalog was generated for visualization purposes, but the normalizing flows can be used to generate catalogs of arbitrarly large size (see Appendix C on benchmarking PZFlow).

\section{Photometric Redshift Estimation}
\label{sec:photo-z}

\subsection{Training an Ensemble for photo-z estimation}

\note{
    Galaxy catalog created in the previous section that has true posteriors. \\
    Ensembles. \\
    Training with error sampling.
}

\subsection{Estimating posteriors}

\note{
    Estimating raw posteriors. \\
    Estimating posteriors with error convolution. \\
    Estimating posteriors with u band marginalization.
}

\subsection{Photo-z metrics}

\note{
    Photo-z point plot. \\
    Binned point metrics vs redshift. \\
    Brier score and other PDF metrics.
}

\appendix

\section{Training details}

\begin{figure*}[t!]
    \script{plot_galaxy_losses.py}
    \begin{centering}
        \includegraphics{figures/galaxy_flow_losses.pdf}
        \caption{
            Training losses for the galaxy flows.
        }
        \label{fig:galaxy-flow-losses}
    \end{centering}
\end{figure*}

\section{LSST Error Model}

- Explain how I adapted the LSST Error Model to work for low-SNR galaxies

\section{Failed experiments}

- CentBeta - harder to train, slower, results aren't better
- convolving Gaussian errors analytically (use the trained flow from above to plot the gradients of the spline or something and show that the necessary assumptions are not valid)

\section{Benchmarking PZFlow}

\bibliography{references.bib}

\end{document}

### 2022-06-06 src/scripts/plot_smooth_color_distribution.py

@@ -1,0 +1,54 @@
+import matplotlib.pyplot as plt
+import pandas as pd
+import paths
+from pzflow import Flow
+
+# load the cosmoDC2 data
+data = pd.read_pickle(paths.data / "cosmoDC2_subset.pkl")
+
+# load the flow
+flow = Flow(file=paths.data / "main_galaxy_flow" / "flow.pzflow.pkl")
+
+# pull out a subset of cosmoDC2 for plotting
+data = data.iloc[:100_000]
+
+# draw a same-sized sample from the flow
+samples = flow.sample(data.shape[0], seed=0)
+
+# create the figure
+fig, (ax1, ax2) = plt.subplots(1, 2, figsize=(6, 2.7), constrained_layout=True, dpi=120)
+
+# set global plot settings
+plot_settings = {
+    "bins": 512,
+    "cmap": "ocean_r",
+    "rasterized": True,
+}
+
+ax1.hist2d(data["redshift"], data["r"] - data["i"], **plot_settings)
+ax2.hist2d(samples["redshift"], samples["r"] - samples["i"], **plot_settings)
+
+# set global axis settings
+ax_settings = {
+    "xlim": (0, 2.5),
+    "ylim": (-0.25, 1.3),
+    "xlabel": "redshift",
+}
+
+# set axis settings
+ax1.set(ylabel="$r - i$", **ax_settings)
+ax2.set(ylabel=" ", **ax_settings)
+
+
+# set the labels
+text_settings = {
+    "x": 0.95,
+    "y": 0.95,
+    "va": "top",
+    "ha": "right",
+}
+ax1.text(s="CosmoDC2", transform=ax1.transAxes, **text_settings)
+ax2.text(s="PZFlow", transform=ax2.transAxes, **text_settings)
+
+# save the figure!
+fig.savefig(paths.figures / "smooth_color_distribution.pdf", dpi=300)

### 2022-06-06 src/scripts/lsst_error_model.py
Added script to generate the pzflow catalog.

@@ -1,0 +1,732 @@
+"""Calculate photometric errors using the LSST Error Model.
+
+See the docstring below for details.
+"""
+from numbers import Number
+from typing import Iterable, List, Optional, Tuple
+
+import numpy as np
+import pandas as pd
+
+
+class LSSTErrorModel:
+    """LSST Model for photometric errors.
+
+    Implements the error model from the LSST Overview Paper:
+    https://arxiv.org/abs/0805.2366
+    Note however that this paper gives the high SNR approximation.
+    By default, this model uses the more accurate version of the error model
+    where Eq. 5 = (N/S)^2, in flux, and the error is Gaussian in flux space.
+    There is a flag allowing you to use the high SNR approximation instead.
+    See the __init__ docstring.
+
+    Create an instance by calling the class, then use the instance as a
+    callable on pandas DataFrames.
+    Example usage:
+    errModel = LSSTErrorModel()
+    data_with_errs = errModel(data)
+    """
+
+    def __init__(
+        self,
+        bandNames: dict = None,
+        tvis: float = None,
+        nYrObs: float = None,
+        nVisYr: dict = None,
+        gamma: dict = None,
+        airmass: float = None,
+        extendedSource: float = None,
+        sigmaSys: float = None,
+        magLim: float = None,
+        ndFlag: float = None,
+        m5: dict = None,
+        Cm: dict = None,
+        msky: dict = None,
+        theta: dict = None,
+        km: dict = None,
+        highSNR: bool = None,
+    ):
+        """Error model from the LSST Overview Paper: https://arxiv.org/abs/0805.2366
+
+        All parameters are optional. To see the default settings, do
+        `LSSTErrorModel().default_settings()`
+
+        By default, this model uses the more accurate version of the error
+        model. See the explanations in the class docstring and the description
+        for highSNR below.
+
+        Note that the dictionary bandNames sets the bands for which this model
+        calculates photometric errors. The dictionary keys are the band names
+        that the error model uses internally to search for parameters, and the
+        corresponding dictionary values are the band names as they appear in
+        your data set.
+
+        By default, the LSST bands are named "u", "g", "r", "i", "z", and "y".
+        You can use the bandNames dictionary to alias them differently.
+
+        For example, if in your DataFrame, the bands are named lsst_u, lsst_g, etc.
+        you can set bandNames = {"u": "lsst_u", "g": "lsst_g", ...}, and the error
+        model will work automatically.
+
+        You can also add other bands to bandNames. For example, if you want to
[…]

### 2022-06-06 src/scripts/plot_smooth_color_distribution.py

@@ -1,3 +1,4 @@
+"""Plot distribution of r-i vs redshift for CosmoDC2 and PZFlow."""
 import matplotlib.pyplot as plt
 import pandas as pd
 import paths

### 2022-06-06 src/scripts/simulate_pzflow_catalog.py

@@ -1,0 +1,43 @@
+import pandas as pd
+import paths
+from lsst_error_model import LSSTErrorModel
+from pzflow import Flow
+
+# generate true photometry (+ redshift)
+flow = Flow(file=paths.data / "main_galaxy_flow" / "flow.pzflow.pkl")
+truth_sample = flow.sample(1_000_000, seed=0)
+
+# generate true ellipticity and size
+ellip_and_size = cond_flow.sample(
+    nsamples=1, conditions=truth_sample, save_conditions=False, seed=1
+)
+
+# simulate observations with the fiducial LSST error model
+err_model = LSSTErrorModel()
+obs_sample = err_model(truth_sample[list("ugrizy")], seed=2)
+
+# relabel the true and observed values
+truth_sample = truth_sample.rename(
+    columns={col: f"true_{col}" for col in truth_sample.columns}
+)
+ellip_and_size = ellip_and_size.rename(
+    columns={col: f"true_{col}" for col in ellip_and_size.columns}
+)
+obs_sample = obs_sample.rename(
+    columns={col: f"obs_{col}" for col in obs_sample.columns}
+)
+
+# combine the two catalogs into a single DataFrame
+pzflow_catalog = pd.concat((truth_sample, obs_sample, ellip_and_size), axis=1)
+pzflow_catalog = pzflow_catalog[
+    ["true_redshift"]
+    + [
+        col
+        for band in list("ugrizy")
+        for col in [f"true_{band}", f"obs_{band}", f"obs_{band}_err"]
+    ]
+    + ["true_ellipticity", "true_size"]
+]
+
+# save the catalog!
+pzflow_catalog.to_pickle(paths.data / "pzflow_catalog.pkl")

### 2022-06-06 src/scripts/simulate_pzflow_catalog.py
Added pzflow_catalog to snakefile and fixed conditional flow loading.

@@ -1,3 +1,4 @@
+"""Simulate a PZFlow-generated catalog of 1e6 galaxies."""
 import pandas as pd
 import paths
 from lsst_error_model import LSSTErrorModel
@@ -8,6 +9,7 @@ flow = Flow(file=paths.data / "main_galaxy_flow" / "flow.pzflow.pkl")
 truth_sample = flow.sample(1_000_000, seed=0)
 
 # generate true ellipticity and size
+cond_flow = Flow(file=paths.data / "conditional_galaxy_flow" / "flow.pzflow.pkl")
 ellip_and_size = cond_flow.sample(
     nsamples=1, conditions=truth_sample, save_conditions=False, seed=1
 )

### 2022-06-07 setup.py
Moved paths and LSST err model to utils. Changed script folder structure.

@@ -1,0 +1,12 @@
+from pathlib import Path
+
+from setuptools import setup
+
+# create the hidden folder that holds the editable utils install
+Path(".utils.egg-info").mkdir(exist_ok=True)
+
+# install the utils package
+setup(
+    name="utils",
+    package_dir={"": "src/utils"},
+)

### 2022-06-07 src/utils/paths.py

@@ -1,8 +1,11 @@
 """Exposes common paths for manipulating datasets and generating figures."""
 from pathlib import Path
 
+import git
+
 # Absolute path to the top level of the repository
-root = Path(__file__).resolve().parents[2].absolute()
+repo = git.Repo(".", search_parent_directories=True)
+root = Path(repo.working_tree_dir).resolve().absolute()
 
 # Absolute path to the `src` folder
 src = root / "src"

### 2022-06-07 setup.py
Fixed modular script setup.

@@ -1,12 +1,0 @@
-from pathlib import Path
-
-from setuptools import setup
-
-# create the hidden folder that holds the editable utils install
-Path(".utils.egg-info").mkdir(exist_ok=True)
-
-# install the utils package
-setup(
-    name="utils",
-    package_dir={"": "src/utils"},
-)

### 2022-06-07 src/scripts/forward_model/plot_galaxy_losses.py

@@ -1,7 +1,10 @@
 """Plot the training losses for the galaxy flows."""
 import matplotlib.pyplot as plt
 import numpy as np
-import paths
+from showyourwork.paths import user as Paths
+
+# instantiate the paths
+paths = Paths()
 
 # load the losses
 main_losses = np.load(paths.data / "main_galaxy_flow" / "losses.npy")

### 2022-06-07 src/scripts/forward_model/plot_main_galaxy_corner.py

@@ -3,9 +3,12 @@ import corner
 import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
-import paths
 from matplotlib.ticker import MaxNLocator
 from pzflow import Flow
+from showyourwork.paths import user as Paths
+
+# instantiate the paths
+paths = Paths()
 
 # set the number of samples to plot
 n_samples = 10_000

### 2022-06-07 src/scripts/forward_model/plot_smooth_color_distribution.py

@@ -1,8 +1,11 @@
 """Plot distribution of r-i vs redshift for CosmoDC2 and PZFlow."""
 import matplotlib.pyplot as plt
 import pandas as pd
-import paths
 from pzflow import Flow
+from showyourwork.paths import user as Paths
+
+# instantiate the paths
+paths = Paths()
 
 # load the cosmoDC2 data
 data = pd.read_pickle(paths.data / "cosmoDC2_subset.pkl")

### 2022-06-07 src/scripts/forward_model/simulate_pzflow_catalog.py

@@ -1,8 +1,11 @@
 """Simulate a PZFlow-generated catalog of 1e6 galaxies."""
 import pandas as pd
-import paths
 from lsst_error_model import LSSTErrorModel
 from pzflow import Flow
+from showyourwork.paths import user as Paths
+
+# instantiate the paths
+paths = Paths()
 
 # generate true photometry (+ redshift)
 flow = Flow(file=paths.data / "main_galaxy_flow" / "flow.pzflow.pkl")

### 2022-06-07 src/scripts/forward_model/train_conditional_galaxy_flow.py

@@ -4,9 +4,12 @@ from pathlib import Path
 import numpy as np
 import optax
 import pandas as pd
-import paths
 from pzflow import Flow
 from pzflow.bijectors import Chain, RollingSplineCoupling, ShiftBounds
+from showyourwork.paths import user as Paths
+
+# instantiate the paths
+paths = Paths()
 
 # load the training data
 data = pd.read_pickle(paths.data / "cosmoDC2_subset.pkl")

### 2022-06-07 src/scripts/forward_model/train_main_galaxy_flow.py

@@ -4,9 +4,12 @@ from pathlib import Path
 import numpy as np
 import optax
 import pandas as pd
-import paths
 from pzflow import Flow
 from pzflow.bijectors import Chain, ColorTransform, RollingSplineCoupling, ShiftBounds
+from showyourwork.paths import user as Paths
+
+# instantiate the paths
+paths = Paths()
 
 # load the training data
 data = pd.read_pickle(paths.data / "cosmoDC2_subset.pkl")

### 2022-06-07 src/scripts/intro/plot_two_moons.py

@@ -1,10 +1,13 @@
 """Create the two moons example figure."""
 import matplotlib.pyplot as plt
 import numpy as np
-import paths
 from matplotlib.gridspec import GridSpec
 from pzflow import Flow
 from pzflow.examples import get_twomoons_data
+from showyourwork.paths import user as Paths
+
+# instantiate the paths
+paths = Paths()
 
 # load the two moons example data
 data = get_twomoons_data()

### 2022-06-07 src/scripts/intro/train_two_moons_flow.py

@@ -1,7 +1,10 @@
 """Train the flow for the two moons example."""
-import paths
 from pzflow import Flow
 from pzflow.examples import get_twomoons_data
+from showyourwork.paths import user as Paths
+
+# instantiate the paths
+paths = Paths()
 
 # load the two moons example data
 data = get_twomoons_data()

### 2022-06-07 src/utils/paths.py

@@ -1,26 +1,0 @@
-"""Exposes common paths for manipulating datasets and generating figures."""
-from pathlib import Path
-
-import git
-
-# Absolute path to the top level of the repository
-repo = git.Repo(".", search_parent_directories=True)
-root = Path(repo.working_tree_dir).resolve().absolute()
-
-# Absolute path to the `src` folder
-src = root / "src"
-
-# Absolute path to the `src/data` folder (contains datasets)
-data = src / "data"
-
-# Absolute path to the `src/static` folder (contains static images)
-static = src / "static"
-
-# Absolute path to the `src/scripts` folder (contains figure/pipeline scripts)
-scripts = src / "scripts"
-
-# Absolute path to the `src/tex` folder (contains the manuscript)
-tex = src / "tex"
-
-# Absolute path to the `src/tex/figures` folder (contains figure output)
-figures = tex / "figures"

### 2022-06-07 src/scripts/forward_model/plot_conditional_galaxy_corner.py
updated ticks and ranges on corner plots.

@@ -38,7 +38,7 @@ corner_settings = {
         (18, 27),
         (18, 27),
         (18, 27),
-        (-0.1, 1.07),
+        (0, 1.07),
         (0, 1.07),
     ],
     "hist_bin_factor": 1,
@@ -62,9 +62,8 @@ for ax in axes[:-2].flatten():
     ax.set_visible(False)
 
 # set redshift and magnitude ticks to integers
-for row in axes[-2:]:
-    for j in range(7):
-        row[j].xaxis.set_major_locator(MaxNLocator(4, integer=True))
+for ax in axes[-2:, :-2].flatten():
+    ax.xaxis.set_major_locator(MaxNLocator(4, integer=True))
 
 # set ellipticity and size ticks
 for ax in [axes[-2, 0], axes[-1, 0]]:
@@ -72,8 +71,16 @@ for ax in [axes[-2, 0], axes[-1, 0]]:
 for ax in [axes[-1, -2], axes[-1, -1]]:
     ax.set(xticks=[0, 0.5, 1.0])
 
-# set ylim on first row
+# remove interior ticks
+for ax in axes[-2, :]:
+    ax.set(xticks=[])
+for ax in axes[-2:, 1:].flatten():
+    ax.set(yticks=[])
+
+# set ylim
 for ax in axes[-2, :-2]:
+    ax.set(ylim=(0, 1.03))
+for ax in axes[-1, :-1]:
     ax.set(ylim=(0, 1.03))
 
 # add a legend

### 2022-06-07 src/scripts/forward_model/plot_main_galaxy_corner.py

@@ -40,7 +40,6 @@ corner_settings = {
     ],
     "hist_bin_factor": 1,
     "labels": test_set.columns,
-    "labelpad": 0.05,
 }
 
 # plot the test set in red
@@ -51,13 +50,18 @@ corner.corner(
     samples.to_numpy(), color="C0", data_kwargs={"ms": 1.5}, **corner_settings
 )
 
-# set ticks to integers
+# set border ticks to integers
 axes = np.array(fig.axes).reshape((7, 7))
-for i in range(1, 7):
-    for j in range(i + 1):
-        axes[i, j].xaxis.set_major_locator(MaxNLocator(4, integer=True))
-        if j < i:
-            axes[i, j].yaxis.set_major_locator(MaxNLocator(4, integer=True))
+for ax in axes[-1, :]:
+    ax.xaxis.set_major_locator(MaxNLocator(4, integer=True))
+for ax in axes[1:, 0]:
+    ax.yaxis.set_major_locator(MaxNLocator(4, integer=True))
+
+# remove interior ticks
+for ax in axes[:-1, :].flatten():
+    ax.set(xticks=[])
+for ax in axes[:, 1:].flatten():
+    ax.set(yticks=[])
 
 # add a legend
 axes[2, 5].plot([], c="C0", label="PZFlow")

### 2022-06-15 src/scripts/forward_model/plot_posteriors.py
Added script to plot galaxy posteriors.

@@ -1,0 +1,105 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+from pzflow import Flow
+from showyourwork.paths import user as Paths
+
+# instantiate the paths
+paths = Paths()
+
+# load the PZFlow generated catalog
+pzflow_catalog = pd.read_pickle(paths.data / "pzflow_catalog.pkl")
+
+# split into truth and observed catalogs
+truth_catalog = pzflow_catalog.filter(like="true_")
+truth_catalog.columns = truth_catalog.columns.str.removeprefix("true_")
+
+obs_catalog = pzflow_catalog.filter(like="obs_")
+obs_catalog.columns = obs_catalog.columns.str.removeprefix("obs_")
+
+no_u_catalog = obs_catalog.copy()
+no_u_catalog["u"] = np.nan
+no_u_catalog = no_u_catalog.drop("u_err", axis=1)
+
+# for nice visualization, let's pull out number 77
+idx = [77]
+truth_catalog = truth_catalog.iloc[idx]
+obs_catalog = obs_catalog.iloc[idx]
+no_u_catalog = no_u_catalog.iloc[idx]
+
+# and let's decrease the u error by factor of 10
+obs_catalog["u_err"] /= 10
+
+# load the flow for posterior estimation
+flow = Flow(file=paths.data / "main_galaxy_flow" / "flow.pzflow.pkl")
+
+# set the redshift grid
+grid = np.linspace(0.1, 0.6, 100)
+
+# calculate posteriors using the true magnitudes
+pdfs_true_mags = flow.posterior(truth_catalog, column="redshift", grid=grid)
+
+# calculate posteriors using the observed magnitudes
+pdfs_obs_mags = flow.posterior(obs_catalog, column="redshift", grid=grid)
+
+# calculate posteriors using error convolution
+pdfs_conv = flow.posterior(
+    obs_catalog, column="redshift", grid=grid, err_samples=10_000, seed=0
+)
+
+# calculate posteriors with u band marginalization
+marg_rules = {
+    "flag": np.nan,
+    "u": lambda row: np.linspace(24, 28, 20),
+}
+pdfs_no_u = flow.posterior(
+    no_u_catalog,
+    column="redshift",
+    grid=grid,
+    err_samples=10_000,
+    seed=0,
+    marg_rules=marg_rules,
+)
+
+# now create the figure
+# the second row is for putting the legend below the plot
[…]

### 2022-06-15 src/tex/ms.tex
Moved paths and LSST err model to utils. Changed script folder structure.
Fixed modular script setup.
Added files missing from last commit

@@
 \begin{figure*}[t!]
-    \script{plot_two_moons.py}
+    \script{intro/plot_two_moons.py}
     \begin{centering}
@@
 \begin{figure*}[t!]
-    \script{plot_main_galaxy_corner.py}
+    \script{forward_model/plot_main_galaxy_corner.py}
     \begin{centering}
@@
 \begin{figure*}[t!]
-    \script{plot_smooth_color_distribution.py}
+    \script{forward_model/plot_smooth_color_distribution.py}
     \begin{centering}
@@
 
-\note{Figure of the posteriors.}
+\begin{figure}[t!]
+    \script{forward_model/plot_posteriors.py}
+    \begin{centering}
+        \includegraphics{figures/posteriors.pdf}
+        \caption{
+            True redshift posteriors for a galaxy representing different amounts of information.
+            The black posterior is calculated with the true magnitudes;
+            the blue posterior is calculated after adding photometric errors;
+            the orange posterior is calculated after adding photometric errors, but with the errors convolved during posterior calculation;
+            the green posteriors is the same as the orange, except with the $u$ band marginalized over.
+        }
+        \label{fig:posteriors}
+    \end{centering}
+\end{figure}
 
@@
 \begin{figure*}[t!]
-    \script{plot_conditional_galaxy_corner.py}
+    \script{forward_model/plot_conditional_galaxy_corner.py}
     \begin{centering}
@@
 \begin{figure*}[t!]
-    \script{plot_galaxy_losses.py}
+    \script{forward_model/plot_galaxy_losses.py}
     \begin{centering}

### 2022-06-21 src/tex/ms.tex
Updated language about posterior plot.

@@
 The black posterior is calculated using the true galaxy magnitudes, while the blue posteriors is calculated after adding photometric errors.
-Calculating the posterior using the noisy photometry results in a biased posterior.
+Calculating the posterior using the noisy photometry results in a posterior with very little support over the true redshift.
 The orange posterior has had the photometric errors convolved as in Equation \ref{eq:err-conv}.
-Convolving the errors de-biases the posterior, but the resultant posterior is broader, reflecting the increase uncertainty due to the photometric errors.
+Convolving the errors broadens the posterior so that the true redshift lies within the support of the posterior.
+This broadening reflects the increased uncertainty due to the photometric errors.
 Finally, the green posterior has had the $u$ band marginalized.
-This posterior is even broader, with a significant low-redshift tail, demonstrating how the loss of information in the $u$ band translates to a loss of information about the redshift.
-\note{WILL NEED TO UPDATE THIS AFTER MAKING THE NEW FIGURE.}
+This posterior favors lower redshifts, although has a high-redshift tail that covers the support of the previous posteriors. 
+This demonstrates how the $u$ band helps constrain the redshift, and the loss of this information leads to greater uncertainty.
 
 Calculating these posteriors enables direct comparison of true redshift posteriors (consistent with the photometric errors and missing bands) with the redshift posteriors estimated by photo-z estimators.
-This allows you to test photo-z estimation on a posterior-per-galaxy basis, which will alleviate the photo-z validation problems discovered by \citet{schmidt2020}.
-
-\note{Point towards the photo-z section for a more comprehensive discussion of posterior calculation.}
+This allows you to test photo-z estimation on a posterior-per-galaxy basis, which will alleviate the photo-z validation problems discovered by \citet{schmidt2020} and will be investigated in future DESC papers.
 

### 2022-07-20 src/scripts/photo-z/calculate_posteriors.py
Added beginning of photo-z work.

@@ -1,0 +1,29 @@
+"""Use the pz flow ensemble to estimate redshift posteriors for test set galaxies."""
+import numpy as np
+from load_pzflow_catalog import load_pzflow_catalog
+from pzflow import FlowEnsemble
+from showyourwork.paths import user as Paths
+
+# instantiate the paths
+paths = Paths()
+
+# load the flow ensemble
+flowEns = FlowEnsemble(file=paths.data / "pz_ensemble" / "pz_ensemble.pzflow.pkl")
+
+# load the test set
+test_set = load_pzflow_catalog(subset="test")
+
+# setup the redshift grid
+grid = np.linspace(0, 3.5, 351)
+
+# calculate posteriors
+pdfs = flowEns.posterior(
+    test_set[:1000],  # remove this for final paper!
+    "redshift",
+    grid,
+    err_samples=10,
+    batch_size=100,
+)
+
+# save the posteriors
+np.savez(paths.data / "redshift_posteriors.npz", grid=grid, pdfs=pdfs)

### 2022-07-20 src/scripts/photo-z/load_pzflow_catalog.py

@@ -1,0 +1,61 @@
+"""Load the PZFlow catalog created by the forward model."""
+import pandas as pd
+from showyourwork.paths import user as Paths
+
+
+def load_pzflow_catalog(
+    truth: bool = False,
+    subset: str = None,
+    train_frac: float = 0.8,
+    include_nans: bool = False,
+) -> pd.DataFrame:
+    """Load the catalog created by the PZFlow forward model.
+
+    Parameters
+    ----------
+    truth: bool, default=False
+        Whether to return the truth catalog. If False, the observed catalog is
+        returned. Note that the true redshifts are returned in both cases.
+    subset: str, default=None
+        If `train` or `test`, this function returns the corresponding sets. If
+        None, the full set is returned.
+    train_frac: float, default=0.8
+        The fraction of the catalog to include in the train set.
+        1 - train_frac is the fraction included in the test set.
+    include_nans: bool, default=False
+        Whether to include galaxies that were not observed in every band.
+    """
+    # instantiate the paths
+    paths = Paths()
+
+    # load the training data
+    data = pd.read_pickle(paths.data / "pzflow_catalog.pkl")
+
+    # rename the columns
+    col_names = {"true_redshift": "redshift"}
+    if truth:
+        col_names = col_names | {f"true_{band}": band for band in "ugrizy"}
+    else:
+        col_names = (
+            col_names
+            | {f"obs_{band}": band for band in "ugrizy"}
+            | {f"obs_{band}_err": f"{band}_err" for band in "ugrizy"}
+        )
+
+    data = data.rename(columns=col_names)[list(col_names.values())]
+
+    # prune the columns
+    data = data[list(col_names.values())]
+
+    # select the subset
+    split_idx = int(train_frac * len(data))
+    if subset == "train":
+        data = data.iloc[:split_idx]
+    elif subset == "test":
+        data = data.iloc[split_idx:]
+
+    if not include_nans:
+        # get the rows without NaNs
+        data = data[~data.isnull().any(axis=1)]
+
+    return data

### 2022-07-20 src/scripts/photo-z/plot_ensemble_posteriors.py

@@ -1,0 +1,51 @@
+"""Plot posterior ensembles for a few galaxies."""
+import matplotlib.pyplot as plt
+import numpy as np
+from load_pzflow_catalog import load_pzflow_catalog
+from pzflow import FlowEnsemble
+from showyourwork.paths import user as Paths
+
+# instantiate the paths
+paths = Paths()
+
+# load the flow ensemble
+flowEns = FlowEnsemble(file=paths.data / "pz_ensemble" / "pz_ensemble.pzflow.pkl")
+
+# load the test set
+test_set = load_pzflow_catalog(subset="test")
+
+# select galaxies for plotting
+idx = [21, 54, 99]
+test_set = test_set.iloc[idx]
+
+# calculate posteriors
+grid = np.linspace(0, 3.5, 351)
+pdf_ens = flowEns.posterior(
+    test_set, "redshift", grid, returnEnsemble=True, err_samples=100, seed=0
+)
+pdf_mean = flowEns.posterior(
+    test_set, "redshift", grid, returnEnsemble=False, err_samples=100, seed=0
+)
+
+# create the figure
+fig, axes = plt.subplots(1, 3, figsize=(7, 2.3), constrained_layout=True, dpi=120)
+
+# loop through the panels
+for i, ax in enumerate(axes):
+    # plot the ensemble of pdfs
+    for j, pdf in enumerate(pdf_ens[i]):
+        ax.plot(grid, pdf, label=f"Flow {j + 1}")
+
+    # plot the ensemble mean
+    ax.plot(grid, pdf_mean[i], c="k", ls="--", label="Mean")
+
+    # plot the true redshift
+    ax.axvline(test_set.iloc[i]["redshift"], c="gray", label="Truth")
+
+    # set labels and limits
+    ax.set(yticks=[], xlabel="redshift", xlim=(grid.min(), grid.max()))
+
+# put a legend in the center
+axes[1].legend(loc="upper left")
+
+fig.savefig(paths.figures / "ensemble_posteriors.pdf")

### 2022-07-20 src/scripts/photo-z/plot_pz_ensemble_corner.py

@@ -1,0 +1,69 @@
+"""Make the corner plot for the pz estimating ensemble."""
+import corner
+import matplotlib.pyplot as plt
+import numpy as np
+from load_pzflow_catalog import load_pzflow_catalog
+from matplotlib.ticker import MaxNLocator
+from pzflow import FlowEnsemble
+from showyourwork.paths import user as Paths
+
+# instantiate the paths
+paths = Paths()
+
+# load the flow ensemble
+flowEns = FlowEnsemble(file=paths.data / "pz_ensemble" / "pz_ensemble.pzflow.pkl")
+
+# load the test set
+test_set = load_pzflow_catalog(subset="test").iloc[:10_000]
+
+# sample from the ensemble
+samples = flowEns.sample(test_set.shape[0], seed=0)
+
+# create the corner plot
+fig = plt.figure(figsize=(7.1, 7.1))
+
+# some global corner settings
+corner_settings = {
+    "fig": fig,
+    "bins": 20,
+    "range": [
+        (-0.1, 3),
+        (20, 29.5),
+        (20, 27.9),
+        (19, 27.5),
+        (18, 27),
+        (18, 27),
+        (18, 27),
+    ],
+    "hist_bin_factor": 1,
+    "labels": test_set.columns,
+}
+
+# plot the test set in red
+corner.corner(test_set.to_numpy(), color="C3", **corner_settings)
+
+# plot the pz estimator samples in blue
+corner.corner(
+    samples.to_numpy(), color="C0", data_kwargs={"ms": 1.5}, **corner_settings
+)
+
+# set border ticks to integers
+axes = np.array(fig.axes).reshape((7, 7))
+for ax in axes[-1, :]:
+    ax.xaxis.set_major_locator(MaxNLocator(4, integer=True))
+for ax in axes[1:, 0]:
+    ax.yaxis.set_major_locator(MaxNLocator(4, integer=True))
+
+# remove interior ticks
+for ax in axes[:-1, :].flatten():
+    ax.set(xticks=[])
+for ax in axes[:, 1:].flatten():
+    ax.set(yticks=[])
+
+# add a legend
+axes[2, 5].plot([], c="C0", label="PZ Estimator Model")
+axes[2, 5].plot([], c="C3", label="PZFlow Catalog")
+axes[2, 5].legend()
+
+# save the figure
+fig.savefig(paths.figures / "pz_estimator_corner.pdf", dpi=600, bbox_inches="tight")

### 2022-07-20 src/scripts/photo-z/train_pz_ensemble.py

@@ -1,0 +1,161 @@
+"""Train the flow ensemble that estimates photometric redshifts."""
+import pickle
+from pathlib import Path
+
+import jax.numpy as jnp
+import numpy as np
+import optax
+from jax import random
+from load_pzflow_catalog import load_pzflow_catalog
+from pzflow import FlowEnsemble
+from pzflow.bijectors import Chain, ColorTransform, RollingSplineCoupling, ShiftBounds
+from showyourwork.paths import user as Paths
+
+"""
+TO DO:
+------
+- fix the NaN issue that requires such a big buffer on the ShiftBounds range.
+- look at the losses. Make sure this training scheme is good. Also figure out a
+    good way to concatenate the losses
+"""
+
+# instantiate the paths
+paths = Paths()
+
+# load the training set of galaxies from the PZFlow catalog
+train_set = load_pzflow_catalog(subset="train")
+
+# set up the bijector
+# we will use the same one used for the forward model
+
+# the first bijector is the color transform
+# we need to tell it which column to use as the reference magnitude
+ref_idx = train_set.columns.get_loc("i")
+# and which columns correspond to the magnitudes we want colors for
+mag_idx = [train_set.columns.get_loc(band) for band in "ugrizy"]
+
+# the next bijector is shift bounds
+# we need to set the mins and maxes
+# I am setting strict limits on redshift, but am adding some padding to
+# the magnitudes and colors so that the flow can sample a little
+colors = -np.diff(train_set[list("ugrizy")].to_numpy())
+mins = np.concatenate(([0, train_set["i"].min() - 0.1], colors.min(axis=0)))
+maxs = np.concatenate(([3, train_set["i"].max() + 0.1], colors.max(axis=0)))
+
+# I will add buffers to the mins and maxs in case that the train set
+# doesn't cover the full range of the test set
+ranges = maxs - mins
+buffer = ranges
+buffer[0] = 0  # except no buffer for redshift!
+mins -= buffer
+maxs += buffer
+
+# finally, the settings for the RQ-RSC
+nlayers = train_set.shape[1]  # layers = number of dimensions
+K = 16  # number of spline knots
+transformed_dim = 1  # only transform one dimension at a time
+
+# chain all the bijectors together
+bijector = Chain(
[…]

### 2022-07-21 src/tex/ms.tex
Added beginning of photo-z work.
Added photo-z point estimate metrics.

@@
     We demonstrate via modeling photometry and redshifts for a galaxy catalog, and estimate photo-z posteriors.
+    Emphasize that while this paper applies the technique to photo-z's, this technique is powerful for any galaxy properties or tabular data in general.
     \end{abstract}
@@
             The data are colored by quadrant to visualize their image in the latent space.
-            You can sample the data distribution by sampling from the uniform distribution, and using $f^{-1}$ to map the back to the data space.
+            You can sample the data distribution by sampling from the uniform distribution, and using $f^{-1}$ to map the samples back to the data space.
         }
@@
 
-To simulate photometric errors, we use the LSST error model of \citet{ivezic2019}, except without the high-SNR assumptions made there (see Appendix B for details).
+To simulate photometric errors, we use the LSST error model of \citet{ivezic2019}, except without the high-SNR assumptions made there (see Appendix \ref{app:error-model} for details).
 
@@
 \note{is the LR the same?}
-We sample a size and ellipticity for each galaxy in the PZFlow catalog created in the previous section (using the true magnitudes), and plot the distribution of these features against the distribution in the test set (Figure \note{N}).
+We sample a size and ellipticity for each galaxy in the PZFlow catalog created in the previous section (using the true magnitudes), and plot the distribution of these features against the distribution in the test set (Figure \ref{fig:conditional-corner}).
 Once again, we see that the normalizing flow does a good job of emulating the CosmoDC2 distribution.
@@
 The final simulated catalog consists of $10^4$ galaxies, each with a redshift, $ugrizy$ magnitudes including photometric errors, a true photo-z posterior consistent with the photometric errors, a size, and an ellipticity.
-This small catalog was generated for visualization purposes, but the normalizing flows can be used to generate catalogs of arbitrarly large size (see Appendix C on benchmarking PZFlow).
+This small catalog was generated for visualization purposes, but the normalizing flows can be used to generate catalogs of arbitrarly large size (see Appendix \ref{app:benchmarking} on benchmarking PZFlow).
 
@@
 
+In addition to forward modeling, normalizing flows are powerful and flexible models for density estimation.
+This makes them useful tools for estimating posterior distributions for galaxy properties, conditioned on observed properties of the galaxy.
+A common example of this in cosmology is photometric redshift estimation, in which you try to estimate the redshift of a galaxy using its magnitude in several photometric bands.
+In this section, we will demonstrate this application to show how normalizing flows can be used for density estimation.
+
 \subsection{Training an Ensemble for photo-z estimation}
 
-\note{
-    Galaxy catalog created in the previous section that has true posteriors. \\
-    Ensembles. \\
-    Training with error sampling.
-}
+When forward modeling in Section \ref{sec:galaxy-catalog}, we wanted a realistic model that captured the relevant correlations between galaxy photometry, redshift, shape, and size.
+However, when estimating redshifts, we do not simply want a realistic model, but rather a model that matches our galaxy sample as closely as possible.
+
+When training deep learning models, the huge parameter space contains many different solutions, corresponding to different local minima in the parameter space.
+In the forward modeling application, we were content with finding a good local minimum, but in this application, we want to marginalize over the different potential models.
+
+A full marginalization over the model parameters would be too computationally expensive, so instead we opt for an ensemble of normalizing flows.
+In other words, we will train multiple normalizing flows under identical conditions, just with different random initializations of the model parameters.
+This will allow the optimization algorithm to explore different basins of attraction in the parameter space, allowing us to marginalize over different potential models.
+In the machine learning literature, this is known as a Deep Ensemble \needscite, a popular method for approximate bayesian marginalization \needscite.
+
+We train an ensemble of 4 normalizing flows, each with the same architecture and training schedule as the flow described in Section \ref{sec:galaxy-catalog}.
+PZFlow makes this easy to do automatically, by simply swapping \texttt{FlowEnsemble} for \texttt{Flow}.
+
+For the training set, we use 100,000 galaxies from the catalog created in Section \ref{sec:galaxy-catalog}.
+Each galaxy in the training set has a true redshift, and observed magnitudes in the $ugrizy$ bands, with corresponding photometric errors.
+To account for the photometric error, we resample the training set from the error distributions on each epoch.
+In other words, each epoch, we resample
+\begin{align}
+    m \sim p(\hat{m}, \sigma_m),
+\end{align}
+where $\hat{m}$ are the observed magnitudes with photometric errors $\sigma_m$, and $p(\hat{m}, \sigma_m)$ is a Gaussian in flux space.
+This allows our ensemble of flows to approximate the distribution $p(z, m)$.
+
+The training curves for the flows in the ensemble can be seen in Appendix \ref{app:training-details}.
 
@@
 
-\note{
-    Estimating raw posteriors. \\
-    Estimating posteriors with error convolution. \\
-    Estimating posteriors with u band marginalization.
+\begin{figure*}[t!]
+    \script{photo-z/plot_ensemble_posteriors.py}
+    \begin{centering}
+        \includegraphics{figures/ensemble_posteriors.pdf}
+        \caption{
+            \note{Write this caption}.
         }
+        \label{fig:ensemble-posteriors}
+    \end{centering}
+\end{figure*}
+
+Each flow estimates the posterior by marginalizing over the error distribution.
+We then average the posteriors from each flow in the ensemble.
+An example of this can be seen in Figure \ref{fig:ensemble-posteriors}.
+Each flow produces a PDF which may contain slightly different features in each case.
+By averaging over the individual posteriors, we select for features that are common between models, while smoothing over features that are present in only a single model.
+Compare these to the true posteriors.
+
+We can calculate posteriors for galaxies with missing magnitudes by marginalizing over the missing magnitudes.
+This might be useful when not every galaxy is observed in every band, in which case you can marginalize over the bands that were not observed, or in cases when a galaxy was observed but not detected in the band.
+You can then marginalize over values for the band that are beyond the limiting magnitude.
+This can also help you understand what information is included in each band and how the inclusion of that band improves photo-z estimation.
+
+For example, we randomly select \note{N} galaxies and estimate redshift posteriors while marginalizing over the $u$ band.
+These posteriors can be seen in Figure \note{N}.
+Mention how the loss of information degrades the posteriors.
+
+At some point in this section, note that we estimate posteriors for different galaxies than we trained on.
 
@@
 
-\note{
-    Photo-z point plot. \\
-    Binned point metrics vs redshift. \\
-    Brier score and other PDF metrics.
+In this section, we calculate common photo-z metrics to evaluate how well the ensemble of flows captures the information present in the catalog.
+We calculate each metric for the full catalog and the sub catalog with missing $u$ bands.
+Note that these metrics are optimistic in the sense that the training set is representative of the test set, which is usually not the case when applied to real galaxies.
+
+\begin{figure}[t!]
+    \script{photo-z/plot_pz_point_estimates.py}
+    \begin{centering}
+        \includegraphics{figures/pz_point_estimates.pdf}
+        \caption{
+            \note{Write this caption}.
         }
+        \label{fig:point-estimates}
+    \end{centering}
+\end{figure}
+
+The most common point of comparison is photo-z point estimates.
+We make the common choice of selecting the mode of the posteriors.
+\note{something something.}
+
+\begin{figure}[t!]
+    \script{photo-z/plot_pit_histogram.py}
+    \begin{centering}
+        \includegraphics{figures/pit_histogram.pdf}
+        \caption{
+            \note{Write this caption}.
+        }
+        \label{fig:pit-histogram}
+    \end{centering}
+\end{figure}
+
+PIT metric.
+\note{something something.}
+Discuss the mode covering behavior of normalizing flows and how this relates to the PIT diagram.
+
+\begin{figure*}[t!]
+    \script{photo-z/plot_binned_metrics.py}
+    \begin{centering}
+        \includegraphics{figures/binned_metrics.pdf}
+        \caption{
+            \note{Write this caption}.
+        }
+        \label{fig:binned-metrics}
+    \end{centering}
+\end{figure*}
+
+Binned point metrics.
+Figure \ref{fig:binned-metrics}
+Defined in the science book.
+We plot against the requirements from the LSST SRD just to provide a reference for the scale.
+Seems to do well up to z~1.5.
+We note that for cosmology, it is okay for the bias to exceed the required limits, as long as the bias can be well determined via some calibration process, e.g. clustering redshifts.
+
+The previous metrics concerned photo-z point estimates, which are insufficient for modern science.
+The DC1 paper demonstrated that these metrics are not sufficient to validate the full posteriors produced by photo-z codes.
+The purpose of this paper is to produce catalogs with true posteriors, which we have done, so that you can evaluate metrics of the full posterior.
+We will not undertake a full analysis of the posteriors here, but will compute a few metrics to point in that direction.
+
+Calculate the KLD and maybe something else?
 
@@
 \section{Training details}
+\label{app:training-details}
 
@@
 \section{LSST Error Model}
+\label{app:error-model}
 
@@
 
-\section{Failed experiments}
-
-- CentBeta - harder to train, slower, results aren't better
-- convolving Gaussian errors analytically (use the trained flow from above to plot the gradients of the spline or something and show that the necessary assumptions are not valid)
-
 \section{Benchmarking PZFlow}
+\label{app:benchmarking}
 

### 2022-07-21 src/scripts/photo-z/plot_binned_metrics.py

@@ -1,0 +1,83 @@
+"""Plot the binned metrics for photo-z point estimates."""
+# %%
+import matplotlib.pyplot as plt
+import numpy as np
+from load_pzflow_catalog import load_pzflow_catalog
+from showyourwork.paths import user as Paths
+
+# instantiate the paths
+paths = Paths()
+
+# load the estimated posteriors
+data = np.load(paths.data / "redshift_posteriors.npz")
+
+grid = data["grid"]
+pdfs = data["pdfs"]
+
+# get point estimates
+z_phot = grid[pdfs.argmax(axis=1)]
+
+# load the true redshifts
+test_set = load_pzflow_catalog(subset="test")
+z_spec = test_set.redshift[: len(z_phot)]
+
+# calculate the weighted error values
+ez = (z_phot - z_spec) / (1 + z_spec)
+
+# %%
+# sort the galaxies by their spectroscopic bins
+dz = 0.25
+bins = np.arange(0, 3.0 + dz, dz)
+bin_idx = np.digitize(z_spec, bins)
+
+# iterate over bins and calculate the metrics
+bin_z = []
+bias = []
+sigma_iqr = []
+mad = []
+outlier_frac = []
+for i in range(len(bins)):
+    # calculate the mean redshift of the bin
+    bin_zs = z_spec[bin_idx == i]
+
+    # if this bin is empty, move on to next bin
+    if len(bin_zs) == 0:
+        continue
+
+    # get the weighted errors for the bin
+    bin_ez = ez[bin_idx == i]
+    bin_z.append(bin_zs.mean())
+
+    # calculate the bias
+    bias.append(np.median(bin_ez))
+
+    # calculate the sigma_iqr
+    q75, q25 = np.percentile(bin_ez, [75.0, 25.0])
+    iqr = q75 - q25
+    sigma_iqr.append(iqr / 1.349)
+
+    # calculate the outlier fraction
+    cut = max(0.06, 3 * sigma_iqr[-1])
+    outlier_frac.append(np.mean(bin_ez > cut))
+
+# %%
+fig, (ax1, ax2, ax3) = plt.subplots(
+    1, 3, figsize=(7, 2), constrained_layout=True, dpi=200
+)
+
+ax1.set(xlabel="redshift", ylabel="bias", xlim=(bins.min(), bins.max()))
[…]

### 2022-07-21 src/scripts/photo-z/plot_ensemble_losses.py

@@ -1,0 +1,22 @@
+"""Plot the training losses for the pz ensemble."""
+# %%
+import pickle
+from pathlib import Path
+
+import numpy as np
+from showyourwork.paths import user as Paths
+
+"""
+This currently doesn't work because I accidentally saved the losses as jax floats.
+I need to re-run training and save the losses as regular floats.
+"""
+
+# instantiate the paths
+paths = Paths()
+
+# %%
+# load the losses
+with open(paths.data / "pz_ensemble/losses.pkl", "rb") as file:
+    losses = pickle.load(file)
+
+# %%

### 2022-07-21 src/scripts/photo-z/plot_pit_histogram.py

@@ -1,0 +1,37 @@
+"""Plot the PIT histogram."""
+import matplotlib.pyplot as plt
+import numpy as np
+from load_pzflow_catalog import load_pzflow_catalog
+from showyourwork.paths import user as Paths
+
+# instantiate the paths
+paths = Paths()
+
+# load the estimated posteriors
+data = np.load(paths.data / "redshift_posteriors.npz")
+
+grid = data["grid"]
+pdfs = data["pdfs"]
+
+# calculate cdfs
+cdfs = pdfs.cumsum(axis=1) * (grid[1] - grid[0])
+
+# get the true redshifts
+test_set = load_pzflow_catalog(subset="test")
+z_spec = test_set.redshift.to_numpy()[: len(cdfs)]
+
+# get PIT values
+idx = np.abs(grid - z_spec[:, None]).argmin(axis=1)
+pits = cdfs[np.arange(len(cdfs)), idx]
+
+# plot the PIT histogram
+fig, ax = plt.subplots(figsize=(3, 2.5), constrained_layout=True)
+ax.hist(pits, bins=20, density=True, histtype="step")
+ax.plot([0, 1], [1, 1], c="k", ls="--", alpha=0.75)
+ax.set(
+    xlim=(0, 1),
+    xlabel="PIT Value",
+    ylabel="Density",
+)
+
+fig.savefig(paths.figures / "pit_histogram.pdf")

### 2022-07-21 src/scripts/photo-z/plot_pz_point_estimates.py

@@ -1,0 +1,36 @@
+"""Plot the photo-z point estimates."""
+import matplotlib.pyplot as plt
+import numpy as np
+from load_pzflow_catalog import load_pzflow_catalog
+from showyourwork.paths import user as Paths
+
+# instantiate the paths
+paths = Paths()
+
+# load the estimated posteriors
+data = np.load(paths.data / "redshift_posteriors.npz")
+
+grid = data["grid"]
+pdfs = data["pdfs"]
+
+# get point estimates
+z_phot = grid[pdfs.argmax(axis=1)]
+
+# load the true redshifts
+test_set = load_pzflow_catalog(subset="test")
+z_spec = test_set.redshift[: len(z_phot)]
+
+# create the figure
+fig, ax = plt.subplots(figsize=(3, 3), constrained_layout=True)
+ax.scatter(z_spec, z_phot, s=1, rasterized=True)
+ax.plot(grid, grid, c="k", lw=1)
+ax.set(
+    xlabel="redshift",
+    ylabel="photo-z",
+    xticks=np.arange(0, 4),
+    yticks=np.arange(0, 4),
+    xlim=(grid.min(), grid.max()),
+    ylim=(grid.min(), grid.max()),
+)
+
+fig.savefig(paths.figures / "pz_point_estimates.pdf", dpi=600)

### 2022-07-21 src/scripts/photo-z/plot_binned_metrics.py
Removed unused variable.

@@ -34,7 +34,6 @@ bin_idx = np.digitize(z_spec, bins)
 bin_z = []
 bias = []
 sigma_iqr = []
-mad = []
 outlier_frac = []
 for i in range(len(bins)):
     # calculate the mean redshift of the bin

### 2022-08-17 src/tex/ms.tex
Updated pz ensemble and point estimate plot.

@@
 In that case, more care should be taken in designing a bijection that maps the data onto the range of the splines.}.
-You can choose a range wider than that covered by the training set if you wish to allow the flow to sample outside the range of the training set.
-
-\note{Discuss the Inverse Softplus Bijection here! If I use it below. Not sure if I will...}
+You can choose a range wider than that covered by the training set if you wish to allow the flow to sample outside the range of the training set
 
@@
 
-\note{
-Discuss latent distributions here.
-Note that most applications use a Normal Distribution, but since the splines have a compact domain, we need to choose a latent distribution with compact support, or else we might generate strange outliers when sampling from the flow.
-Some text for this might already be in the creation module draft.
-}
+In principle, with a sufficiently expressive bijection, the choice of latent distribution does not matter as long as it is a distribution in which you can easily sample and calculate densities.
+However, in practice, bijections are limited in expressiveness, i.e. they cannot necessarily transform any arbitrary data distribution into any arbitrary latent distribution.
+
+For example, the splines of RQ-RSCs only transform samples in the range [-B, B].
+Sampling from a latent distribution with support outside this range will therefore result in strange outliers and incorrect boundary conditions.
+One can apply a transformation to the latent samples before they are fed into the RQ-RSC to ensure that they lie within the support of the splines, but it is simpler to use a compact latent distribution whose support matches that of the spline layers.
+A simple choice would be the uniform distribution over [-B, B].
+
+Additionally, as no bijector is perfect, the structure of the latent distribution will not be completely erased in the translation from latent to data distribution.
+Thus, the latent distribution can be viewed as a prior or inductive bias on samples from the data distribution \citep{jaini2020}.
+It is therefore advantageous to select a latent distribution whose features match some of the structure in the data.
+
+A latent distribution that can achieve both desiderata is the Beta distribution, i.e.
+$u \sim \mathrm{Beta}(\alpha, \beta)$, where $\alpha, \beta > 0$ are learnable parameters\footnote{In practice, it is easier to learn $\log\alpha$ and $\log\beta$ to ensure that $\alpha, \beta > 0$.}.
+This distribution is compact, and by varying $\alpha$ and $\beta$ this distribution can take on a wide variety of shapes with different means, skews, and kurtoses, allowing the inductive bias of the prior to adapt to structure in the data during training.
+However, as RQ-RSCs are defined on the domain $[-B, B]$, it is more convenient to use a modified Beta distribution, which we name the \emph{Centered Beta distribution}:
+\begin{align}
+    \text{CentBeta}(u | \alpha, \beta, B) = 2B\left(\text{Beta}(u|\alpha, \beta) - \frac{1}{2}\right).
+\end{align}
+
+In general, as long as sampling and density evaluation are tractable, one can use any parameterization of the latent distribution that matches some desired structure in the data and learn the distribution parameters during training.
+We give this generalization the name \emph{latent-adaptive flows} (LAFs; inspired by the Tail Adaptive Flows of \citealt{jaini2020}).
+Our experiments indicate that learnable latent distributions can improve training loss, but require more care in training.
+
+Note that while we discussed univariate distributions above, these considerations generalize easily to multiple dimensions.
+Each of these distributions have multivariate generalizations that can be used when modeling higher-dimensional data.
+The full multivariate latent distribution can also be assembled by taking the product of multiple univariate distributions\footnote{Note that while the latent variables will be independent, the data variables will still have correlations imprinted by the bijections.}.
+This may even be desired if different dimensions of the data have different structure that you wish to encode in the latent distribution.
 
@@
 
-See Figure \ref{fig:sphere} for an example of circular splines used to model the population density of the Earth \note{cite the data set here}.
-The angles are shifted so that $\phi=\pm \pi$ corresponds to the Prime Meridian.
-You can see that the circular splines have no problem with smoothly modeling the data around this discontinuity in $\phi$.
-
 The circular spline construction above is easily generalized to n-spheres and n-tori: $S^n \to [-1, 1]^{n-1} \times S^1$ and $T^n \to (S^1)^n$ (see \citealt{rezende2020} for more details).
@@
 
-\note{Try to get the discrete example working again! Add a figure here.}
-
 \section{PZFlow}
@@
 The code is easily installable for the Python Package Index\footnote{\url{https://pypi.org/project/pzflow/}} (PyPI) and is hosted on Github,\footnote{\url{https://github.com/jfcrenshaw/pzflow}}.
-The documentation\footnote{\note{URL}} includes tutorial notebooks covering all of the features demonstrated in this paper.
+The documentation\footnote{\url{https://jfcrenshaw.github.io/pzflow/}} includes tutorial notebooks demonstrating the features mentioned in this paper on different example problems.
 
@@
 \begin{itemize}
-    \item \needscite used PZFlow to build a metric for observing strategy optimization based on information theory;
-    \item \needscite used PZFlow to forward model galaxy data with true redshift posteriors in order to evaluate the impact of survey incompleteness and spec-z errors on photo-z estimation;
-    \item \needscite used PZFlow to smooth high-redshift artifacts in simulations of supernova host galaxy simulations.
+    \item \citet{malz2021} used PZFlow to build a metric for observing strategy optimization based on information theory;
+    \item \citet{stylianou2022} used PZFlow to forward model galaxy data with true redshift posteriors in order to evaluate the impact of survey incompleteness and spec-z errors on photo-z estimation;
+    \item \citet{lokken2022} used PZFlow to smooth high-redshift artifacts in simulations of host galaxies for supernovae and other transients.
 \end{itemize}
@@
 
-In Section \ref{sec:fwd-model} we construct a normalizing flow to model the galaxy redshifts and photometry.
-In Section \note{LABEL}, we calculate posteriors?
+In Section \ref{sec:fwd-model} we construct a normalizing flow to model the galaxy redshifts and photometry and generate a new simulated catalog.
+In Section \ref{sec:true-posteriors}, we calculate true redshift posteriors for the new catalog.
 In Section \ref{sec:fwd-model-conditional} we build a conditional flow to add additional properties to the catalog.
@@
 
-To create a generative model of galaxy redshifts and photometry, we use the true redshifts and $ugrizy$ magnitudes from the DESC CosmoDC2 simulation (\needscite).
+To create a generative model of galaxy redshifts and photometry, we use the true redshifts and $ugrizy$ magnitudes from the DESC CosmoDC2 simulation \citep{dc2, cosmodc2}.
 We selected all galaxies from CosmoDC2 with at most one band beyond the LSST 10-year 5-sigma point source depths.
@@
 \end{align}
-\note{I might add an InverseSoftPlus}
 We will explain each layer of the bijection in the order they are applied to the input data.
@@
 
-We trained the normalizing flow for 150 epochs using the Adam optimizer \citep{adam} (with a learning rate of $10^{-3}$, that decreased by a factor of 10 every 50 epochs) to minimize the negative log-likelihood,
-\begin{align}
-    \mathcal{L} = - \, \mathbb{E}[ \, \log p(x) \, ],
-\end{align}
-where the expectation is performed over galaxies in the training set and $p(x)$ is defined in Equation \ref{eq:px}.
-Training took 7 minutes on a Tesla P100 12GB GPU.
-
-We assess the results by drawing $10^4$ galaxies from the trained flow, and plotting their distribution against  $10^4$ galaxies from the test set (Figure \ref{fig:main-corner}).
+After training the flow (see Appendix \ref{app:training-details}), we assess the results by drawing $10^4$ galaxies from the trained flow, and plotting their distribution against  $10^4$ galaxies from the test set (Figure \ref{fig:main-corner}).
 We see that the normalizing flow has done an excellent job of reproducing the distribution of DESC DC2, without any unusual artifacts or outliers.
@@
 You may wish to marginalize over all values of $n$ if the galaxy was not observed in that band.
-This might occur, for example, if you include photometry from Euclid, which will not have complete coverage of the LSST catalog \needscite.
+This might occur, for example, if you include photometry from Euclid, which will not have complete coverage of the LSST catalog \citep{euclid}.
 You may also wish to marginalize over all values beyond a limiting magnitude if the galaxy was observed, but not detected in that band.
@@
 \subsection{Additional properties with conditional flows}
-\label{sec:fwd-model-condtional}
+\label{sec:fwd-model-conditional}
 
@@
 
-The conditional flow is trained for \note{N} epochs using the same loss function and optimizer as before.
-\note{is the LR the same?}
-We sample a size and ellipticity for each galaxy in the PZFlow catalog created in the previous section (using the true magnitudes), and plot the distribution of these features against the distribution in the test set (Figure \ref{fig:conditional-corner}).
+After training the flow (see Appendix \ref{app:training-details}), we sample a size and ellipticity for each galaxy in the PZFlow catalog created in the previous section (using the true magnitudes), and plot the distribution of these features against the distribution in the test set (Figure \ref{fig:conditional-corner}).
 Once again, we see that the normalizing flow does a good job of emulating the CosmoDC2 distribution.
@@
 The final simulated catalog consists of $10^4$ galaxies, each with a redshift, $ugrizy$ magnitudes including photometric errors, a true photo-z posterior consistent with the photometric errors, a size, and an ellipticity.
-This small catalog was generated for visualization purposes, but the normalizing flows can be used to generate catalogs of arbitrarly large size (see Appendix \ref{app:benchmarking} on benchmarking PZFlow).
+This small catalog was generated for visualization purposes, but the normalizing flows can be used to generate catalogs of arbitrarily large size.
 
@@
 In addition to forward modeling, normalizing flows are powerful and flexible models for density estimation.
-This makes them useful tools for estimating posterior distributions for galaxy properties, conditioned on observed properties of the galaxy.
+This makes them useful tools for estimating posterior distributions for galaxy properties, conditioned on observed features of the galaxy.
 A common example of this in cosmology is photometric redshift estimation, in which you try to estimate the redshift of a galaxy using its magnitude in several photometric bands.
@@
 This will allow the optimization algorithm to explore different basins of attraction in the parameter space, allowing us to marginalize over different potential models.
-In the machine learning literature, this is known as a Deep Ensemble \needscite, a popular method for approximate bayesian marginalization \needscite.
+In the machine learning literature, this is known as a Deep Ensemble \citep{lakshminarayanan2017}, a popular method for approximate bayesian marginalization \citep{wilson2020,fort2020}.
 
@@
 For the training set, we use 100,000 galaxies from the catalog created in Section \ref{sec:galaxy-catalog}.
-Each galaxy in the training set has a true redshift, and observed magnitudes in the $ugrizy$ bands, with corresponding photometric errors.
-To account for the photometric error, we resample the training set from the error distributions on each epoch.
+Each galaxy in the training set has a true redshift and observed magnitudes in the $ugrizy$ bands, with corresponding photometric errors.
+To account for the photometric error, at the start of each training epoch, we resample the training set from the photometric error distribution.
 In other words, each epoch, we resample
@@
         \caption{
-            \note{Write this caption}.
+            The ensemble of posteriors for three example galaxies.
+            Flows 1-4 label the individual posteriors produced by each of the flows that make up the ensemble.
+            The dashed black line is the mean of these individual posteriors and is the value used by the ensemble.
+            The vertical gray line labeled ``Truth'' denotes the true redshift of the galaxy.
+            Note the posteriors for most galaxies are not this broad or multimodal. 
         }
@@
 
-Each flow estimates the posterior by marginalizing over the error distribution.
+Each flow in the ensemble estimates the posterior for each galaxy by marginalizing over the photometric errors:
+\begin{align}
+    p(z| \hat{m}, \sigma_m) \propto \int dm \, p(z, m) \, p(m| \hat{m}, \sigma_m),
+\end{align}
+which is estimated by sampling $m \sim p(\hat{m}, \sigma_m)$ and averaging $p(z, m)$ over these samples.
 We then average the posteriors from each flow in the ensemble.
-An example of this can be seen in Figure \ref{fig:ensemble-posteriors}.
+
+An example of this for three galaxies can be seen in Figure \ref{fig:ensemble-posteriors}.
 Each flow produces a PDF which may contain slightly different features in each case.
@@
 We can calculate posteriors for galaxies with missing magnitudes by marginalizing over the missing magnitudes.
-This might be useful when not every galaxy is observed in every band, in which case you can marginalize over the bands that were not observed, or in cases when a galaxy was observed but not detected in the band.
-You can then marginalize over values for the band that are beyond the limiting magnitude.
-This can also help you understand what information is included in each band and how the inclusion of that band improves photo-z estimation.
+This is done by averaging $p(z, m)$ over a grid of values for the missing bands. 
+This might be useful when not every galaxy is observed in every band, in which case you can marginalize over the bands that were not observed.
+Or if a galaxy was observed but not detected in the band, you can marginalize over values for the band that are beyond the limiting magnitude.
+You can also marginalize over bands for which you have a magnitude to understand what information about redshift is included in that band.
 
+\note{Think about what I want to do here.}
 For example, we randomly select \note{N} galaxies and estimate redshift posteriors while marginalizing over the $u$ band.
@@
 Mention how the loss of information degrades the posteriors.
-
-At some point in this section, note that we estimate posteriors for different galaxies than we trained on.
 
@@
         \caption{
-            \note{Write this caption}.
+            Comparing the photo-z point estimates to the true redshifts in the catalog.
+            The black diagonal line represents perfect photo-z performance.
         }
@@
 
-The most common point of comparison is photo-z point estimates.
-We make the common choice of selecting the mode of the posteriors.
-\note{something something.}
+The most common metrics for photo-z estimation concern photo-z point estimates, which are a compression of the photo-z posterior to a single redshift estimate.
+We make the common choice of selecting the mode of the posteriors\footnote{The mean redshift is often a poor choice, since photo-z posteriors are generally multimodal, and so the mean value can lie between two modes at a redshift with very small probability density.}.
+We compute metrics of the quantity $\Delta z = (z_\text{phot} - z_\text{true}) / (1 + z_\text{true})$, where the denominator accounts for naturally greater uncertainties at high redshift.
+
+\begin{figure*}[t!]
+    \script{photo-z/plot_binned_metrics.py}
+    \begin{centering}
+        \includegraphics{figures/binned_metrics.pdf}
+        \caption{
+            The bias, scatter, and catastrophic outlier fraction of the photo-z point estimates as a function of true galaxy redshift.
+            The dashed black lines represent the science requirements for LSST cosmology as stated in the LSST SRD \needscite.
+            These lines are to provide a sense of scale for these metrics.
+            You can see that PZFlow meets the bias and scatter requirements up to redshift $\sim$ 1.5, while meeting the outlier fraction requirements for all redshifts.
+            We note that individual redshifts do not actually need to meet the bias requirement as long as the bias can be well calibrated via some other source, e.g. galaxy clustering.
+        }
+        \label{fig:binned-metrics}
+    \end{centering}
+\end{figure*}
+
+Figure \ref{fig:binned-metrics} shows the metrics from \citet{descSRD} as a function of true redshift.
+The bias is defined as the median of $\Delta z$; the scatter is defined as $\text{IQR} / 1.349$, where IQR is the interquartile range of $\Delta z$; the outlier fraction is defined as the fraction of galaxies for which $\Delta z$ is greater than three times the $\Delta z$ scatter.
+As a point of comparison, we plot the requirements for LSST cosmology in gray.
+
+Like many photo-z estimators, our estimator performs well to a redshift of approximately 1.5.
+At higher redshifts, our estimator does not meet the bias and scatter requirements, because there is very little training data in this redshift range.
+We note however that for many cosmology application, it is okay for the bias to exceed the required limits, as long as the bias can be well determined via some calibration process \citep{newman2022}.
 
@@
         \caption{
-            \note{Write this caption}.
+            The histogram of the probability integral transform (PIT) values of the photo-z posteriors.
+            This figure characterizes the calibration of the estimated posteriors.
+            In other words, does the true redshift lie at the center/in the tails of the posteriors as often as one would expect?
+            Ideally, this histogram should be flat, matching the dashed black line.
+            The fact that this distribution has a central peak indicates that the flow ensemble produces posteriors that are a little too conservative in their predictions.
+            The slight rightward tilt indicates that the posteriors also have a negative redshift bias.
         }
@@
 
-PIT metric.
-\note{something something.}
-Discuss the mode covering behavior of normalizing flows and how this relates to the PIT diagram.
+Another common metric is the probability integral transform (PIT) (see e.g. \citealt{schmidt2020, dey2022}).
+The PIT metric is used to determine if posteriors are well calibrated, i.e. if true values fall within X\% confidence intervals X\% of the time.
+The PIT histogram for our estimator is shown in Figure \ref{fig:pit-histogram}.
+Ideally, this histogram would be uniform and match the dashed horizontal line.
+The fact that the histogram bulges at the center indicates that our estimator is too conservative -- i.e. the posteriors it produces are too broad.
+This can be explained by the fact that normalizing flows exhibit mode covering behavior (the opposite of the mode collapse seen in GANsl; \citealt{salimans2016}).
+In other words, because normalizing flows are trained via maximum likelihood estimation on the training samples, they receive very high penalties for missing any modes in the data.
+As a result, they tend to conservatively spread out their density, in order to avoid any missed modes.
+This results in overly conservative posterior predictions.
+The low values at the edges of the PIT histogram indicate the relative rarity of catastrophic outliers, which is also reflected in the far right panel of Figure \ref{fig:binned-metrics}, where you can see that our estimator meets the requirement on the outlier fraction at all redshifts.
 
-\begin{figure*}[t!]
-    \script{photo-z/plot_binned_metrics.py}
-    \begin{centering}
-        \includegraphics{figures/binned_metrics.pdf}
-        \caption{
-            \note{Write this caption}.
-        }
-        \label{fig:binned-metrics}
-    \end{centering}
-\end{figure*}
+The previous metrics analyze photo-z performance for point estimates, which are insufficient for modern cosmology \citep{newman2022}, and for ensembles of posteriors, which is often not a good indicator of photo-z performance for science applications \citep{schmidt2020}.
+One of the purposes of this paper is to create a galaxy catalog for which each galaxy has a true redshift posterior, which will enable more comprehensive evaluation of photo-z estimators.
 
-Binned point metrics.
-Figure \ref{fig:binned-metrics}
-Defined in the science book.
-We plot against the requirements from the LSST SRD just to provide a reference for the scale.
-Seems to do well up to z~1.5.
-We note that for cosmology, it is okay for the bias to exceed the required limits, as long as the bias can be well determined via some calibration process, e.g. clustering redshifts.
+\note{NEED TO FINISH WRITING THIS}
+We will not undertake a full analysis of the posteriors here, but will compute a few metrics to point in that direction.
+Calculate the KLD and maybe something else?
+Point towards RAIL.
 
-The previous metrics concerned photo-z point estimates, which are insufficient for modern science.
-The DC1 paper demonstrated that these metrics are not sufficient to validate the full posteriors produced by photo-z codes.
-The purpose of this paper is to produce catalogs with true posteriors, which we have done, so that you can evaluate metrics of the full posterior.
-We will not undertake a full analysis of the posteriors here, but will compute a few metrics to point in that direction.
-
-Calculate the KLD and maybe something else?
+\section{Conclusion}
 
@@
             Training losses for the galaxy flows.
+            After epoch 50 and 100, you can see a drop in the loss due to the decrease in the learning rate.
         }
@@
 
+\note{the main flow:}
+We trained the normalizing flow for 150 epochs using the Adam optimizer \citep{adam} (with a learning rate of $10^{-3}$, that decreased by a factor of 10 every 50 epochs) to minimize the negative log-likelihood,
+\begin{align}
+    \mathcal{L} = - \, \mathbb{E}[ \, \log p(x) \, ],
+\end{align}
+where the expectation is performed over galaxies in the training set and $p(x)$ is defined in Equation \ref{eq:px}.
+Training took 7 minutes on a Tesla P100 12GB GPU.
+
+\note{
+    Need to add the curves for the ensembles.
+    Move discussion of epochs and learning rates here too?    
+}
+
 \section{LSST Error Model}
@@
 
-- Explain how I adapted the LSST Error Model to work for low-SNR galaxies
+We estimate photometric errors for LSST using the error model from \citet{ivezic2019}, except we do not make the high SNR approximation used there.
 
-\section{Benchmarking PZFlow}
-\label{app:benchmarking}
+\citet{ivezic2019} calculates a Gaussian photometric error in magnitude space using Equation 5 from that paper: 
+\begin{align}
+    \sigma_\text{rand}^2 = (0.04 - \gamma) x + \gamma x^2 ~~ (\text{mag}^2),
+\end{align}
+where $x \equiv 10 ^ {0.4(m - m_5)}$, $m_5$ is the 5$\sigma$ depth for the band, and $\gamma$ is a band-dependent parameter listed in Table 2 of the same paper.
+This formula is based on the approximation
+\begin{align}
+    \sigma_\text{rand} = 2.5 \log_{10} \left(1 + \frac{N}{S} \right) \approx \frac{N}{S},
+\end{align}
+where $S$ and $N$ are the flux signal and noise, respectively.
+This approximation is valid in the high signal-to-noise regime.
 
-\bibliography{references.bib}
+To generalize this model to the low signal-to-noise regime, we abandon this latter approximation.
+For each magnitude, $m$, we calculate the corresponding flux, $S = 10^{-m/2.5}$, and the Gaussian flux noise,
+\begin{align}
+    N = S \, \sqrt{(0.04 - \gamma) x + \gamma x^2}, \label{eq:noise}.
+\end{align}
+For very bright galaxies, we impose a systematic floor corresponding to an error of 0.005 mags.
+We then sample an observed flux, $S_\text{obs} \sim \mathcal{N}(S, N)$, and calculate the observed magnitude, $m_\text{obs} = -2.5 \log_{10} S_\text{obs}$.
+
+For the reported photometric error, we do not use $N$, as this provides a deterministic link to the true magnitude via Equation \ref{eq:noise}.
+Instead, we apply Equation \ref{eq:noise} again, this time using the observed magnitude instead of the true magnitude, yielding the estimated noise, $N_\text{obs}$.
+Finally, we report the photometric error in magnitudes:
+\begin{align}
+    \sigma_m = 2.5 \log_{10} \left(1 + \frac{N_\text{obs}}{S_\text{obs}} \right).
+\end{align}
+
+\bibliography{bib.bib}
 

### 2022-08-17 src/scripts/photo-z/calculate_posteriors.py

@@ -18,7 +18,7 @@ grid = np.linspace(0, 3.5, 351)
 
 # calculate posteriors
 pdfs = flowEns.posterior(
-    test_set[:1000],  # remove this for final paper!
+    test_set[:100_000],
     "redshift",
     grid,
     err_samples=10,

### 2022-08-17 src/scripts/photo-z/plot_binned_metrics.py

@@ -70,7 +70,7 @@ ax1.plot(bins, +0.003 * np.ones(len(bins)), c="k", ls="--", alpha=0.75)
 ax1.plot(bins, -0.003 * np.ones(len(bins)), c="k", ls="--", alpha=0.75)
 
 ax2.set(
-    xlabel="redshift", ylabel="$\sigma_{\mathrm{IQR}}$", xlim=(bins.min(), bins.max())
+    xlabel="redshift", ylabel="scatter", xlim=(bins.min(), bins.max())
 )
 ax2.plot(bin_z, sigma_iqr)
 ax2.plot(bins, 0.02 * np.ones(len(bins)), c="k", ls="--", alpha=0.75)

### 2022-08-17 src/scripts/photo-z/plot_pz_point_estimates.py

@@ -18,19 +18,23 @@ z_phot = grid[pdfs.argmax(axis=1)]
 
 # load the true redshifts
 test_set = load_pzflow_catalog(subset="test")
-z_spec = test_set.redshift[: len(z_phot)]
+z_spec = test_set.redshift[: len(z_phot)].to_numpy()
+
+# select the non-zero z_phots
+idx = np.where(z_phot > 0)
 
 # create the figure
-fig, ax = plt.subplots(figsize=(3, 3), constrained_layout=True)
-ax.scatter(z_spec, z_phot, s=1, rasterized=True)
-ax.plot(grid, grid, c="k", lw=1)
+fig, ax = plt.subplots(figsize=(3.3, 2.6), constrained_layout=True, dpi=200)
+hb = ax.hexbin(z_spec[idx], z_phot[idx], extent=(0, 3, 0, 3), gridsize=205, bins="log")
+cb = fig.colorbar(hb, ax=ax)
+cb.set_label("Galaxies")
 ax.set(
-    xlabel="redshift",
+    xlabel="true redshift",
     ylabel="photo-z",
     xticks=np.arange(0, 4),
     yticks=np.arange(0, 4),
-    xlim=(grid.min(), grid.max()),
-    ylim=(grid.min(), grid.max()),
+    xlim=(0, 3),
+    ylim=(0, 3),
 )
 
 fig.savefig(paths.figures / "pz_point_estimates.pdf", dpi=600)

### 2022-08-17 src/scripts/photo-z/train_pz_ensemble.py

@@ -11,14 +11,6 @@ from pzflow import FlowEnsemble
 from pzflow.bijectors import Chain, ColorTransform, RollingSplineCoupling, ShiftBounds
 from showyourwork.paths import user as Paths
 
-"""
-TO DO:
-------
-- fix the NaN issue that requires such a big buffer on the ShiftBounds range.
-- look at the losses. Make sure this training scheme is good. Also figure out a
-    good way to concatenate the losses
-"""
-
 # instantiate the paths
 paths = Paths()
 
@@ -40,7 +32,7 @@ mag_idx = [train_set.columns.get_loc(band) for band in "ugrizy"]
 # the magnitudes and colors so that the flow can sample a little
 colors = -np.diff(train_set[list("ugrizy")].to_numpy())
 mins = np.concatenate(([0, train_set["i"].min() - 0.1], colors.min(axis=0)))
-maxs = np.concatenate(([3, train_set["i"].max() + 0.1], colors.max(axis=0)))
+maxs = np.concatenate(([4, train_set["i"].max() + 0.1], colors.max(axis=0)))
 
 # I will add buffers to the mins and maxs in case that the train set
 # doesn't cover the full range of the test set

### 2022-11-30 src/tex/ms.tex
Manuscript updates.

@@
 
-% Abstract
 \begin{abstract}
-    Normalizing flows are useful for modeling high-dimensional probability distributions.
-    We give some background and introduce PZFlow, a Python package for statistical modeling of tabular data.
-    We demonstrate via modeling photometry and redshifts for a galaxy catalog, and estimate photo-z posteriors.
-    Emphasize that while this paper applies the technique to photo-z's, this technique is powerful for any galaxy properties or tabular data in general.
+    Normalizing flows are powerful tools for learning high-dimensional probability distributions from samples thereof.
+    We introduce PZFlow, a Python package for the statistical modeling of tabular data using normalizing flows.
+    We use PZFlow to model the seven dimensional space of galaxy redshift and LSST photometry.
+    Using these normalizing flows, we forward model galaxy catalogs in which each galaxy property has an associated ``true'' posterior distribution.
+    We discuss the importance of these posteriors for the comprehensive evaluation of the redshift posteriors produced by photometric redshift (photo-z) estimators.
+    We also demonstrate the use of an ensemble of normalizing flows for photo-z estimation.
+    While we focus on photo-z estimation and validation, we emphasize that these methods are applicable to any galaxy properties, or any other tabular data.
 \end{abstract}
 
-% Main body with filler text
 \section{Introduction}
@@
 Astronomical data represent realizations of complex probability distributions.
-Need to write more here.
+A common goal in research is to infer the underlying distribution from a limited set of noisy data, and to use this distribution to estimate posterior distributions over galaxy and population parameters.
+An important example is photometric redshift (photo-z) estimation, where galaxy redshift posteriors are estimated from galaxy photometry, using a model informed by a training set of spectroscopic galaxy data \citep{newman2022}.
+These redshift posteriors are then used in analyses that estimate posterior distributions for the cosmological parameters of interest \citep{descSRD}.
 
-While direct access to the probability distributions that characterize our data is usually impossible, machine learning provides powerful tools to learn these distributions from samples thereof.
-Generative Adversarial Networks (GANs; \needscite) and Variational Autoencoders (VAEs; \needscite) are two popular methods that provide generative models trained on samples.
-However, neither of these methods allow exact evaluation of probability densities.
+Knowledge of the underlying probability distributions is also valuable for simulation and forward modeling astronomical data sets.
+Simulating data sets with realistic statistical properties enables methodological development and the calibration of systematic uncertainties.
+Forward modeling also facilitates data augmentation (e.g.
+\citealt{lokken2022}) and the creation of large data challenges (e.g.
+\citealt{kessler2019, dc2, cosmodc2}), which are becoming more prevalent in the big data era of astronomy.
 
-Normalizing flows are a deep learning tool for density estimation and generative modeling that allow efficient sampling and exact likelihood calculation.
-The ability to generate samples from a probability distribution with respect to which you can calculate exact likelihoods is very powerful for generative modeling, as you can provide a notion of ``true'' posteriors over sample properties when simulating new data.
-For example, when simulating galaxy catalogs for photometric redshift (photo-z) estimation, you can calculate the ``true'' redshift posterior for each galaxy.
-This is useful for ...
+Inferring the underlying probability distribution from a high-dimensional data set is a difficult problem, and many machine learning tools have been developed for this task.
+Two popular examples from deep learning are Generative Adversarial Networks (GANs; \citealt{goodfellow2014}) and Variational Autoencoders (VAEs; \citealt{kingma2014}).
+Both methods excel at forward modeling complex data sets, but neither allow exact likelihood inference.
+
+Normalizing flows are a deep learning tool that excel at forward modeling, but also allow exact likelihood inference.
+They operate by learning an invertible transformation of the data distribution into a simpler, tractable distribution (such as a normal distribution, hence the name).
+This allows sampling and likelihood inference to be performed within the context of the simple distribution, with the normalizing flow acting as a translator between these samples and likelihoods, and their corresponding values in the context of the data distribution.
+
+Once trained, a normalizing flow can forward model data with realistic statistical correlations, and since normalizing flows provide exact likelihood inference, each property of the generated data has a ``true'' posterior.
+When calculating these posteriors, you can convolve error distributions, and marginalize over chosen properties.
+Access to ``true'' posteriors is very valuable for the development and validation of analysis pipelines that include posterior estimation, such as the photo-z estimators used in much of astrophysics and cosmology.
 
@@
 
-\note{[need to work in citations to Francois' paper on Deep Generative Models for Galaxy Image Sims, and any other examples of normalizing flows in the astronomy literature.]}
-
-Describe outline of the paper\dots
-
-Mention PZFlow and note that all of the NFs in this paper were created using it.
-
-Maybe mention ShowYourWork and how the scripts and data are all available at the links in the paper.
+In this paper, we provide the background on normalizing flows (Section \ref{sec:nf}) required to understand PZFlow (Section \ref{sec:pzflow}).
+We then use PZFlow to simulate a galaxy catalog (Section \ref{sec:galaxy-catalog}), where each object has photometry (including errors), size, ellipticity, redshift, and a true redshift posterior.
+We also demonstrate using PZFlow as a density estimator, via the example of photo-z estimation (Section \ref{sec:photo-z}).
+We conclude in Section \ref{sec:conclusion}.
 
@@
 Figure \ref{fig:two-moons} shows an example of a normalizing flow that transforms the \texttt{scikit-learn} two moons distribution into a uniform distribution.
-The data points are colored by quadrant to visualize their image under the bijection $f$.
+The data points are colored by quadrant to visualize their image under $f$.
 
@@
 \subsection{Choosing a latent distribution}
+\label{sec:latent}
 
@@
             The vertical gray line labeled ``Truth'' denotes the true redshift of the galaxy.
-            Note the posteriors for most galaxies are not this broad or multimodal. 
+            Note these galaxies were specifically chosen for their broad, multimodal posteriors.
+            The posteriors of most galaxies are unimodal with sharp peaks.
         }
@@
             Comparing the photo-z point estimates to the true redshifts in the catalog.
-            The black diagonal line represents perfect photo-z performance.
         }
@@
             The bias, scatter, and catastrophic outlier fraction of the photo-z point estimates as a function of true galaxy redshift.
-            The dashed black lines represent the science requirements for LSST cosmology as stated in the LSST SRD \needscite.
+            The dashed black lines represent the science requirements for LSST cosmology as stated in the LSST SRD \citep{descSRD}.
             These lines are to provide a sense of scale for these metrics.
@@
 The fact that the histogram bulges at the center indicates that our estimator is too conservative -- i.e. the posteriors it produces are too broad.
-This can be explained by the fact that normalizing flows exhibit mode covering behavior (the opposite of the mode collapse seen in GANsl; \citealt{salimans2016}).
-In other words, because normalizing flows are trained via maximum likelihood estimation on the training samples, they receive very high penalties for missing any modes in the data.
-As a result, they tend to conservatively spread out their density, in order to avoid any missed modes.
+This can be explained by the fact that normalizing flows exhibit mode covering behavior (the opposite of the mode collapse seen in GANs; \citealt{salimans2016}).
+In other words, because normalizing flows are trained by maximizing the likelihood of the training samples, they receive very high penalties for missing any modes in the data.
+As a result, they tend to conservatively spread out their density, in order to avoid missing any modes.
 This results in overly conservative posterior predictions.
@@
 The previous metrics analyze photo-z performance for point estimates, which are insufficient for modern cosmology \citep{newman2022}, and for ensembles of posteriors, which is often not a good indicator of photo-z performance for science applications \citep{schmidt2020}.
-One of the purposes of this paper is to create a galaxy catalog for which each galaxy has a true redshift posterior, which will enable more comprehensive evaluation of photo-z estimators.
-
-\note{NEED TO FINISH WRITING THIS}
-We will not undertake a full analysis of the posteriors here, but will compute a few metrics to point in that direction.
-Calculate the KLD and maybe something else?
-Point towards RAIL.
+The methods introduced in this paper allow one to create galaxy catalogs for which each galaxy has a true redshift posterior, which will enable more comprehensive evaluation of photo-z estimators.
+A comprehensive evaluation of the full posteriors from photo-z estimators is a goal of our future work.
 
 \section{Conclusion}
+\label{sec:conclusion}
 

### 2022-11-30 src/scripts/photo-z/plot_pz_point_estimates.py

@@ -24,7 +24,7 @@ z_spec = test_set.redshift[: len(z_phot)].to_numpy()
 idx = np.where(z_phot > 0)
 
 # create the figure
-fig, ax = plt.subplots(figsize=(3.3, 2.6), constrained_layout=True, dpi=200)
+fig, ax = plt.subplots(figsize=(3.3, 2.7), constrained_layout=True, dpi=200)
 hb = ax.hexbin(z_spec[idx], z_phot[idx], extent=(0, 3, 0, 3), gridsize=205, bins="log")
 cb = fig.colorbar(hb, ax=ax)
 cb.set_label("Galaxies")

### 2022-11-30 src/tex/ms.tex
Added plot with marginalized u

@@
 
+\begin{figure*}[t!]
+    \script{photo-z/plot_ensemble_posteriors.py}
+    \begin{centering}
+        \includegraphics{figures/ensemble_posteriors.pdf}
+        \caption{
+            The ensemble of posteriors for three example galaxies.
+            Flows 1-4 label the individual posteriors produced by each of the flows that make up the ensemble.
+            The dashed black line is the mean of these individual posteriors and is the value used by the ensemble.
+            The vertical gray line labeled ``Truth'' denotes the true redshift of the galaxy.
+            Note these galaxies were specifically chosen for their broad, multimodal posteriors.
+            The posteriors of most galaxies are unimodal with sharp peaks.
+        }
+        \label{fig:ensemble-posteriors}
+    \end{centering}
+\end{figure*}
+
 In addition to forward modeling, normalizing flows are powerful and flexible models for density estimation.
@@
 
-\begin{figure*}[t!]
-    \script{photo-z/plot_ensemble_posteriors.py}
-    \begin{centering}
-        \includegraphics{figures/ensemble_posteriors.pdf}
-        \caption{
-            The ensemble of posteriors for three example galaxies.
-            Flows 1-4 label the individual posteriors produced by each of the flows that make up the ensemble.
-            The dashed black line is the mean of these individual posteriors and is the value used by the ensemble.
-            The vertical gray line labeled ``Truth'' denotes the true redshift of the galaxy.
-            Note these galaxies were specifically chosen for their broad, multimodal posteriors.
-            The posteriors of most galaxies are unimodal with sharp peaks.
-        }
-        \label{fig:ensemble-posteriors}
-    \end{centering}
-\end{figure*}
-
 Each flow in the ensemble estimates the posterior for each galaxy by marginalizing over the photometric errors:
@@
 
-\note{Think about what I want to do here.}
-For example, we randomly select \note{N} galaxies and estimate redshift posteriors while marginalizing over the $u$ band.
-These posteriors can be seen in Figure \note{N}.
-Mention how the loss of information degrades the posteriors.
+\begin{figure}[t!]
+    \script{photo-z/plot_marginalized_posterior.py}
+    \begin{centering}
+        \includegraphics{figures/posterior_marginalized.pdf}
+        \caption{
+            Redshift posteriors for an example galaxy.
+            The vertical red line indicates the true redshift.
+            The black posterior is estimated using the information from all 6 LSST bands (and includes convolution of the photometric errors).
+            The gold posterior has had the $u$ band marginalized, which opens up a second high-redshift mode.
+        }
+        \label{fig:posterior_marginalized}
+    \end{centering}
+\end{figure}
+
+Figure \ref{fig:posterior_marginalized} compares posteriors with and without $u$ band information for an example galaxy.
+It is clear how losing the $u$ band information increases the uncertainty in the redshift inference.
+In this instance, a second potential redshift mode is created.
+Conversely, this demonstrates how the addition of $u$ band information rules out the higher redshift solution for this galaxy.
 
@@
 Note that these metrics are optimistic in the sense that the training set is representative of the test set, which is usually not the case when applied to real galaxies.
+
+The most common metrics for photo-z estimation concern photo-z point estimates, which are a compression of the photo-z posterior to a single redshift estimate.
+We make the common choice of selecting the mode of the posteriors\footnote{The mean redshift is often a poor choice, since photo-z posteriors are generally multimodal, and so the mean value can lie between two modes at a redshift with very small probability density.}.
+We compute metrics of the quantity $\Delta z = (z_\text{phot} - z_\text{true}) / (1 + z_\text{true})$, where the denominator accounts for naturally greater uncertainties at high redshift.
+
+Figure \ref{fig:binned-metrics} shows the metrics from \citet{descSRD} as a function of true redshift.
+The bias is defined as the median of $\Delta z$; the scatter is defined as $\text{IQR} / 1.349$, where IQR is the interquartile range of $\Delta z$; the outlier fraction is defined as the fraction of galaxies for which $\Delta z$ is greater than three times the $\Delta z$ scatter.
+As a point of comparison, we plot the requirements for LSST cosmology in gray.
+
+Like many photo-z estimators, our estimator performs well to a redshift of approximately 1.5.
+At higher redshifts, our estimator does not meet the bias and scatter requirements, because there is very little training data in this redshift range.
+We note however that for many cosmology application, it is okay for the bias to exceed the required limits, as long as the bias can be well determined via some calibration process \citep{newman2022}.
 
@@
 
-The most common metrics for photo-z estimation concern photo-z point estimates, which are a compression of the photo-z posterior to a single redshift estimate.
-We make the common choice of selecting the mode of the posteriors\footnote{The mean redshift is often a poor choice, since photo-z posteriors are generally multimodal, and so the mean value can lie between two modes at a redshift with very small probability density.}.
-We compute metrics of the quantity $\Delta z = (z_\text{phot} - z_\text{true}) / (1 + z_\text{true})$, where the denominator accounts for naturally greater uncertainties at high redshift.
+Another common metric is the probability integral transform (PIT) (see e.g. \citealt{schmidt2020, dey2022}).
+The PIT metric is used to determine if posteriors are well calibrated, i.e. if true values fall within X\% confidence intervals X\% of the time.
+The PIT histogram for our estimator is shown in Figure \ref{fig:pit-histogram}.
+Ideally, this histogram would be uniform and match the dashed horizontal line.
+The fact that the histogram bulges at the center indicates that our estimator is too conservative -- i.e. the posteriors it produces are too broad.
+This can be explained by the fact that normalizing flows exhibit mode covering behavior (the opposite of the mode collapse seen in GANs; \citealt{salimans2016}).
+In other words, because normalizing flows are trained by maximizing the likelihood of the training samples, they receive very high penalties for missing any modes in the data.
+As a result, they tend to conservatively spread out their density, in order to avoid missing any modes.
+This results in overly conservative posterior predictions.
+The low values at the edges of the PIT histogram indicate the relative rarity of catastrophic outliers, which is also reflected in the far right panel of Figure \ref{fig:binned-metrics}, where you can see that our estimator meets the requirement on the outlier fraction at all redshifts.
 
@@
 
-Figure \ref{fig:binned-metrics} shows the metrics from \citet{descSRD} as a function of true redshift.
-The bias is defined as the median of $\Delta z$; the scatter is defined as $\text{IQR} / 1.349$, where IQR is the interquartile range of $\Delta z$; the outlier fraction is defined as the fraction of galaxies for which $\Delta z$ is greater than three times the $\Delta z$ scatter.
-As a point of comparison, we plot the requirements for LSST cosmology in gray.
-
-Like many photo-z estimators, our estimator performs well to a redshift of approximately 1.5.
-At higher redshifts, our estimator does not meet the bias and scatter requirements, because there is very little training data in this redshift range.
-We note however that for many cosmology application, it is okay for the bias to exceed the required limits, as long as the bias can be well determined via some calibration process \citep{newman2022}.
+The previous metrics analyze photo-z performance for point estimates, which are insufficient for modern cosmology \citep{newman2022}, and for ensembles of posteriors, which is often not a good indicator of photo-z performance for science applications \citep{schmidt2020}.
+The methods introduced in this paper allow one to create galaxy catalogs for which each galaxy has a true redshift posterior, which will enable more comprehensive evaluation of photo-z estimators.
+A comprehensive evaluation of the full posteriors from photo-z estimators is a goal of our future work.
 
@@
 \end{figure}
-
-Another common metric is the probability integral transform (PIT) (see e.g. \citealt{schmidt2020, dey2022}).
-The PIT metric is used to determine if posteriors are well calibrated, i.e. if true values fall within X\% confidence intervals X\% of the time.
-The PIT histogram for our estimator is shown in Figure \ref{fig:pit-histogram}.
-Ideally, this histogram would be uniform and match the dashed horizontal line.
-The fact that the histogram bulges at the center indicates that our estimator is too conservative -- i.e. the posteriors it produces are too broad.
-This can be explained by the fact that normalizing flows exhibit mode covering behavior (the opposite of the mode collapse seen in GANs; \citealt{salimans2016}).
-In other words, because normalizing flows are trained by maximizing the likelihood of the training samples, they receive very high penalties for missing any modes in the data.
-As a result, they tend to conservatively spread out their density, in order to avoid missing any modes.
-This results in overly conservative posterior predictions.
-The low values at the edges of the PIT histogram indicate the relative rarity of catastrophic outliers, which is also reflected in the far right panel of Figure \ref{fig:binned-metrics}, where you can see that our estimator meets the requirement on the outlier fraction at all redshifts.
-
-The previous metrics analyze photo-z performance for point estimates, which are insufficient for modern cosmology \citep{newman2022}, and for ensembles of posteriors, which is often not a good indicator of photo-z performance for science applications \citep{schmidt2020}.
-The methods introduced in this paper allow one to create galaxy catalogs for which each galaxy has a true redshift posterior, which will enable more comprehensive evaluation of photo-z estimators.
-A comprehensive evaluation of the full posteriors from photo-z estimators is a goal of our future work.
 

### 2022-11-30 src/scripts/forward_model/plot_posteriors.py

@@ -1,3 +1,4 @@
+"""Plot the true, convolved, etc posteriors for a galaxy."""
 import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
@@ -98,8 +99,8 @@ ax1.plot(grid, pdfs_no_u[0], label="$u$ marginalized")
 ax1.set(xlabel="redshift", xlim=(0.1, 0.6), yticks=[])
 
 # put the legend in axis 2
-h, l = ax1.get_legend_handles_labels()
-ax2.legend(h, l, ncol=2, fontsize=9, handlelength=1.5, loc=10, frameon=False)
+handles, labels = ax1.get_legend_handles_labels()
+ax2.legend(handles, labels, ncol=2, fontsize=9, handlelength=1.5, loc=10, frameon=False)
 ax2.axis(False)  # hide the axes
 
 fig.savefig(paths.figures / "posteriors.pdf")

### 2022-11-30 src/scripts/photo-z/calculate_posteriors_without_u.py

@@ -1,0 +1,37 @@
+"""Caculate posteriors but marginalize over the u band."""
+import numpy as np
+from load_pzflow_catalog import load_pzflow_catalog
+from pzflow import FlowEnsemble
+from showyourwork.paths import user as Paths
+
+# instantiate the paths
+paths = Paths()
+
+# load the flow ensemble
+flowEns = FlowEnsemble(file=paths.data / "pz_ensemble" / "pz_ensemble.pzflow.pkl")
+
+# load the test set
+test_set = load_pzflow_catalog(subset="test")
+test_set = test_set[:20]
+
+# delete the u band magnitudes
+test_set["u"] *= np.nan
+
+# setup the redshift grid
+grid = np.linspace(0, 3.5, 351)
+
+# calculate posteriors
+pdfs = flowEns.posterior(
+    test_set,
+    "redshift",
+    grid,
+    err_samples=10,
+    batch_size=100,
+    marg_rules={
+        "flag": np.nan, 
+        "u": lambda row: np.linspace(row["g"]-2, row["g"]+1, 10),
+    }
+)
+
+# save the posteriors
+np.savez(paths.data / "redshift_posteriors_without_u.npz", grid=grid, pdfs=pdfs)

### 2022-11-30 src/scripts/photo-z/plot_marginalized_posterior.py

@@ -1,0 +1,47 @@
+"""Plot the posterior with u band marginalized."""
+import matplotlib.pyplot as plt
+import numpy as np
+from load_pzflow_catalog import load_pzflow_catalog
+from showyourwork.paths import user as Paths
+
+# instantiate the paths
+paths = Paths()
+
+# load the test set
+test_set = load_pzflow_catalog(subset="test")
+test_set = test_set[:20]
+
+# load the posteriors
+grid = np.load(paths.data / "redshift_posteriors.npz")["grid"]
+pdfs0 = np.load(paths.data / "redshift_posteriors.npz")["pdfs"][:20]
+pdfs1 = np.load(paths.data / "redshift_posteriors_without_u.npz")["pdfs"]
+
+# create the figure
+fig, (ax1, ax2) = plt.subplots(
+    2,
+    1,
+    figsize=(2.5, 3.4),
+    dpi=120,
+    constrained_layout=True,
+    gridspec_kw={
+        "height_ratios": [4, 1],
+    },
+)
+
+# plot the true redshift
+z = test_set["redshift"].iloc[11]
+ax1.axvline(z, c="C3", ls="--", label="True redshift")
+
+# plot the posteriors
+ax1.plot(grid, pdfs0[11], c="k", label="Posterior with all bands")
+ax1.plot(grid, pdfs1[11], c="C1", label="Posterior with $u$ marginalized")
+
+# set the x label and limits, and remove the yticks
+ax1.set(xlabel="redshift", xlim=(0, 3), yticks=[])
+
+# put the legend in axis 2
+handles, labels = ax1.get_legend_handles_labels()
+ax2.legend(handles, labels, ncol=1, fontsize=9, handlelength=1.5, loc=10, frameon=False)
+ax2.axis(False)  # hide the axes
+
+fig.savefig(paths.figures / "posterior_marginalized.pdf")

### 2022-11-30 src/tex/ms.tex
Updated appendix on training.

@@
 We train an ensemble of 4 normalizing flows, each with the same architecture and training schedule as the flow described in Section \ref{sec:galaxy-catalog}.
-PZFlow makes this easy to do automatically, by simply swapping \texttt{FlowEnsemble} for \texttt{Flow}.
+PZFlow makes this easy to do automatically, by simply swapping \texttt{FlowEnsemble} for \texttt{Flow} in the code.
 
@@
             Training losses for the galaxy flows.
-            After epoch 50 and 100, you can see a drop in the loss due to the decrease in the learning rate.
+            Left: losses for the normal flow.
+            After epochs 50 and 100, you can see a drop in the loss due to the decrease in the learning rate.
+            Right: losses for the conditional flow.
+            After epochs 150 and 300, you can see a drop in the loss due to the decrease in the learning rate.
         }
@@
 
-\note{the main flow:}
-We trained the normalizing flow for 150 epochs using the Adam optimizer \citep{adam} (with a learning rate of $10^{-3}$, that decreased by a factor of 10 every 50 epochs) to minimize the negative log-likelihood,
+\begin{figure}[t!]
+    \script{photo-z/plot_ensemble_losses.py}
+    \begin{centering}
+        \includegraphics{figures/ensemble_losses.pdf}
+        \caption{
+            Training losses for the four flows in the flow ensemble.
+            You can see that each of the flow achieves a different minimum loss.
+            After epochs 50 and 100, you can see a drop in the loss due to the decrease in the learning rate.
+        }
+        \label{fig:ensemble-losses}
+    \end{centering}
+\end{figure}
+
+In this section we list some technical details of training the normalizing flows.
+Every flow is trained via minimizing the negative log-likelihood
 \begin{align}
@@
 where the expectation is performed over galaxies in the training set and $p(x)$ is defined in Equation \ref{eq:px}.
+
+For the main flow in Section \ref{sec:galaxy-catalog}, we trained for 150 epochs.
+We used the Adam optimizer \citep{adam}, starting with a learning rate of $10^{-3}$.
+We decreased the learning rate by a factor of 10 every 50 epochs.
 Training took 7 minutes on a Tesla P100 12GB GPU.
+The training loss for this flow is in the left panel of Figure \ref{fig:galaxy-flow-losses}.
 
-\note{
-    Need to add the curves for the ensembles.
-    Move discussion of epochs and learning rates here too?    
-}
+For the conditional flow in \ref{sec:galaxy-catalog}, we trained for 450 epochs.
+Again, we used Adam with an initial learning rate of $10^{-3}$.
+We decreased the learning rate by a factor of 10 every 150 epochs.
+The training loss for this flow is in the right panel of Figure \ref{fig:galaxy-flow-losses}.
+
+For each of the flows that make up the flow ensemble in \ref{sec:photo-z}, we trained for 150 epochs using the Adam optimizer.
+We started each with a learning rate of $10^{-4}$, and decreased the learning rate by a factor of 10 every 50 epochs.
+The training loss for the ensemble is in Figure \ref{fig:ensemble-losses}.
+You can see that each of the flows achieves a different minimum loss.
+Apparently, each has found a different potential solution in the neural network's parameter space.
 

### 2022-11-30 src/scripts/photo-z/plot_ensemble_losses.py

@@ -1,22 +1,22 @@
 """Plot the training losses for the pz ensemble."""
-# %%
 import pickle
-from pathlib import Path
 
-import numpy as np
+import matplotlib.pyplot as plt
 from showyourwork.paths import user as Paths
-
-"""
-This currently doesn't work because I accidentally saved the losses as jax floats.
-I need to re-run training and save the losses as regular floats.
-"""
 
 # instantiate the paths
 paths = Paths()
 
-# %%
 # load the losses
 with open(paths.data / "pz_ensemble/losses.pkl", "rb") as file:
     losses = pickle.load(file)
 
-# %%
+# plot the losses
+fig, ax = plt.subplots(figsize=(3, 3), constrained_layout=True)
+for flow, loss in losses.items():
+    ax.plot(loss, alpha=0.5, label=flow)
+ax.legend()
+ax.set(xlabel="Epochs", ylabel="Loss", ylim=(-1.6, 0))
+
+# save the figure
+fig.savefig(paths.figures / "ensemble_losses.pdf")

### 2022-12-01 src/tex/ms.tex
PhotErr.

@@
 
+\author{Ziang Yan}
+
+\author{Andrew Connolly}
+
+\author{Francois Lanusse}
+
 \author{Alex Malz}
@@
 
-\author{Francois Lanusse}
-
 \collaboration{99}{(The LSST Dark Energy Science Collaboration)}
@@
 \begin{abstract}
-    Normalizing flows are powerful tools for learning high-dimensional probability distributions from samples thereof.
+    Normalizing flows are powerful tools for learning high-dimensional probability distributions from samples thereof, and have many applications in astronomy, including posterior estimation and forward modeling.
     We introduce PZFlow, a Python package for the statistical modeling of tabular data using normalizing flows.
-    We use PZFlow to model the seven dimensional space of galaxy redshift and LSST photometry.
-    Using these normalizing flows, we forward model galaxy catalogs in which each galaxy property has an associated ``true'' posterior distribution.
-    We discuss the importance of these posteriors for the comprehensive evaluation of the redshift posteriors produced by photometric redshift (photo-z) estimators.
+    We use PZFlow to model photometric galaxy catalogs including redshifts, LSST photometry, size, and ellipticity.
+    We generate synthetic galaxy catalogs in which each galaxy has a ``true'' redshift posterior distribution, and discuss the importance of these true posteriors for the comprehensive evaluation of the redshift posteriors produced by photometric redshift (photo-z) estimators.
     We also demonstrate the use of an ensemble of normalizing flows for photo-z estimation.
-    While we focus on photo-z estimation and validation, we emphasize that these methods are applicable to any galaxy properties, or any other tabular data.
+    While we focus on photo-z estimation and validation, we emphasize that these methods are applicable to any galaxy properties, and any other tabular data.
 \end{abstract}
@@
         \caption{
-            A normalizing flow demonstrated on the two moons data set from \texttt{scikit-learn}.
+            A normalizing flow demonstrated on the two moons data set from scikit-learn.
             The two moons data on the left is mapped onto a two dimensional uniform distribution by the bijection $f$.
@@
 
-Figure \ref{fig:two-moons} shows an example of a normalizing flow that transforms the \texttt{scikit-learn} two moons distribution into a uniform distribution.
+Figure \ref{fig:two-moons} shows an example of a normalizing flow that transforms the scikit-learn \citep{sklearn} two moons distribution into a uniform distribution.
 The data points are colored by quadrant to visualize their image under $f$.
@@
 The documentation\footnote{\url{https://jfcrenshaw.github.io/pzflow/}} includes tutorial notebooks demonstrating the features mentioned in this paper on different example problems.
-
-PZFlow is a package for building normalizing flows in Python, with a focus on high-performance out-of-the-box modeling of high dimensional tabular data.
-Data is handled in Pandas DataFrames \citep{pandas} while the normalizing flows are implemented in Jax \citep{jax}, which allows efficient, parallelizable, GPU-enabled calculations for very large data sets.
-The code is easily installable from the Python Package Index\footnote{\url{https://pypi.org/project/pzflow/}} (PyPI) and is hosted on Github,\footnote{\url{https://github.com/jfcrenshaw/pzflow}} including several tutorial notebooks.
 
@@
 
-\section{Conclusion}
+\section{Conclusion \& Summary}
 \label{sec:conclusion}
+
+In this paper we introduced PZFlow, a normalizing flow package in Python, designed for the statistical modeling of astronomical data sets.
+We used PZFlow to forward model galaxy catalogs that include photometry, redshifts, sizes, and ellipticities.
+In addition, each galaxy in our catalog has a true redshift posterior, which can be convolved with measurement errors.
+These true posteriors allow a direct evaluation of the posteriors produced by photo-z estimators.
+A similar comparison can be made to posterior estimates for any other galaxy property.
+
+Direct evaluation of photo-z posteriors is vital for future cosmology analyses which must use all of the information incorporated in the full redshift posteriors \citep{newman2022}.
+This is because only the full redshift posteriors allow one to account for degeneracies in color-redshift space, which will otherwise bias cosmological inference.
+Previous evaluations of photo-z performance focused on point estimates and metrics for ensembles of posteriors, but \citet{schmidt2020} showed these analyses are inadequate for modern cosmology.
+This work enables a future analysis of photo-z estimators on a per-posterior basis, which is a major goal of the LSST DESC.
+
+In addition to forward modeling galaxy catalogs, we demonstrated PZFlow's utility as a density estimator that can be applied to photo-z estimation and other statistical analyses.
+PZFlow achieves high accuracy with relatively little fine tuning and with very few modeling assumptions.
+PZFlow will be a powerful tool for the statistical analysis of tabular astronomical data.
+
+Note that the code to reproduce the figures in this paper is hosted publicly on Github\footnote{\url{https://github.com/jfcrenshaw/pzflow-paper}}, and the code for each individual figure can be reached by clicking on the Github logo in the margin next to the figure.
+
+\begin{acknowledgements}
+    JFC is supported by...
+\end{acknowledgements}
+
+\software{
+    adam \citep{adam},
+    corner \citep{corner},
+    dill \citep{dill},
+    jax \citep{jax},
+    jupyter \citep{jupyter},
+    matplotlib \citep{matplotlib},
+    numpy \citep{numpy},
+    pandas \citep{pandas,pandas-software},
+    scipy \citep{scipy},
+    showyourwork \citep{showyourwork},
+    scikit-learn \citep{sklearn}
+}
 
@@
             Training losses for the four flows in the flow ensemble.
-            You can see that each of the flow achieves a different minimum loss.
-            After epochs 50 and 100, you can see a drop in the loss due to the decrease in the learning rate.
+            We have zoomed in to the bottom of the loss curve so you can see that each of the flows converges to a different minimum loss.
         }
@@
 
-We estimate photometric errors for LSST using the error model from \citet{ivezic2019}, except we do not make the high SNR approximation used there.
+We estimate photometric errors for LSST using a generalization of the error model from \citet{ivezic2019}.
+To derive the error model, we start with the noise-to-signal ratio (NSR) for an object with photon count $C$ and background noise $N_0$ (which depends on seeing, read-out noise, etc.):
+\begin{align}
+    \text{NSR}^2 = \frac{N_0^2 + C}{C^2}.
+\end{align}
+If we define $C=C_5$ when $\text{NSR}= 1/5$, then we can solve for $N_0$ and write
+\begin{align}
+    \text{NSR}^2 = \frac{1}{C_5} \left( \frac{C_5}{C} \right) + \left[ \left( \frac{1}{5} \right)^2 - \frac{1}{C_5} \right] \left( \frac{C_5}{C} \right)^2.
+\end{align}
+Defining $x = C_5/C = 10^{(m-m_5)/2.5}$ and $\gamma = 1/5^2 - 1/C_5$, we have
+\begin{align}
+    \text{NSR}^2 = (0.04 - \gamma) \, x + \gamma \, x^2 ~~ (\text{mag}^2),
+\end{align}
+which is Equation 5 from \citet{ivezic2019}.
+Values for the band-dependent parameter $\gamma$ can be found in Table 2 of the same paper.
 
-\citet{ivezic2019} calculates a Gaussian photometric error in magnitude space using Equation 5 from that paper: 
+In the high signal-to-noise (SNR) limit, $\text{NSR} \ll 1$, and we can approximate
 \begin{align}
-    \sigma_\text{rand}^2 = (0.04 - \gamma) x + \gamma x^2 ~~ (\text{mag}^2),
+    \sigma_\text{rand} = 2.5 \log_{10}(1 + \text{NSR}) \approx \text{NSR}.
+    \label{eq:err}
 \end{align}
-where $x \equiv 10 ^ {0.4(m - m_5)}$, $m_5$ is the 5$\sigma$ depth for the band, and $\gamma$ is a band-dependent parameter listed in Table 2 of the same paper.
-This formula is based on the approximation
-\begin{align}
-    \sigma_\text{rand} = 2.5 \log_{10} \left(1 + \frac{N}{S} \right) \approx \frac{N}{S},
-\end{align}
-where $S$ and $N$ are the flux signal and noise, respectively.
-This approximation is valid in the high signal-to-noise regime.
+This latter approximation is made by \citet{ivezic2019}, and errors are assumed to be Gaussian in magnitude space.
+In contrast, we use the exact form of Equation \ref{eq:err}, and model errors as Gaussian in flux space.
+Note that after the photometric errors are applied, the error is re-calculated from the ``observed'' flux, and this new error is reported as the estimated photometric error.
+If the original photometric error were reported, it would provide a deterministic link to the original flux.
 
-To generalize this model to the low signal-to-noise regime, we abandon this latter approximation.
-For each magnitude, $m$, we calculate the corresponding flux, $S = 10^{-m/2.5}$, and the Gaussian flux noise,
-\begin{align}
-    N = S \, \sqrt{(0.04 - \gamma) x + \gamma x^2}, \label{eq:noise}.
-\end{align}
-For very bright galaxies, we impose a systematic floor corresponding to an error of 0.005 mags.
-We then sample an observed flux, $S_\text{obs} \sim \mathcal{N}(S, N)$, and calculate the observed magnitude, $m_\text{obs} = -2.5 \log_{10} S_\text{obs}$.
-
-For the reported photometric error, we do not use $N$, as this provides a deterministic link to the true magnitude via Equation \ref{eq:noise}.
-Instead, we apply Equation \ref{eq:noise} again, this time using the observed magnitude instead of the true magnitude, yielding the estimated noise, $N_\text{obs}$.
-Finally, we report the photometric error in magnitudes:
-\begin{align}
-    \sigma_m = 2.5 \log_{10} \left(1 + \frac{N_\text{obs}}{S_\text{obs}} \right).
-\end{align}
+We have implemented this error model, along with several other extensions to it, in the Python package PhotErr, which is available on the Python Package Index\footnote{\url{https://pypi.org/project/photerr/}} (PyPI), and Github\footnote{\url{https://github.com/jfcrenshaw/photerr}}.
+The extensions include different methods for handling non-detections, methods for modeling errors of extended objects \citep{vandenbusch2020,kuijken2019}, and error models for the Roman and Euclid space telescopes \citep{roman,euclid,graham2020}, however, these extensions are not used in this paper.
 

### 2022-12-01 src/scripts/forward_model/simulate_pzflow_catalog.py

@@ -1,6 +1,6 @@
 """Simulate a PZFlow-generated catalog of 1e6 galaxies."""
 import pandas as pd
-from lsst_error_model import LSSTErrorModel
+from photerr import LsstErrorModel
 from pzflow import Flow
 from showyourwork.paths import user as Paths
 
@@ -18,8 +18,8 @@ ellip_and_size = cond_flow.sample(
 )
 
 # simulate observations with the fiducial LSST error model
-err_model = LSSTErrorModel()
-obs_sample = err_model(truth_sample[list("ugrizy")], seed=2)
+err_model = LsstErrorModel(sigLim=1)
+obs_sample = err_model(truth_sample[list("ugrizy")], random_state=2)
 
 # relabel the true and observed values
 truth_sample = truth_sample.rename(

### 2022-12-01 src/scripts/photo-z/load_pzflow_catalog.py

@@ -1,4 +1,5 @@
 """Load the PZFlow catalog created by the forward model."""
+import numpy as np
 import pandas as pd
 from showyourwork.paths import user as Paths
 
@@ -7,7 +8,7 @@ def load_pzflow_catalog(
     truth: bool = False,
     subset: str = None,
     train_frac: float = 0.8,
-    include_nans: bool = False,
+    finite: bool = True,
 ) -> pd.DataFrame:
     """Load the catalog created by the PZFlow forward model.
 
@@ -22,8 +23,8 @@ def load_pzflow_catalog(
     train_frac: float, default=0.8
         The fraction of the catalog to include in the train set.
         1 - train_frac is the fraction included in the test set.
-    include_nans: bool, default=False
-        Whether to include galaxies that were not observed in every band.
+    finite: bool, default=True
+        Whether to only include galaxies whose properties are all finite.
     """
     # instantiate the paths
     paths = Paths()
@@ -54,8 +55,8 @@ def load_pzflow_catalog(
     elif subset == "test":
         data = data.iloc[split_idx:]
 
-    if not include_nans:
+    if finite:
         # get the rows without NaNs
-        data = data[~data.isnull().any(axis=1)]
+        data = data[np.isfinite(data).all(axis=1)]
 
     return data

### 2022-12-01 src/scripts/photo-z/plot_ensemble_losses.py

@@ -16,7 +16,7 @@ fig, ax = plt.subplots(figsize=(3, 3), constrained_layout=True)
 for flow, loss in losses.items():
     ax.plot(loss, alpha=0.5, label=flow)
 ax.legend()
-ax.set(xlabel="Epochs", ylabel="Loss", ylim=(-1.6, 0))
+ax.set(xlabel="Epochs", ylabel="Loss", ylim=(-1.8, -1))
 
 # save the figure
 fig.savefig(paths.figures / "ensemble_losses.pdf")

### 2022-12-01 src/scripts/photo-z/plot_ensemble_posteriors.py

@@ -15,7 +15,7 @@ flowEns = FlowEnsemble(file=paths.data / "pz_ensemble" / "pz_ensemble.pzflow.pkl
 test_set = load_pzflow_catalog(subset="test")
 
 # select galaxies for plotting
-idx = [21, 54, 99]
+idx = [39, 52, 44]
 test_set = test_set.iloc[idx]
 
 # calculate posteriors
@@ -46,6 +46,6 @@ for i, ax in enumerate(axes):
     ax.set(yticks=[], xlabel="redshift", xlim=(grid.min(), grid.max()))
 
 # put a legend in the center
-axes[1].legend(loc="upper left")
+axes[1].legend(loc="upper right")
 
 fig.savefig(paths.figures / "ensemble_posteriors.pdf")

### 2022-12-01 src/scripts/photo-z/plot_marginalized_posterior.py

@@ -29,12 +29,13 @@ fig, (ax1, ax2) = plt.subplots(
 )
 
 # plot the true redshift
-z = test_set["redshift"].iloc[11]
+idx = 10
+z = test_set["redshift"].iloc[idx]
 ax1.axvline(z, c="C3", ls="--", label="True redshift")
 
 # plot the posteriors
-ax1.plot(grid, pdfs0[11], c="k", label="Posterior with all bands")
-ax1.plot(grid, pdfs1[11], c="C1", label="Posterior with $u$ marginalized")
+ax1.plot(grid, pdfs0[idx], c="k", label="Posterior with all bands")
+ax1.plot(grid, pdfs1[idx], c="C1", label="Posterior with $u$ marginalized")
 
 # set the x label and limits, and remove the yticks
 ax1.set(xlabel="redshift", xlim=(0, 3), yticks=[])

### 2022-12-01 src/tex/ms.tex
Fixed dependency typo.

@@
     We introduce PZFlow, a Python package for the statistical modeling of tabular data using normalizing flows.
-    We use PZFlow to model photometric galaxy catalogs including redshifts, LSST photometry, size, and ellipticity.
-    We generate synthetic galaxy catalogs in which each galaxy has a ``true'' redshift posterior distribution, and discuss the importance of these true posteriors for the comprehensive evaluation of the redshift posteriors produced by photometric redshift (photo-z) estimators.
-    We also demonstrate the use of an ensemble of normalizing flows for photo-z estimation.
+    We use PZFlow to model photometric galaxy catalogs including redshifts, LSST photometry, size, and ellipticity, and generate a synthetic catalog.
+    In this catalog, each galaxy has a \emph{true} redshift posterior.
+    We discuss the importance of these true posteriors for the comprehensive evaluation of redshift posteriors produced by photometric redshift (photo-z) estimators.
+    We also demonstrate the use of an ensemble of normalizing flows for density estimation, applied to photo-z estimation.
     While we focus on photo-z estimation and validation, we emphasize that these methods are applicable to any galaxy properties, and any other tabular data.
@@
 Astronomical data represent realizations of complex probability distributions.
-A common goal in research is to infer the underlying distribution from a limited set of noisy data, and to use this distribution to estimate posterior distributions over galaxy and population parameters.
+A common goal in research is to infer the underlying distribution from a limited set of noisy data, and use this distribution to estimate posterior distributions over galaxy and population parameters.
 An important example is photometric redshift (photo-z) estimation, where galaxy redshift posteriors are estimated from galaxy photometry, using a model informed by a training set of spectroscopic galaxy data \citep{newman2022}.
-These redshift posteriors are then used in analyses that estimate posterior distributions for the cosmological parameters of interest \citep{descSRD}.
+These redshift posteriors are then used to estimate posterior distributions for cosmological parameters \citep{descSRD}.
 
-Knowledge of the underlying probability distributions is also valuable for simulation and forward modeling astronomical data sets.
+Knowledge of the probability distributions underlying our data is also valuable for simulation and forward modeling astronomical data sets.
 Simulating data sets with realistic statistical properties enables methodological development and the calibration of systematic uncertainties.
@@
 Inferring the underlying probability distribution from a high-dimensional data set is a difficult problem, and many machine learning tools have been developed for this task.
-Two popular examples from deep learning are Generative Adversarial Networks (GANs; \citealt{goodfellow2014}) and Variational Autoencoders (VAEs; \citealt{kingma2014}).
+Two popular deep learning examples are Generative Adversarial Networks (GANs; \citealt{goodfellow2014}) and Variational Autoencoders (VAEs; \citealt{kingma2014}).
 Both methods excel at forward modeling complex data sets, but neither allow exact likelihood inference.
 
-Normalizing flows are a deep learning tool that excel at forward modeling, but also allow exact likelihood inference.
-They operate by learning an invertible transformation of the data distribution into a simpler, tractable distribution (such as a normal distribution, hence the name).
-This allows sampling and likelihood inference to be performed within the context of the simple distribution, with the normalizing flow acting as a translator between these samples and likelihoods, and their corresponding values in the context of the data distribution.
+Normalizing flows, on the other hand, are a deep learning tool that excel at forward modeling, while also allowing exact likelihood inference.
+They operate by learning an invertible transformation of the data distribution into a simpler, tractable latent distribution (such as a normal distribution, hence the name \emph{normalizing} flow).
+This allows sampling and likelihood inference to be performed within the context of the latent distribution, with the normalizing flow acting as a translator between the latent samples and likelihoods, and their corresponding values in the context of the data distribution.
 
-Once trained, a normalizing flow can forward model data with realistic statistical correlations, and since normalizing flows provide exact likelihood inference, each property of the generated data has a ``true'' posterior.
+Once trained, a normalizing flow can forward model data with realistic statistical correlations, and since normalizing flows provide exact likelihood inference, each property of the generated data has a \emph{true} posterior.
 When calculating these posteriors, you can convolve error distributions, and marginalize over chosen properties.
-Access to ``true'' posteriors is very valuable for the development and validation of analysis pipelines that include posterior estimation, such as the photo-z estimators used in much of astrophysics and cosmology.
+Catalogs with \emph{true} posteriors are useful for the testing and validation of analysis pipelines that estimate posteriors, such as the photo-z estimators used in much of astrophysics and cosmology.
 
-To facilitate the statistical modeling of galaxy catalogs and other astronomical data sets, we have developed PZFlow, a normalizing flow package for python.
+To facilitate the statistical modeling of galaxy catalogs and other astronomical data sets, we have developed PZFlow, a normalizing flow package for Python.
 With relatively little tuning required by the user, PZFlow can provide a generative model for any tabular data, including continuous and discrete variables, and variables with Euclidean or periodic topology.
@@
 In this paper, we provide the background on normalizing flows (Section \ref{sec:nf}) required to understand PZFlow (Section \ref{sec:pzflow}).
-We then use PZFlow to simulate a galaxy catalog (Section \ref{sec:galaxy-catalog}), where each object has photometry (including errors), size, ellipticity, redshift, and a true redshift posterior.
+We then use PZFlow to simulate a galaxy catalog (Section \ref{sec:galaxy-catalog}), where each object has photometry, size, ellipticity, redshift, and a true redshift posterior.
 We also demonstrate using PZFlow as a density estimator, via the example of photo-z estimation (Section \ref{sec:photo-z}).

### 2022-12-02 src/tex/ms.tex
Changes to text. Diagram for conditional flow. Jax from conda.

@@
 
-\shorttitle{Modeling Galaxy Catalogues with Normalizing Flows}
+% aliases for the DC2 PZ paper
+\defcitealias{schmidt2020}{Schmidt \& Malz et al.}
+\newcommand{\citePZp}{(\citetalias{schmidt2020} \citeyear{schmidt2020})\xspace}
+\newcommand{\citePZt}{\citetalias{schmidt2020} (\citeyear{schmidt2020})\xspace}
+\newcommand{\citePZa}{\citetalias{schmidt2020} \citeyear{schmidt2020}\xspace}
+
+\shorttitle{Galaxy Catalogues with Normalizing Flows}
 \shortauthors{Crenshaw et al.}
@@
 Normalizing flows model complex, high-dimensional probability distributions by learning a mapping from the data distribution to a tractable latent distribution\footnote{Some of the machine learning literature defines the mapping in the opposite direction.}.
-Often the latent distribution is a standard Normal distribution, and so the mapping ``normalizes'' the data, hence the name ``noramlizing flow''.
+Often the latent distribution is a standard Normal distribution, and so the mapping ``normalizes'' the data, hence the name ``normalizing flow''.
 This mapping allows us to sample and evaluate densities using the latent distribution, rather than the unknown data distribution.
@@
 
-If we further assume that is invertible, we can sample from the data distribution by appling $f^{-1}$ to samples from the latent distribution:
+If we further assume that $f$ is invertible, we can sample from the data distribution by appling $f^{-1}$ to samples from the latent distribution:
 \begin{align}
@@
 Section \ref{sec:latent} discusses how to choose an appropriate latent distribution for your data;
-Section \ref{sec:conditional} describe how to condition a normalizing flow on some conditional variables;
+Section \ref{sec:conditional} describes how to build a flow that models a conditional distribution;
 Section \ref{sec:periodic} explains how to model data with periodic topology;
@@
 
-The bijection of a normalizing flow must be powerful enough to model complex relationships in data, while simultaneously possessing an efficiently computable Jacobian determinant.
+The bijection of a normalizing flow must be powerful enough to model complex relationships in data, while remaining invertible and simultaneously possessing an efficiently computable Jacobian determinant.
 This latter constraint is the primary difficulty in designing a normalizing flow.
@@
     \begin{tikzpicture}[thick, outer sep=0]
-        \node [draw, rectangle, minimum height=2cm, minimum width=1.2cm, fill=blue!20] (1) {$x^{}_{1:d}$};
-        \node [draw, rectangle, minimum height=2cm, minimum width=1.2cm, fill=red!20, below = 0 of 1] (2) {$x^{}_{d+1:D}$};
-        \node [draw, circle, minimum size=8mm, fill=gray!20, right = 2cm of 1] (3) {$=$};
-        \node [draw, circle, minimum size=8mm, fill=gray!20, right = 2cm of 2] (4) {$g$};
-        \node [draw, rectangle, minimum height=2cm, minimum width=1.2cm, fill=blue!20, right = 1.5cm of 3] (5) {$y^{}_{1:d}$};
-        \node [draw, rectangle, minimum height=2cm, minimum width=1.2cm, fill=red!20, below = 0 of 5] (6) {$y^{}_{d+1:D}$};
-        \draw[->] (1) -- (3);
-        \draw[->] (2) -- (4);
-        \draw[->] (3) -- (5);
-        \draw[->] (4) -- (6);
-        \draw[->] (1) -- (4) node [midway, draw, circle, fill=white, minimum size=8mm, fill=gray!20] {$m$};
+        % blue boxes
+        \node [draw, rectangle, minimum height=2cm, minimum width=1.28cm, fill=blue!20] (1) {$x^{}_{1:d}$};
+        \node [draw, circle, minimum size=8mm, fill=gray!20, right = 2cm of 1] (2) {$=$};
+        \node [draw, rectangle, minimum height=2cm, minimum width=1.28cm, fill=blue!20, right = 1.5cm of 2] (3) {$y^{}_{1:d}$};
+        \draw[->] (1) -- (2);
+        \draw[->] (2) -- (3);
+
+        % red boxes
+        \node [draw, rectangle, minimum height=2cm, minimum width=1.28cm, fill=red!20, below = 0 of 1] (4) {$x^{}_{d+1:D}$};
+        \node [draw, circle, minimum size=8mm, fill=gray!20, right = 2cm of 4] (5) {$g$};
+        \node [draw, rectangle, minimum height=2cm, minimum width=1.28cm, fill=red!20, below = 0 of 3] (6) {$y^{}_{d+1:D}$};
+        \draw[->] (4) -- (5);
+        \draw[->] (5) -- (6);
+
+        % coupling function
+        \draw[->] (1) -- (5) node [midway, draw, circle, fill=white, minimum size=8mm, fill=gray!20] (7) {$m$};
+
     \end{tikzpicture}
@@
 
-The obvious limitation of a coupling later is that only a subset of the data variables are transformed.
+The obvious limitation of a coupling layer is that only a subset of the data dimensions are transformed.
 This is overcome by stacking multiple coupling layers in succession, and switching which variables belong to which partition.
@@
 $K \to \infty$), RQ-RSCs can model any differentiable, monotonic function on $[-B, B]^D$ and can thus model arbitrarily complex distributions in this region.
-In practice, we find very good performance for diverse data sets with much smaller $K$.
+In practice, we find very good performance for diverse data sets with $K=16$.
 
-Note that you can also specify a different value of $K$ for each of the $D$ spline layers in order to individually control the resolution of each dimension.
+Note you can specify a different value of $K$ for each of the $D$ spline layers in order to individually control the resolution of each dimension.
 
@@
 It is therefore useful to define a \emph{Shift Bounds} bijection, which shifts the original range of each dimension to match the domain of the splines.
-Note that this shift must be set at training time, with the assumption that future test data will lie within the same bounds\footnote{This bijection may be poorly suited to data with large outliers.
-In that case, more care should be taken in designing a bijection that maps the data onto the range of the splines.}.
+Note this shift must be set at training time, with the assumption that future test data will lie within the same bounds\footnote{While this sounds quite restrictive, neural networks are typically pretty bad at extrapolating beyond the bounds of the training set anyway.}.
 You can choose a range wider than that covered by the training set if you wish to allow the flow to sample outside the range of the training set
@@
 For an example of building an application-specific data processing bijection, see the \emph{Color Transform} bijection defined in Section \ref{sec:fwd-model}, which maps galaxy magnitudes to galaxy colors.
-See section \ref{sec:discrete} for data processing bijections, which enable the flow to model discrete data.
+See section \ref{sec:discrete} for data processing bijections that enable modeling of discrete data.
 
@@
 
-Additionally, as no bijector is perfect, the structure of the latent distribution will not be completely erased in the translation from latent to data distribution.
+Additionally, as no bijection is perfect, the structure of the latent distribution will not be completely erased in the translation from latent to data distribution.
 Thus, the latent distribution can be viewed as a prior or inductive bias on samples from the data distribution \citep{jaini2020}.
@@
 
-The bijectors and latent distributions discussed above can be easily adapted to directly learn conditional probability distributions:
-you only need to make the replacement $f(x) \to f(x;y)$, where $y$ is a vector of conditions \citep{winkler2019}.
-In practice, as the bijection $f$ is parameterized by a neural network, this can be simply achieved by appending the vector $y$ to the neural network's input.
+\begin{figure}
+    \centering
+    \begin{tikzpicture}[thick, outer sep=0]
+        % blue boxes
+        \node [draw, rectangle, minimum height=2cm, minimum width=1.28cm, fill=blue!20] (1) {$x^{}_{1:d}$};
+        \node [draw, circle, minimum size=8mm, fill=gray!20, right = 2cm of 1] (2) {$=$};
+        \node [draw, rectangle, minimum height=2cm, minimum width=1.28cm, fill=blue!20, right = 1.5cm of 2] (3) {$y^{}_{1:d}$};
+        \draw[->] (1) -- (2);
+        \draw[->] (2) -- (3);
 
-While $p(x|y)$ is technically encoded within $p(x,y)$, which can be learned with a regular normalizing flow, directly modeling $p(x|y)$ with a conditional flow has a few benefits.
+        % red boxes
+        \node [draw, rectangle, minimum height=2cm, minimum width=1.28cm, fill=red!20, below = 0 of 1] (4) {$x^{}_{d+1:D}$};
+        \node [draw, circle, minimum size=8mm, fill=gray!20, right = 2cm of 4] (5) {$g$};
+        \node [draw, rectangle, minimum height=2cm, minimum width=1.28cm, fill=red!20, below = 0 of 3] (6) {$y^{}_{d+1:D}$};
+        \draw[->] (4) -- (5);
+        \draw[->] (5) -- (6);
+
+        % coupling function
+        \draw[->] (1) -- (5) node [midway, draw, circle, fill=white, minimum size=8mm, fill=gray!20] (7) {$m$};
+
+        % conditional variables
+        \node [draw, rectangle, minimum height=1.28cm, minimum width=1.28cm, fill=green!20, above = 0.25 of 1] (8) {$z^{}_{1:L}$};
+        \draw[-] (8.south east) -- (7.north west);
+
+    \end{tikzpicture}
+    \caption{
+        Diagram of a \emph{conditional} coupling layer.
+        The first partition, $x^{}_{1:d}$, is passed through the layer unchanged.
+        The second partition, $x^{}_{d+1:D}$, is transformed by the coupling law $g$, which is parameterized by the coupling function $m$ applied to the first partition \emph{and} the conditional variables $z^{}_{1:L}$.
+        The conditional variables are \emph{never} altered by the flow.
+    }
+    \label{fig:conditional-coupling}
+\end{figure}
+
+The bijections and latent distributions discussed above can be easily adapted to directly learn conditional probability distributions:
+you only need to make the replacement $f(x) \to f(x;z)$, where $z$ is a vector of conditions \citep{winkler2019}.
+This is illustrated in Figure \ref{fig:conditional-coupling}, which is a modification of Figure \ref{fig:coupling} to include the input of conditional variables to the coupling function $m$.
+In practice, since $m$ is usually a neural network, this amounts to just appending the conditions $z$ to the inputs of the neural network.
+
+While $p(x|z)$ is technically encoded within $p(x,z)$, which can be learned with a regular normalizing flow, directly modeling $p(x|z)$ with a conditional flow has a few benefits.
 Training is typically faster, since the latent distribution has a smaller number of dimensions.
-You can also draw samples of $x$ at fixed values of the conditions $y$, and you can calculate $p(x|y)$ without having to numerically calculate and divide by $p(y)$, which can be computationally expensive.
+You can also draw samples of $x$ at fixed values of the conditions $z$, and you can calculate $p(x|z)$ without having to numerically calculate and divide by $p(z)$, which can be computationally expensive.
 
@@
 In addition to the continuous variables described above, normalizing flows can also be used to model discrete variables.
-This can be achieved by ``dequantizing'' the discrete dimensions of the data, which can then be mapped onto continuous latent distributions using regular continuous bijectors.
+This can be achieved by ``dequantizing'' the discrete dimensions of the data, which can then be mapped onto continuous latent distributions using regular continuous bijections.
 Dequantization consists of adding some kind of continuous noise to the discrete dimensions, transforming them into continuous dimensions.
@@
 
-While the dequantizers are not technically bijections, they can be treated as data processing bijections and be chained together with the other bijectors in your normalizing flow.
+While the dequantizers are not technically bijections, they can be treated as data processing bijections and be chained together with the other bijections in your normalizing flow.
 
@@
 Data is handled in Pandas DataFrames \citep{pandas}, while the normalizing flows are implemented in Jax \citep{jax}, which allows for efficient, parallelizable, GPU-enabled calculations for very large data sets.
-The code is easily installable for the Python Package Index\footnote{\url{https://pypi.org/project/pzflow/}} (PyPI) and is hosted on Github,\footnote{\url{https://github.com/jfcrenshaw/pzflow}}.
+The code is easily installable from the Python Package Index\footnote{\url{https://pypi.org/project/pzflow/}} (PyPI) and is hosted on Github,\footnote{\url{https://github.com/jfcrenshaw/pzflow}}.
 The documentation\footnote{\url{https://jfcrenshaw.github.io/pzflow/}} includes tutorial notebooks demonstrating the features mentioned in this paper on different example problems.
@@
 The rest of this paper will demonstrate using PZFlow for the statistical modeling of galaxy catalogs.
-Section \ref{sec:galaxy-catalog} uses PZFlow to forward model a photometric catalog, including spectroscopic redshifts, Rubin photometry, ``true'' photo-z posteriors, ellipticities, and sizes.
+Section \ref{sec:galaxy-catalog} uses PZFlow to forward model a galaxy catalog, including photometry, spectroscopic redshifts (spec-z's), \emph{true} photo-z posteriors, ellipticities, and sizes.
 Section \ref{sec:photo-z} uses PZFlow for photo-z estimation, demonstrating the power of PZFlow as a density estimator, including numerous useful features for photo-z estimation.
@@
 
-In this section, we apply PZFlow to forward model a photometric galaxy catalog for the Rubin Observatory.
-The advantage of using a catalog generated from a normalizing flow is that we have direct access to the exact distribution from which the data is drawn, enabling us to calculate truth values for derived statistical products, such as the true photo-z redshift posterior for each galaxy.
+In this section, we use PZFlow to forward model a photometric galaxy catalog for the Vera Rubin Observatory's Legacy Survey of Space and Time (LSST; \citealt{ivezic2019}).
+The advantage of using a catalog generated from a normalizing flow is that we have direct access to the exact distribution from which the data is drawn, enabling us to calculate true values for derived statistical products, such as the \emph{true} photo-z redshift posterior for each galaxy.
 
@@
 In Section \ref{sec:true-posteriors}, we calculate true redshift posteriors for the new catalog.
-In Section \ref{sec:fwd-model-conditional} we build a conditional flow to add additional properties to the catalog.
+In Section \ref{sec:fwd-model-conditional} we build a conditional flow to add additional galaxy properties to the catalog.
 
@@
             Distribution of the CosmoDC2 test set compared to the distribution learned by PZFlow.
-            The close overlap of every pair-wise distribution demonstrates that PZFlow was able to learn the structure present in CosmoDC2 with high fidelity.
+            The close overlap of every pair-wise distribution demonstrates that PZFlow has learned the distribution in CosmoDC2 with high fidelity.
         }
@@
 
-To create a generative model of galaxy redshifts and photometry, we use the true redshifts and $ugrizy$ magnitudes from the DESC CosmoDC2 simulation \citep{dc2, cosmodc2}.
+To create a generative model of galaxy redshifts and photometry, we use the true redshifts and $ugrizy$ magnitudes from the CosmoDC2 simulation \citep{dc2, cosmodc2} of the LSST Dark Energy Science Collaboration (DESC).
 We selected all galaxies from CosmoDC2 with at most one band beyond the LSST 10-year 5-sigma point source depths.
-Of these, we selected $10^6$ galaxies and split them intro training and test sets consisting of 80\% and 20\% of the galaxies, respectively.
+Of these, we randomly selected $10^6$ galaxies and split them intro training and test sets consisting of 80\% and 20\% of the galaxies, respectively.
 
-For the latent distribution we use a 7 dimensional Uniform distribution over the range $[-5, 5]$.
+For the latent distribution we use a 7 dimensional Uniform distribution over the range $[-5, 5]$\footnote{The choice of 5 was arbitrary. Any other positive value would work just as well.}.
 To map the data onto the latent distribution, we use the following bijection:
@@
 
-The first layer of the bijection is a Color Transform.
-This bijection converts galaxy magnitudes to colors, but keeps the $i$ band magnitude as a proxy for the apparent luminosity:
+The first layer of the bijection is the Color Transform, a data processing bijection designed specifically for this task.
+The Color Transform converts galaxy magnitudes to colors, but keeps the $i$ band magnitude as a proxy for the apparent luminosity:
 \begin{multline}
@@
 We use $D=7$ layers to transform all 7 dimensions of our data, and set $B=5$ to match the support of the latent distribution.
-We use the coupling function (with two hidden layers of 128 neurons) described in \citet{durkan2019}.
+We use the coupling function (a feedforward neural network with two hidden layers of 128 neurons) described in \citet{durkan2019}.
 We use $K=16$ spline knots.
@@
 After training the flow (see Appendix \ref{app:training-details}), we assess the results by drawing $10^4$ galaxies from the trained flow, and plotting their distribution against  $10^4$ galaxies from the test set (Figure \ref{fig:main-corner}).
-We see that the normalizing flow has done an excellent job of reproducing the distribution of DESC DC2, without any unusual artifacts or outliers.
+We see the normalizing flow has done an excellent job of reproducing the distribution of galaxies in CosmoDC2, without any unusual artifacts or outliers.
 In addition, Figure \ref{fig:smooth-color-dist} compares the distribution of galaxy $r-i$ vs redshift.
@@
 With this normalizing flow, we have an efficient CosmoDC2 emulator that produces a smooth distribution of realistic galaxies up to high-redshifts.
-
-To simulate photometric errors, we use the LSST error model of \citet{ivezic2019}, except without the high-SNR assumptions made there (see Appendix \ref{app:error-model} for details).
+We use this emulator to generate a catalog with $10^4$ galaxies.
+We add photometric errors using the LSST error model of our PhotErr package (see Appendix \ref{app:error-model}).
+Importantly, since we have access to the probability distribution from which the galaxies were generated, we can calculate \emph{true} redshift posteriors for each galaxy.
+This is the subject of the next section.
 
@@
 Averaging $p(z|\hat{m})$ over the samples $\hat{m}$ yields $p(z|m, \sigma_m)$.
+This is the ``best case scenario'' redshift posterior that photo-z estimators hope to estimate.
 
@@
 You may wish to marginalize over all values of $n$ if the galaxy was not observed in that band.
-This might occur, for example, if you include photometry from Euclid, which will not have complete coverage of the LSST catalog \citep{euclid}.
+This might occur, for example, if you include photometry from the Euclid Space Telescope \citep{euclid}, which will not have complete coverage of the LSST catalog.
 You may also wish to marginalize over all values beyond a limiting magnitude if the galaxy was observed, but not detected in that band.
@@
 
-Figure \ref{fig:posteriors} shows a number of redshift posteriors to visualize the effects of error convolution and band marginalization.
+To visualize the impacts of error convolution and band marginalization, Figure \ref{fig:posteriors} shows a number of redshift posteriors for a single example galaxy.
 The black posterior is calculated using the true galaxy magnitudes, while the blue posteriors is calculated after adding photometric errors.
-Calculating the posterior using the noisy photometry results in a posterior with very little support over the true redshift.
+Calculating the posterior using the noisy photometry results in a biased posterior.
 The orange posterior has had the photometric errors convolved as in Equation \ref{eq:err-conv}.
@@
 
-Calculating these posteriors enables direct comparison of true redshift posteriors (consistent with the photometric errors and missing bands) with the redshift posteriors estimated by photo-z estimators.
-This allows you to test photo-z estimation on a posterior-per-galaxy basis, which will alleviate the photo-z validation problems discovered by \citet{schmidt2020} and will be investigated in future DESC papers.
+Calculating these posteriors enables direct comparison of true redshift posteriors (consistent with photometric errors and missing bands) with the redshift posteriors estimated by photo-z estimators.
+This is important, as modern cosmology analyses are beginning to rely on full redshift posteriors \citep{newman2022}.
+\citePZt showed that popular metrics for evaluating photo-z estimators using ensembles of photo-z posteriors can be misleading, and are not well suited to our science.
+PZFlow catalogs with \emph{true} redshift posteriors provide a path forward by enabling the evaluation of photo-z estimators on a per-posterior basis.
 
@@
         \caption{
-            Distribution of the ellipticities and sizes of the galaxies in the CosmoDC2 test set compared to the distribution learned by PZFlow.
-            The close overlap of every pair-wise distribution demonstrates that PZFlow was able to learn the structure present in CosmoDC2 with high fidelity.
+            Conditional distributions of the ellipticity and size of the galaxies in the CosmoDC2 test set compared to the distribution learned by PZFlow.
+            The close overlap of every pair-wise distribution demonstrates that PZFlow has learned the distribution in CosmoDC2 with high fidelity.
         }
@@
 
-For the latent distribution, we once more use a Uniform distribution over the range $[-5, 5]$.
+For the latent distribution, we again use a Uniform distribution over the range $[-5, 5]$.
 For the bijection, we use
@@
 \end{align}
-The RQ-RSC acts on the two dimensional space of size and ellipticity, but also takes the galaxy redshift and magnitudes as inputs.
+The RQ-RSC acts on the two dimensional space of size and ellipticity, but also takes the galaxy redshift and magnitudes as inputs (see Figure \ref{fig:conditional-coupling}).
 The redshifts and magnitudes are transformed to have zero mean and unit variance before being input to the neural network\footnote{These variables are standard scaled instead of mapped onto the domain [-5, 5], because the neural network that parameterizes the splines has no limit on inputs, unlike the splines themselves, which are limited to the range [-5, 5].} that parameterizes the splines.
@@
 
-After training the flow (see Appendix \ref{app:training-details}), we sample a size and ellipticity for each galaxy in the PZFlow catalog created in the previous section (using the true magnitudes), and plot the distribution of these features against the distribution in the test set (Figure \ref{fig:conditional-corner}).
-Once again, we see that the normalizing flow does a good job of emulating the CosmoDC2 distribution.
+After training the flow (see Appendix \ref{app:training-details}), we sample a size and ellipticity for each galaxy in the PZFlow catalog created in the previous section (conditioned on the true magnitudes), and plot the distribution of these features against the distribution in the test set (Figure \ref{fig:conditional-corner}).
+Once again, we see the normalizing flow does a good job of emulating the CosmoDC2 galaxy distribution.
 
@@
             Note these galaxies were specifically chosen for their broad, multimodal posteriors.
-            The posteriors of most galaxies are unimodal with sharp peaks.
+            The posteriors of most galaxies are sharp and unimodal.
         }
@@
 This makes them useful tools for estimating posterior distributions for galaxy properties, conditioned on observed features of the galaxy.
-A common example of this in cosmology is photometric redshift estimation, in which you try to estimate the redshift of a galaxy using its magnitude in several photometric bands.
-In this section, we will demonstrate this application to show how normalizing flows can be used for density estimation.
+A common example of this in cosmology is photometric redshift estimation, in which you estimate the redshift of a galaxy using its magnitude in several photometric bands.
+In this section, we use PZFlow as a photo-z estimator to demonstrate using normalizing flows for density estimation.
 
@@
 
-A full marginalization over the model parameters would be too computationally expensive, so instead we opt for an ensemble of normalizing flows.
-In other words, we will train multiple normalizing flows under identical conditions, just with different random initializations of the model parameters.
-This will allow the optimization algorithm to explore different basins of attraction in the parameter space, allowing us to marginalize over different potential models.
-In the machine learning literature, this is known as a Deep Ensemble \citep{lakshminarayanan2017}, a popular method for approximate bayesian marginalization \citep{wilson2020,fort2020}.
+A full marginalization over the model parameters would be too computationally expensive, so instead we approximate this marginalization using an ensemble of normalizing flows.
+In other words, we train multiple normalizing flows under identical conditions, using different random initializations of the model parameters.
+This allows the optimization algorithm to explore different basins of attraction in the parameter space.
+In the machine learning literature, this is known as a Deep Ensemble \citep{lakshminarayanan2017}, and is a popular method for approximate bayesian marginalization \citep{wilson2020,fort2020}.
 
-We train an ensemble of 4 normalizing flows, each with the same architecture and training schedule as the flow described in Section \ref{sec:galaxy-catalog}.
-PZFlow makes this easy to do automatically, by simply swapping \texttt{FlowEnsemble} for \texttt{Flow} in the code.
+We train an ensemble of 4 normalizing flows, each with the same architecture and training schedule as the regular flow described in Section \ref{sec:galaxy-catalog}.
+With PZFlow, this is as simple as swapping \texttt{FlowEnsemble} for \texttt{Flow} in the code.
 
@@
 Each galaxy in the training set has a true redshift and observed magnitudes in the $ugrizy$ bands, with corresponding photometric errors.
-To account for the photometric error, at the start of each training epoch, we resample the training set from the photometric error distribution.
-In other words, each epoch, we resample
+To account for the photometric error, at the start of each training epoch, we resample the training set from the photometric error distributions.
+In other words, each epoch, for each galaxy, we sample
 \begin{align}
@@
 This allows our ensemble of flows to approximate the distribution $p(z, m)$.
-
-The training curves for the flows in the ensemble can be seen in Appendix \ref{app:training-details}.
+For more details on training the ensemble, see Appendix \ref{app:training-details}.
 
@@
 
-Each flow in the ensemble estimates the posterior for each galaxy by marginalizing over the photometric errors:
+For each flow in the ensemble, we estimate $p(z, m)$ by marginalizing over the photometric errors:
 \begin{align}
@@
 which is estimated by sampling $m \sim p(\hat{m}, \sigma_m)$ and averaging $p(z, m)$ over these samples.
-We then average the posteriors from each flow in the ensemble.
+We then average the $p(z, m)$ from each flow, and normalize with respect to the redshift grid.
+This provides a redshift posterior for each galaxy.
 
-An example of this for three galaxies can be seen in Figure \ref{fig:ensemble-posteriors}.
+Posteriors for three galaxies can be seen in Figure \ref{fig:ensemble-posteriors}.
 Each flow produces a PDF which may contain slightly different features in each case.
 By averaging over the individual posteriors, we select for features that are common between models, while smoothing over features that are present in only a single model.
-Compare these to the true posteriors.
+\note{Compare these to the true posteriors.}
 
-We can calculate posteriors for galaxies with missing magnitudes by marginalizing over the missing magnitudes.
-This is done by averaging $p(z, m)$ over a grid of values for the missing bands. 
-This might be useful when not every galaxy is observed in every band, in which case you can marginalize over the bands that were not observed.
-Or if a galaxy was observed but not detected in the band, you can marginalize over values for the band that are beyond the limiting magnitude.
-You can also marginalize over bands for which you have a magnitude to understand what information about redshift is included in that band.
+We can calculate posteriors for galaxies with missing magnitudes by marginalizing over the missing magnitudes, as described in Section \ref{sec:true-posteriors}.
+An example is shown in Figure \ref{fig:posterior_marginalized}, which compares a posterior calculated using every LSST band to a posterior with the $u$ band marginalized.
+Sacrificing the information in the $u$ band increases the uncertainty of the redshift inference.
+In this instance, a second potential higher-redshift mode is created.
+Conversely, this demonstrates how the additon of $u$ band information rules out the higher redshift solution for this galaxy.
 
@@
 
-Figure \ref{fig:posterior_marginalized} compares posteriors with and without $u$ band information for an example galaxy.
-It is clear how losing the $u$ band information increases the uncertainty in the redshift inference.
-In this instance, a second potential redshift mode is created.
-Conversely, this demonstrates how the addition of $u$ band information rules out the higher redshift solution for this galaxy.
-
 \subsection{Photo-z metrics}
 
-In this section, we calculate common photo-z metrics to evaluate how well the ensemble of flows captures the information present in the catalog.
-We calculate each metric for the full catalog and the sub catalog with missing $u$ bands.
-Note that these metrics are optimistic in the sense that the training set is representative of the test set, which is usually not the case when applied to real galaxies.
+In this section, we evaluate the performance of PZFlow using common photo-z metrics.
+Note these metrics are optimistic in the sense that the training set is representative of the test set, which is usually not the case in modern cosmology applications.
 
 The most common metrics for photo-z estimation concern photo-z point estimates, which are a compression of the photo-z posterior to a single redshift estimate.
-We make the common choice of selecting the mode of the posteriors\footnote{The mean redshift is often a poor choice, since photo-z posteriors are generally multimodal, and so the mean value can lie between two modes at a redshift with very small probability density.}.
+We make the common choice of selecting the mode of the posteriors\footnote{The mean redshift is a poor choice, since photo-z posteriors are often multimodal, and so the mean value can lie between two modes at a redshift with very small probability density.}.
 We compute metrics of the quantity $\Delta z = (z_\text{phot} - z_\text{true}) / (1 + z_\text{true})$, where the denominator accounts for naturally greater uncertainties at high redshift.
 
-Figure \ref{fig:binned-metrics} shows the metrics from \citet{descSRD} as a function of true redshift.
-The bias is defined as the median of $\Delta z$; the scatter is defined as $\text{IQR} / 1.349$, where IQR is the interquartile range of $\Delta z$; the outlier fraction is defined as the fraction of galaxies for which $\Delta z$ is greater than three times the $\Delta z$ scatter.
-As a point of comparison, we plot the requirements for LSST cosmology in gray.
+Figure \ref{fig:point-estimates} compares the photo-z point estimates to the true redshifts.
+The point estimates for most galaxies lie along the diagonal, indicating strong performance.
+There are the common photo-z ``wings'', indicating redshifts where important spectral features are transitioning between neighboring photometric bands.
+There is also a small population of high-redshift objects that were mistakenly identified as very low redshift objects.
+This point estimate plot is comparable to other high-performance machine learning photo-z estimators when provided with representative training sets \needscite.
 
-Like many photo-z estimators, our estimator performs well to a redshift of approximately 1.5.
+Figure \ref{fig:binned-metrics} shows the photo-z point estimate metrics from the LSST DESC Science Requirements Document (SRD; \citealt{descSRD}) as a function of true redshift.
+The \emph{bias} is defined as the median of $\Delta z$; the \emph{scatter} is defined as $\text{IQR} / 1.349$, where IQR is the interquartile range of $\Delta z$; the \emph{outlier fraction} is defined as the fraction of galaxies for which $\Delta z$ is greater than three times the scatter.
+The requirements from the SRD are plotted in black to provide a sense of scale.
+
+Like many photo-z estimators, PZFlow performs well to a redshift of approximately 1.5.
 At higher redshifts, our estimator does not meet the bias and scatter requirements, because there is very little training data in this redshift range.
-We note however that for many cosmology application, it is okay for the bias to exceed the required limits, as long as the bias can be well determined via some calibration process \citep{newman2022}.
+We note however that for many cosmology applications, it is okay for the bias to exceed the required limits, as long as the bias can be well determined via some calibration process \citep{newman2022}.
 
@@
         \caption{
-            Comparing the photo-z point estimates to the true redshifts in the catalog.
+            Photo-z point estimates (maximum a posteriori) vs true redshift for galaxies in the test set.
         }
@@
 
-Another common metric is the probability integral transform (PIT) (see e.g. \citealt{schmidt2020, dey2022}).
+Another common metric is the probability integral transform (PIT) (see e.g. \citePZa, \citealt{dey2022}).
 The PIT metric is used to determine if posteriors are well calibrated, i.e. if true values fall within X\% confidence intervals X\% of the time.
@@
 The low values at the edges of the PIT histogram indicate the relative rarity of catastrophic outliers, which is also reflected in the far right panel of Figure \ref{fig:binned-metrics}, where you can see that our estimator meets the requirement on the outlier fraction at all redshifts.
+There is also a slight rightward tilt.
+This indicates a small negative bias, which reflects the intrinsic prior towards smaller redshifts, as this is where the majority of galaxies in the training set lie.
+This negative bias is also visible in the far left panel of Figure \ref{fig:binned-metrics}.
 
@@
         \caption{
-            The bias, scatter, and catastrophic outlier fraction of the photo-z point estimates as a function of true galaxy redshift.
-            The dashed black lines represent the science requirements for LSST cosmology as stated in the LSST SRD \citep{descSRD}.
+            The bias, scatter, and outlier fraction of the photo-z point estimates as a function of true galaxy redshift.
+            The dashed black lines represent the requirements for LSST cosmology as stated in the LSST DESC SRD \citep{descSRD}.
             These lines are to provide a sense of scale for these metrics.
@@
 
-The previous metrics analyze photo-z performance for point estimates, which are insufficient for modern cosmology \citep{newman2022}, and for ensembles of posteriors, which is often not a good indicator of photo-z performance for science applications \citep{schmidt2020}.
-The methods introduced in this paper allow one to create galaxy catalogs for which each galaxy has a true redshift posterior, which will enable more comprehensive evaluation of photo-z estimators.
-A comprehensive evaluation of the full posteriors from photo-z estimators is a goal of our future work.
+The previous metrics analyze photo-z performance for point estimates, which are insufficient for modern cosmology \citep{newman2022}, and for ensembles of posteriors, which is often misleading and not a good indicator of performance for science applications \citePZp.
+The methods introduced in this paper enable the creation of galaxy catalogs for which each galaxy has a \emph{true} redshift posterior, which will enable more comprehensive evaluation of photo-z estimators.
+An evaluation of the full posteriors from many photo-z estimators is a major goal of future DESC papers.
 
@@
 
-In this paper we introduced PZFlow, a normalizing flow package in Python, designed for the statistical modeling of astronomical data sets.
+In this paper we introduced PZFlow, a normalizing flow package in Python, designed for the statistical modeling of tabular astronomical data.
 We used PZFlow to forward model galaxy catalogs that include photometry, redshifts, sizes, and ellipticities.
-In addition, each galaxy in our catalog has a true redshift posterior, which can be convolved with measurement errors.
+In addition, each galaxy in our catalog has a \emph{true} redshift posterior, which can be convolved with measurement errors.
 These true posteriors allow a direct evaluation of the posteriors produced by photo-z estimators.
-A similar comparison can be made to posterior estimates for any other galaxy property.
+A similar comparison can be made to posterior estimates for any other galaxy properties.
 
@@
 This is because only the full redshift posteriors allow one to account for degeneracies in color-redshift space, which will otherwise bias cosmological inference.
-Previous evaluations of photo-z performance focused on point estimates and metrics for ensembles of posteriors, but \citet{schmidt2020} showed these analyses are inadequate for modern cosmology.
-This work enables a future analysis of photo-z estimators on a per-posterior basis, which is a major goal of the LSST DESC.
+Previous evaluations of photo-z performance have focused on point estimates and metrics for ensembles of posteriors, but \citePZt demonstrated these metrics are misleading and inadequate for modern cosmology.
+This work enables future analysis of photo-z estimators on a per-posterior basis, which is a major goal of the LSST DESC.
 
@@
 PZFlow achieves high accuracy with relatively little fine tuning and with very few modeling assumptions.
-PZFlow will be a powerful tool for the statistical analysis of tabular astronomical data.
+This makes PZFlow a powerful tool for the statistical analysis of tabular astronomical data.
 
-Note that the code to reproduce the figures in this paper is hosted publicly on Github\footnote{\url{https://github.com/jfcrenshaw/pzflow-paper}}, and the code for each individual figure can be reached by clicking on the Github logo in the margin next to the figure.
+The code to reproduce the figures in this paper is hosted publicly on Github\footnote{\url{https://github.com/jfcrenshaw/pzflow-paper}}, and the code for each individual figure can be found by clicking on the Github logo in the margin next to that figure.
 
@@
 
-For each of the flows that make up the flow ensemble in \ref{sec:photo-z}, we trained for 150 epochs using the Adam optimizer.
+For each of the flows that make up the flow ensemble in Section \ref{sec:photo-z}, we trained for 150 epochs using the Adam optimizer.
 We started each with a learning rate of $10^{-4}$, and decreased the learning rate by a factor of 10 every 50 epochs.
@@
 We have implemented this error model, along with several other extensions to it, in the Python package PhotErr, which is available on the Python Package Index\footnote{\url{https://pypi.org/project/photerr/}} (PyPI), and Github\footnote{\url{https://github.com/jfcrenshaw/photerr}}.
-The extensions include different methods for handling non-detections, methods for modeling errors of extended objects \citep{vandenbusch2020,kuijken2019}, and error models for the Roman and Euclid space telescopes \citep{roman,euclid,graham2020}, however, these extensions are not used in this paper.
+The extensions include different methods for handling non-detections, methods for modeling errors of extended objects \citep{vandenbusch2020,kuijken2019}, and error models for the Roman and Euclid space telescopes \citep{roman,euclid,graham2020}.
+These extensions, however, are not used in this paper.
 

### 2022-12-02 src/scripts/forward_model/train_conditional_galaxy_flow.py

@@ -45,13 +45,13 @@ flow = Flow(
 # train for three rounds of 150 epochs
 # after each round, decrease the learning rate by a factor of 10
 opt = optax.adam(1e-3)
-losses = flow.train(train_set, epochs=150, optimizer=opt, seed=0)
+losses = flow.train(train_set, epochs=150, optimizer=opt, seed=0, verbose=True)
 
 opt = optax.adam(1e-4)
-losses += flow.train(train_set, epochs=150, optimizer=opt, seed=1)
+losses += flow.train(train_set, epochs=150, optimizer=opt, seed=1, verbose=True)
 
 opt = optax.adam(1e-5)
-losses += flow.train(train_set, epochs=150, optimizer=opt, seed=2)
+losses += flow.train(train_set, epochs=150, optimizer=opt, seed=2, verbose=True)
 
 # save some info with the model
 flow.info = (

### 2022-12-02 src/scripts/forward_model/train_main_galaxy_flow.py

@@ -56,13 +56,13 @@ flow = Flow(train_set.columns, bijector=bijector)
 # train for three rounds of 50 epochs
 # after each round, decrease the learning rate by a factor of 10
 opt = optax.adam(1e-3)
-losses = flow.train(data, epochs=50, optimizer=opt, seed=0)
+losses = flow.train(data, epochs=50, optimizer=opt, seed=0, verbose=True)
 
 opt = optax.adam(1e-4)
-losses += flow.train(data, epochs=50, optimizer=opt, seed=1)
+losses += flow.train(data, epochs=50, optimizer=opt, seed=1, verbose=True)
 
 opt = optax.adam(1e-5)
-losses += flow.train(data, epochs=50, optimizer=opt, seed=2)
+losses += flow.train(data, epochs=50, optimizer=opt, seed=2, verbose=True)
 
 # save some info with the model
 flow.info = (

### 2022-12-03 src/scripts/forward_model/plot_galaxy_losses.py
More citations. Figure fixes.

@@ -15,7 +15,7 @@ fig, (ax1, ax2) = plt.subplots(1, 2, figsize=(7, 3), constrained_layout=True)
 
 main_epochs = np.arange(len(main_losses))
 ax1.plot(main_epochs, main_losses)
-ax1.set(xlabel="Epochs", ylabel="Loss", title="Main flow")
+ax1.set(xlabel="Epochs", ylabel="Loss", title="Regular flow")
 
 conditional_epochs = np.arange(len(conditional_losses))
 ax2.plot(conditional_epochs[10:], conditional_losses[10:])

### 2022-12-03 src/scripts/photo-z/plot_binned_metrics.py

@@ -19,7 +19,7 @@ z_phot = grid[pdfs.argmax(axis=1)]
 
 # load the true redshifts
 test_set = load_pzflow_catalog(subset="test")
-z_spec = test_set.redshift[: len(z_phot)]
+z_spec = test_set.redshift.iloc[: len(z_phot)]
 
 # calculate the weighted error values
 ez = (z_phot - z_spec) / (1 + z_spec)

### 2022-12-03 src/scripts/photo-z/plot_ensemble_losses.py

@@ -13,8 +13,8 @@ with open(paths.data / "pz_ensemble/losses.pkl", "rb") as file:
 
 # plot the losses
 fig, ax = plt.subplots(figsize=(3, 3), constrained_layout=True)
-for flow, loss in losses.items():
-    ax.plot(loss, alpha=0.5, label=flow)
+for i, loss in enumerate(losses.values()):
+    ax.plot(loss, alpha=0.5, label=f"Flow {i+1}")
 ax.legend()
 ax.set(xlabel="Epochs", ylabel="Loss", ylim=(-1.8, -1))
 

### 2022-12-03 src/scripts/photo-z/plot_pz_point_estimates.py

@@ -18,7 +18,7 @@ z_phot = grid[pdfs.argmax(axis=1)]
 
 # load the true redshifts
 test_set = load_pzflow_catalog(subset="test")
-z_spec = test_set.redshift[: len(z_phot)].to_numpy()
+z_spec = test_set.redshift.iloc[: len(z_phot)].to_numpy()
 
 # select the non-zero z_phots
 idx = np.where(z_phot > 0)

### 2022-12-03 src/scripts/intro/train_two_moons_flow.py
Fixed directory inputs in snakefile. Two moons now verbose.

@@ -11,7 +11,7 @@ data = get_twomoons_data()
 
 # create and train a flow to model the two moons data
 flow = Flow(data.columns)
-_ = flow.train(data)
+_ = flow.train(data, verbose=True)
 
 # save the trained flow
 flow.save(paths.data / "two_moons_flow.pzflow.pkl")

### 2022-12-06 src/scripts/photo-z/plot_ensemble_posteriors.py
Final tweaks before sending to coauthors.

@@ -1,4 +1,5 @@
 """Plot posterior ensembles for a few galaxies."""
+# %%
 import matplotlib.pyplot as plt
 import numpy as np
 from load_pzflow_catalog import load_pzflow_catalog
@@ -15,7 +16,7 @@ flowEns = FlowEnsemble(file=paths.data / "pz_ensemble" / "pz_ensemble.pzflow.pkl
 test_set = load_pzflow_catalog(subset="test")
 
 # select galaxies for plotting
-idx = [39, 52, 44]
+idx = [87, 42, 99]
 test_set = test_set.iloc[idx]
 
 # calculate posteriors
@@ -49,3 +50,5 @@ for i, ax in enumerate(axes):
 axes[1].legend(loc="upper right")
 
 fig.savefig(paths.figures / "ensemble_posteriors.pdf")
+
+# %%

### 2022-12-06 src/scripts/photo-z/plot_marginalized_posterior.py

@@ -29,7 +29,7 @@ fig, (ax1, ax2) = plt.subplots(
 )
 
 # plot the true redshift
-idx = 10
+idx = 18
 z = test_set["redshift"].iloc[idx]
 ax1.axvline(z, c="C3", ls="--", label="True redshift")
 

### 2022-12-09 src/tex/ms.tex
Fixed showyourwork dependency.
Author contributions.

@@
 % Define document class
-\documentclass[twocolumn]{aastex631}
+\documentclass[twocolumn,twocolappendix]{aastex631}
 \usepackage{showyourwork}
+
+\usepackage{lipsum}
 
@@
 
+\newcommand{\Illinois}{Department of Astronomy, University of Illinois at Urbana-Champaign, 1002 W. Green St., IL 61801, USA}
+\newcommand{\NCSA}{National Center for Supercomputing Applications, Urbana, IL, 61801, USA}
+\newcommand{\CAPS}{Center for AstroPhysical Surveys, Urbana, IL, 61801, USA}
+\newcommand{\NSF}{National Science Foundation Graduate Research Fellow}
+
 \shorttitle{Galaxy Catalogues with Normalizing Flows}
@@
 % Author list
-\correspondingauthor{John Franklin Crenshaw}
+\correspondingauthor{John~Franklin~Crenshaw}
 \email{jfc20@uw.edu}
 
-\author[0000-0002-2495-3514]{John Franklin Crenshaw}
-\affiliation{DIRAC Institute and Department of Physics, University of Washington, Seattle, WA 98195, USA}
+\author[0000-0002-2495-3514]{John~Franklin~Crenshaw}
+\affiliation{DIRAC Institute and the Department of Physics, University of Washington, Seattle, WA 98195, USA}
 
-\author{J. Bryce Kalmbach}
+\author[0000-0002-6825-5283]{J.~Bryce~Kalmbach}
+\affiliation{DIRAC Institute and the Department of Astronomy, University of Washington, Seattle, WA 98195, USA}
 
-\author{Alexander Gagliano}
+\author[0000-0003-4906-8447]{Alexander~Gagliano}
+\affiliation{\Illinois}
+\affiliation{\NCSA}
+\affiliation{\CAPS}
+\affiliation{\NSF}
 
-\author{Ziang Yan}
+\author[0000-0001-8043-5378]{Ziang Yan}
+\affiliation{Ruhr University Bochum, Faculty of Physics and Astronomy, Astronomical Institute (AIRUB), German Centre for Cosmological Lensing, 44780 Bochum, Germany}
 
-\author{Andrew Connolly}
+\author[0000-0001-5576-8189]{Andrew~J.~Connolly}
+\affiliation{DIRAC Institute and the Department of Astronomy, University of Washington, Seattle, WA 98195, USA}
 
-\author{Francois Lanusse}
+\author{François~Lanusse}
 
-\author{Alex Malz}
+\author{Alex~I.~Malz}
 
-\author{Sam Schmidt}
+\author[0000-0002-5091-0470]{Samuel~J.~Schmidt}
+\affiliation{Department of Physics and Astronomy, University of California, One Shields Avenue, Davis, CA 95616, USA}
 
-\collaboration{99}{(The LSST Dark Energy Science Collaboration)}
+\collaboration{99}{The LSST Dark Energy Science Collaboration}
 
@@
     Normalizing flows are powerful tools for learning high-dimensional probability distributions from samples thereof, and have many applications in astronomy, including posterior estimation and forward modeling.
-    We introduce PZFlow, a Python package for the statistical modeling of tabular data using normalizing flows.
+    We introduce PZFlow, a Python package for statistical modeling of tabular data using normalizing flows.
     We use PZFlow to model photometric galaxy catalogs including redshifts, LSST photometry, size, and ellipticity, and generate a synthetic catalog.
@@
 \label{sec:intro}
+
+\begin{figure*}[t]
+    \script{intro/plot_two_moons.py}
+    \begin{centering}
+        \includegraphics{figures/twomoons_example.pdf}
+        \caption{
+            A normalizing flow demonstrated on the two moons data set from scikit-learn.
+            The two moons data on the left is mapped onto a two dimensional uniform distribution by the bijection $f$.
+            The data are colored by quadrant to visualize their image in the latent space.
+            You can sample the data distribution by sampling from the uniform distribution, and using $f^{-1}$ to map the samples back to the data space.
+        }
+        \label{fig:two-moons}
+    \end{centering}
+\end{figure*}
 
@@
 
-Once trained, a normalizing flow can forward model data with realistic statistical correlations, and since normalizing flows provide exact likelihood inference, each property of the generated data has a \emph{true} posterior.
+Normalizing flows have gained popularity as tools for efficient and flexible sampling for parameter inference (e.g., \citealt{dai2022,dacunha2022,hassan2022}).
+In this paper, we focus on photo-z's and using normalizing flows to generate synthetic catalogs.
+Since normalizing flows provide exact likelihood inference, the properties of objects in these catalogs have \emph{true} posteriors.
 When calculating these posteriors, you can convolve error distributions, and marginalize over chosen properties.
 Catalogs with \emph{true} posteriors are useful for the testing and validation of analysis pipelines that estimate posteriors, such as the photo-z estimators used in much of astrophysics and cosmology.
+Previous evaluations of these estimators have focused on comparing point estimates to true values (e.g., \citealt{hildebrandt2010,sanchez2014,graham2018}), or evaluating ensembles of posteriors \citePZp.
+Normalizing flow catalogs with \emph{true} posteriors open up a new, more comprehensive avenue for evaluation of these estimators by enabling posterior-to-posterior comparison.
 
@@
 \label{sec:nf}
-
-\begin{figure*}[t!]
-    \script{intro/plot_two_moons.py}
-    \begin{centering}
-        \includegraphics{figures/twomoons_example.pdf}
-        \caption{
-            A normalizing flow demonstrated on the two moons data set from scikit-learn.
-            The two moons data on the left is mapped onto a two dimensional uniform distribution by the bijection $f$.
-            The data are colored by quadrant to visualize their image in the latent space.
-            You can sample the data distribution by sampling from the uniform distribution, and using $f^{-1}$ to map the samples back to the data space.
-        }
-        \label{fig:two-moons}
-    \end{centering}
-\end{figure*}
 
@@
 
-\begin{figure*}[t!]
+\begin{figure*}[t]
     \script{forward_model/plot_main_galaxy_corner.py}
@@
 
-\begin{figure*}[t!]
+\begin{figure*}[t]
     \script{forward_model/plot_smooth_color_distribution.py}
@@
 
-\begin{figure}[t!]
+\begin{figure}[t]
     \script{forward_model/plot_posteriors.py}
@@
 
-To visualize the impacts of error convolution and band marginalization, Figure \ref{fig:posteriors} shows a number of redshift posteriors for a single example galaxy.
-The black posterior is calculated using the true galaxy magnitudes, while the blue posteriors is calculated after adding photometric errors.
-Calculating the posterior using the noisy photometry results in a biased posterior.
-The orange posterior has had the photometric errors convolved as in Equation \ref{eq:err-conv}.
-Convolving the errors broadens the posterior so that the true redshift lies within the support of the posterior.
-This broadening reflects the increased uncertainty due to the photometric errors.
-Finally, the green posterior has had the $u$ band marginalized.
-This posterior favors lower redshifts, although has a high-redshift tail that covers the support of the previous posteriors. 
-This demonstrates how the $u$ band helps constrain the redshift, and the loss of this information leads to greater uncertainty.
-
-Calculating these posteriors enables direct comparison of true redshift posteriors (consistent with photometric errors and missing bands) with the redshift posteriors estimated by photo-z estimators.
-This is important, as modern cosmology analyses are beginning to rely on full redshift posteriors \citep{newman2022}.
-\citePZt showed that popular metrics for evaluating photo-z estimators using ensembles of photo-z posteriors can be misleading, and are not well suited to our science.
-PZFlow catalogs with \emph{true} redshift posteriors provide a path forward by enabling the evaluation of photo-z estimators on a per-posterior basis.
-
-\subsection{Additional properties with conditional flows}
-\label{sec:fwd-model-conditional}
-
-\begin{figure*}[t!]
+% this figure is so high just to move it ahead one page in the document
+\begin{figure*}[t]
     \script{forward_model/plot_conditional_galaxy_corner.py}
@@
 \end{figure*}
+
+To visualize the impacts of error convolution and band marginalization, Figure \ref{fig:posteriors} shows a number of redshift posteriors for a single example galaxy.
+The black posterior is calculated using the true galaxy magnitudes, while the blue posteriors is calculated after adding photometric errors.
+Calculating the posterior using the noisy photometry results in a biased posterior.
+The orange posterior has had the photometric errors convolved as in Equation \ref{eq:err-conv}.
+Convolving the errors broadens the posterior so that the true redshift lies within the support of the posterior.
+This broadening reflects the increased uncertainty due to the photometric errors.
+Finally, the green posterior has had the $u$ band marginalized.
+This posterior favors higher redshifts, but still comfortably agrees with the true redshift.
+This demonstrates how the $u$ band helps constrain the redshift, and the loss of this information leads to greater uncertainty.
+
+Calculating these posteriors enables direct comparison of true redshift posteriors (consistent with photometric errors and missing bands) with the redshift posteriors estimated by photo-z estimators.
+This is important, as modern cosmology analyses are beginning to increasingly rely on full redshift posteriors \citep{mandelbaum2008,newman2022}.
+\citePZt showed that popular metrics for evaluating photo-z estimators using ensembles of photo-z posteriors can be misleading, and are not well suited to our science.
+PZFlow catalogs with \emph{true} redshift posteriors provide a path forward by enabling the evaluation of photo-z estimators on a per-posterior basis.
+
+\subsection{Additional properties with conditional flows}
+\label{sec:fwd-model-conditional}
 
@@
 
-\begin{figure*}[t!]
+\begin{figure*}[t]
     \script{photo-z/plot_ensemble_posteriors.py}
@@
 By averaging over the individual posteriors, we select for features that are common between models, while smoothing over features that are present in only a single model.
-\note{Compare these to the true posteriors.}
+We can also treat the ensemble of posteriors as a distribution over possible posteriors, which will allow for more consistent error calibration in cosmological analyses \citep{zhang2023}.
 
@@
 
-\begin{figure}[t!]
+\begin{figure}[t]
     \script{photo-z/plot_marginalized_posterior.py}
@@
             The black posterior is estimated using the information from all 6 LSST bands (and includes convolution of the photometric errors).
-            The gold posterior has had the $u$ band marginalized, which opens up a second high-redshift mode.
+            The orange posterior has had the $u$ band marginalized, which opens up a second high-redshift mode.
         }
@@
 
-In this section, we evaluate the performance of PZFlow using common photo-z metrics.
-Note these metrics are optimistic in the sense that the training set is representative of the test set, which is usually not the case in modern cosmology applications.
-
-The most common metrics for photo-z estimation concern photo-z point estimates, which are a compression of the photo-z posterior to a single redshift estimate.
-We make the common choice of selecting the mode of the posteriors\footnote{The mean redshift is a poor choice, since photo-z posteriors are often multimodal, and so the mean value can lie between two modes at a redshift with very small probability density.}.
-We compute metrics of the quantity $\Delta z = (z_\text{phot} - z_\text{true}) / (1 + z_\text{true})$, where the denominator accounts for naturally greater uncertainties at high redshift.
-
-Figure \ref{fig:point-estimates} compares the photo-z point estimates to the true redshifts.
-The point estimates for most galaxies lie along the diagonal, indicating strong performance.
-There are the common photo-z ``wings'', indicating redshifts where important spectral features are transitioning between neighboring photometric bands.
-There is also a small population of high-redshift objects that were mistakenly identified as very low redshift objects.
-This point estimate plot is comparable to other high-performance machine learning photo-z estimators when provided with representative training sets \needscite.
-
-Figure \ref{fig:binned-metrics} shows the photo-z point estimate metrics from the LSST DESC Science Requirements Document (SRD; \citealt{descSRD}) as a function of true redshift.
-The \emph{bias} is defined as the median of $\Delta z$; the \emph{scatter} is defined as $\text{IQR} / 1.349$, where IQR is the interquartile range of $\Delta z$; the \emph{outlier fraction} is defined as the fraction of galaxies for which $\Delta z$ is greater than three times the scatter.
-The requirements from the SRD are plotted in black to provide a sense of scale.
-
-Like many photo-z estimators, PZFlow performs well to a redshift of approximately 1.5.
-At higher redshifts, our estimator does not meet the bias and scatter requirements, because there is very little training data in this redshift range.
-We note however that for many cosmology applications, it is okay for the bias to exceed the required limits, as long as the bias can be well determined via some calibration process \citep{newman2022}.
-
-\begin{figure}[t!]
+\begin{figure}[t]
     \script{photo-z/plot_pz_point_estimates.py}
@@
 \end{figure}
+
+\begin{figure*}[t]
+    \script{photo-z/plot_binned_metrics.py}
+    \begin{centering}
+        \includegraphics{figures/binned_metrics.pdf}
+        \caption{
+            The bias, scatter, and outlier fraction of the photo-z point estimates as a function of true galaxy redshift.
+            The dashed black lines represent the requirements for LSST cosmology as stated in the LSST DESC SRD \citep{descSRD}.
+            These lines are to provide a sense of scale for these metrics.
+            You can see that PZFlow meets the bias and scatter requirements up to redshift $\sim$ 1.5, while meeting the outlier fraction requirements for all redshifts.
+            We note that individual redshifts do not actually need to meet the bias requirement as long as the bias can be well calibrated via some other source, e.g. galaxy clustering.
+        }
+        \label{fig:binned-metrics}
+    \end{centering}
+\end{figure*}
+
+\begin{figure}[t]
+    \script{photo-z/plot_pit_histogram.py}
+    \begin{centering}
+        \includegraphics{figures/pit_histogram.pdf}
+        \caption{
+            The probability integral transform (PIT) histogram for PZFlow photo-z posteriors.
+            The PIT characterizes the calibration of the estimated posteriors, with the horizontal black line indicating perfect calibration.
+        }
+        \label{fig:pit-histogram}
+    \end{centering}
+\end{figure}
+
+In this section, we evaluate the performance of PZFlow using common photo-z metrics.
+Note these metrics are optimistic in the sense that the training set is representative of the test set, which is usually not the case in modern cosmology applications.
+
+The most common metrics for photo-z estimation concern photo-z point estimates, which are a compression of the photo-z posterior to a single redshift estimate (e.g., \citealt{hildebrandt2010,sanchez2014}).
+We make the common choice of selecting the mode of the posteriors\footnote{The mean redshift is a poor choice, since photo-z posteriors are often multimodal, and so the mean value can lie between two modes at a redshift with very small probability density.}.
+We compute metrics of the quantity $\Delta z = (z_\text{phot} - z_\text{true}) / (1 + z_\text{true})$, where the denominator accounts for naturally greater uncertainties at high redshift.
+
+Figure \ref{fig:point-estimates} compares the photo-z point estimates to the true redshifts.
+The point estimates for most galaxies lie along the diagonal, indicating strong performance.
+There are the common photo-z ``wings'', indicating redshifts where important spectral features are transitioning between neighboring photometric bands.
+There is also a small population of high-redshift objects that were mistakenly identified as very low redshift objects.
+This point estimate plot is comparable to other high-performance machine learning photo-z estimators when provided with representative training sets \citep{sanchez2014}.
+
+Figure \ref{fig:binned-metrics} shows the photo-z point estimate metrics from the LSST DESC Science Requirements Document (SRD; \citealt{descSRD}) as a function of true redshift.
+The \emph{bias} is defined as the median of $\Delta z$; the \emph{scatter} is defined as $\text{IQR} / 1.349$, where IQR is the interquartile range of $\Delta z$; the \emph{outlier fraction} is defined as the fraction of galaxies for which $\Delta z$ is greater than three times the scatter.
+The requirements from the SRD are plotted in black to provide a sense of scale.
+
+Like many photo-z estimators, PZFlow performs well to a redshift of approximately 1.5.
+At higher redshifts, our estimator does not meet the bias and scatter requirements, because there is very little training data in this redshift range.
+We note however that for many cosmology applications, it is okay for the bias to exceed the required limits, as long as the bias can be well determined via some calibration process \citep{newman2015}.
 
@@
 
-\begin{figure*}[t!]
-    \script{photo-z/plot_binned_metrics.py}
-    \begin{centering}
-        \includegraphics{figures/binned_metrics.pdf}
-        \caption{
-            The bias, scatter, and outlier fraction of the photo-z point estimates as a function of true galaxy redshift.
-            The dashed black lines represent the requirements for LSST cosmology as stated in the LSST DESC SRD \citep{descSRD}.
-            These lines are to provide a sense of scale for these metrics.
-            You can see that PZFlow meets the bias and scatter requirements up to redshift $\sim$ 1.5, while meeting the outlier fraction requirements for all redshifts.
-            We note that individual redshifts do not actually need to meet the bias requirement as long as the bias can be well calibrated via some other source, e.g. galaxy clustering.
-        }
-        \label{fig:binned-metrics}
-    \end{centering}
-\end{figure*}
-
 The previous metrics analyze photo-z performance for point estimates, which are insufficient for modern cosmology \citep{newman2022}, and for ensembles of posteriors, which is often misleading and not a good indicator of performance for science applications \citePZp.
 The methods introduced in this paper enable the creation of galaxy catalogs for which each galaxy has a \emph{true} redshift posterior, which will enable more comprehensive evaluation of photo-z estimators.
-An evaluation of the full posteriors from many photo-z estimators is a major goal of future DESC papers.
-
-\begin{figure}[t!]
-    \script{photo-z/plot_pit_histogram.py}
-    \begin{centering}
-        \includegraphics{figures/pit_histogram.pdf}
-        \caption{
-            The histogram of the probability integral transform (PIT) values of the photo-z posteriors.
-            This figure characterizes the calibration of the estimated posteriors.
-            In other words, does the true redshift lie at the center/in the tails of the posteriors as often as one would expect?
-            Ideally, this histogram should be flat, matching the dashed black line.
-            The fact that this distribution has a central peak indicates that the flow ensemble produces posteriors that are a little too conservative in their predictions.
-            The slight rightward tilt indicates that the posteriors also have a negative redshift bias.
-        }
-        \label{fig:pit-histogram}
-    \end{centering}
-\end{figure}
+An full evaluation of photo-z estimators on a posterior-by-posterior basis is a major goal of the LSST DESC.
 
@@
 
-The code to reproduce the figures in this paper is hosted publicly on Github\footnote{\url{https://github.com/jfcrenshaw/pzflow-paper}}, and the code for each individual figure can be found by clicking on the Github logo in the margin next to that figure.
+This paper was written using the showyourwork\footnote{\url{https://show-your.work/}} workflow manager.
+The code to reproduce this paper is hosted publicly on Github\footnote{\url{https://github.com/jfcrenshaw/pzflow-paper}}, and the code for each individual figure can be found by clicking on the Github logo in the margin next to that figure.
 
 \begin{acknowledgements}
-    JFC is supported by...
+    Author contributions are listed below: \\
+    J.~F.~Crenshaw: created PZFlow, designed experiments, wrote paper. \\
+    J.~B.~Kalmbach: wrote early normalizing flow code that evolved into PZFlow; photo-z estimation. \\
+    A.~Gagliano: validated code, developed use cases and associated tutorials; revised manuscript text. \\
+    Z.~Yan: contributed to PZFlow and PhotErr. \\
+    A.~J.~Connolly: discussion during development. \\
+    F.~Lanusse: wrote early normalizing flow code that evolved into PZFlow. \\
+    A.~I.~Malz: consulted on design and testing of PZFlow. \\
+    S.~J.~Schimdt: consulted on design and testing of PZFlow and PhotErr. \\
 \end{acknowledgements}
@@
 \label{app:training-details}
-
-\begin{figure*}[t!]
-    \script{forward_model/plot_galaxy_losses.py}
-    \begin{centering}
-        \includegraphics{figures/galaxy_flow_losses.pdf}
-        \caption{
-            Training losses for the galaxy flows.
-            Left: losses for the normal flow.
-            After epochs 50 and 100, you can see a drop in the loss due to the decrease in the learning rate.
-            Right: losses for the conditional flow.
-            After epochs 150 and 300, you can see a drop in the loss due to the decrease in the learning rate.
-        }
-        \label{fig:galaxy-flow-losses}
-    \end{centering}
-\end{figure*}
-
-\begin{figure}[t!]
-    \script{photo-z/plot_ensemble_losses.py}
-    \begin{centering}
-        \includegraphics{figures/ensemble_losses.pdf}
-        \caption{
-            Training losses for the four flows in the flow ensemble.
-            We have zoomed in to the bottom of the loss curve so you can see that each of the flows converges to a different minimum loss.
-        }
-        \label{fig:ensemble-losses}
-    \end{centering}
-\end{figure}
 
@@
 Apparently, each has found a different potential solution in the neural network's parameter space.
+
+\begin{figure*}[t!]
+    \script{forward_model/plot_galaxy_losses.py}
+    \begin{centering}
+        \includegraphics{figures/galaxy_flow_losses.pdf}
+        \caption{
+            Training losses for the normalizing flows.
+            Left: losses for the regular flow.
+            After epochs 50 and 100, you can see a drop in the loss due to the decrease in the learning rate.
+            Right: losses for the conditional flow.
+            After epochs 150 and 300, you can see a drop in the loss due to the decrease in the learning rate.
+        }
+        \label{fig:galaxy-flow-losses}
+    \end{centering}
+\end{figure*}
+
+\begin{figure}[t!]
+    \script{photo-z/plot_ensemble_losses.py}
+    \begin{centering}
+        \includegraphics{figures/ensemble_losses.pdf}
+        \caption{
+            Training losses for the four flows in the flow ensemble.
+            We have zoomed in to the bottom of the loss curve so you can see that each of the flows converges to a different minimum loss.
+        }
+        \label{fig:ensemble-losses}
+    \end{centering}
+\end{figure}
 
@@
 
-We have implemented this error model, along with several other extensions to it, in the Python package PhotErr, which is available on the Python Package Index\footnote{\url{https://pypi.org/project/photerr/}} (PyPI), and Github\footnote{\url{https://github.com/jfcrenshaw/photerr}}.
-The extensions include different methods for handling non-detections, methods for modeling errors of extended objects \citep{vandenbusch2020,kuijken2019}, and error models for the Roman and Euclid space telescopes \citep{roman,euclid,graham2020}.
-These extensions, however, are not used in this paper.
+We have implemented this error model, along with several other extensions, in the Python package PhotErr, which is available on the Python Package Index\footnote{\url{https://pypi.org/project/photerr/}} (PyPI), and Github\footnote{\url{https://github.com/jfcrenshaw/photerr}}.
+The extensions include different methods for handling non-detections, methods for modeling errors of extended objects (using models from \citealt{vandenbusch2020,kuijken2019}), and error models for the Roman and Euclid space telescopes \citep{roman,euclid,graham2020}.
 
-\bibliography{bib.bib}
+\bibliography{bib,nfbib}
 

### 2022-12-09 src/tex/ms.tex
Starting funding acknowledgements.

@@
 
-\newcommand{\Illinois}{Department of Astronomy, University of Illinois at Urbana-Champaign, 1002 W. Green St., IL 61801, USA}
-\newcommand{\NCSA}{National Center for Supercomputing Applications, Urbana, IL, 61801, USA}
-\newcommand{\CAPS}{Center for AstroPhysical Surveys, Urbana, IL, 61801, USA}
-\newcommand{\NSF}{National Science Foundation Graduate Research Fellow}
+\newcommand{\dirac}{DIRAC Institute and the Department of Astronomy, University of Washington, Seattle, WA 98195, USA}
 
@@
 \author[0000-0002-2495-3514]{John~Franklin~Crenshaw}
-\affiliation{DIRAC Institute and the Department of Physics, University of Washington, Seattle, WA 98195, USA}
+\affiliation{\dirac}
+\affiliation{Department of Physics, University of Washington, Seattle, WA 98195, USA}
 
 \author[0000-0002-6825-5283]{J.~Bryce~Kalmbach}
-\affiliation{DIRAC Institute and the Department of Astronomy, University of Washington, Seattle, WA 98195, USA}
+\affiliation{\dirac}
 
 \author[0000-0003-4906-8447]{Alexander~Gagliano}
-\affiliation{\Illinois}
-\affiliation{\NCSA}
-\affiliation{\CAPS}
-\affiliation{\NSF}
+\affiliation{Department of Astronomy, University of Illinois at Urbana-Champaign, 1002 W. Green St., IL 61801, USA}
+\affiliation{National Center for Supercomputing Applications, Urbana, IL, 61801, USA}
+\affiliation{Center for AstroPhysical Surveys, Urbana, IL, 61801, USA}
+\affiliation{National Science Foundation Graduate Research Fellow}
 
@@
 \begin{acknowledgements}
+    J.~F.~Crenshaw, B.~J.~Kalmbach, and A.~J.~Connolly acknowledge support from the DiRAC Institute in the Department of Astronomy at the University of Washington.
+    The DIRAC Institute is supported through generous gifts from the Charles and Lisa Simonyi Fund for Arts and Sciences, and the Washington Research Foundation.
+    Z.~Yan acknowledges support from the Max Planck Society and the Alexander von Humboldt Foundation in the framework of the Max Planck-Humboldt Research Award endowed by the German Federal Ministry of Education and Research.
+
     Author contributions are listed below: \\

### 2023-01-06 src/tex/ms.tex
Fixed dirac affil.
Moving figures.
Removed Francois as author

@@
 \author[0000-0001-5576-8189]{Andrew~J.~Connolly}
-\affiliation{DIRAC Institute and the Department of Astronomy, University of Washington, Seattle, WA 98195, USA}
-
-\author{François~Lanusse}
+\affiliation{\dirac}
 
 \author{Alex~I.~Malz}
+\affiliation{\note{Alex's affiliation}}
 
@@
 
-Since we have direct access to the probability distribution from which the photometry and redshifts are drawn, we can calculate the true redshift posterior for each galaxy: $p(z|m)$ where $m$ is the vector of galaxy magnitudes.
-We note that this is not an estimate, like what would be returned by a photo-z estimator, but rather the truth, obtained from the model that generated the photometry and redshifts in the first place.
-
-In addition to calculating the true posterior, $p(z|m)$, we can calculate a true posterior that is consistent with the photometric errors:
-\begin{align}
-    p(z|m, \sigma_m) = \int p(z|\hat{m}) p(\hat{m}|m, \sigma_m) d\hat{m},
-    \label{eq:err-conv}
-\end{align}
-where $\sigma_m$ is the vector of photometric errors returned by the error model, and $\hat{m}$ are possible noisy observations of the true magnitudes, $m$.
-We can evaluate this integral numerically by sampling $\hat{m} \sim p(\hat{m}|m, \sigma_m)$, evaluating $p(z, \hat{m})$ on a redshift grid, and normalizing to obtain $p(z|\hat{m})$.
-Averaging $p(z|\hat{m})$ over the samples $\hat{m}$ yields $p(z|m, \sigma_m)$.
-This is the ``best case scenario'' redshift posterior that photo-z estimators hope to estimate.
-
-We can also marginalize over any missing bands, $n$:
-\begin{align}
-    p(z|\hat{m}) = \frac{1}{p(\hat{m})} \int p(z, n, \hat{m}) dn,
-\end{align}
-which can be calculated by evaluating $p(z, n, \hat{m})$ on a grid of $z$ and $n$, summing over $n$ to yield $p(z, \hat{m})$, and normalizing with respect to redshift to yield $p(z|\hat{m})$.
-You can once again average over $\hat{m}$ samples to convolve the photometric errors.
-You may wish to marginalize over all values of $n$ if the galaxy was not observed in that band.
-This might occur, for example, if you include photometry from the Euclid Space Telescope \citep{euclid}, which will not have complete coverage of the LSST catalog.
-You may also wish to marginalize over all values beyond a limiting magnitude if the galaxy was observed, but not detected in that band.
-This might occur, for example, in the low wavelength bands of Lyman dropout galaxies observed by LSST.
-
 \begin{figure}[t]
@@
 
+Since we have direct access to the probability distribution from which the photometry and redshifts are drawn, we can calculate the true redshift posterior for each galaxy: $p(z|m)$ where $m$ is the vector of galaxy magnitudes.
+We note that this is not an estimate, like what would be returned by a photo-z estimator, but rather the truth, obtained from the model that generated the photometry and redshifts in the first place.
+
+In addition to calculating the true posterior, $p(z|m)$, we can calculate a true posterior that is consistent with the photometric errors:
+\begin{align}
+    p(z|m, \sigma_m) = \int p(z|\hat{m}) p(\hat{m}|m, \sigma_m) d\hat{m},
+    \label{eq:err-conv}
+\end{align}
+where $\sigma_m$ is the vector of photometric errors returned by the error model, and $\hat{m}$ are possible noisy observations of the true magnitudes, $m$.
+We can evaluate this integral numerically by sampling $\hat{m} \sim p(\hat{m}|m, \sigma_m)$, evaluating $p(z, \hat{m})$ on a redshift grid, and normalizing to obtain $p(z|\hat{m})$.
+Averaging $p(z|\hat{m})$ over the samples $\hat{m}$ yields $p(z|m, \sigma_m)$.
+This is the ``best case scenario'' redshift posterior that photo-z estimators hope to estimate.
+
+We can also marginalize over any missing bands, $n$:
+\begin{align}
+    p(z|\hat{m}) = \frac{1}{p(\hat{m})} \int p(z, n, \hat{m}) dn,
+\end{align}
+which can be calculated by evaluating $p(z, n, \hat{m})$ on a grid of $z$ and $n$, summing over $n$ to yield $p(z, \hat{m})$, and normalizing with respect to redshift to yield $p(z|\hat{m})$.
+You can once again average over $\hat{m}$ samples to convolve the photometric errors.
+You may wish to marginalize over all values of $n$ if the galaxy was not observed in that band.
+This might occur, for example, if you include photometry from the Euclid Space Telescope \citep{euclid}, which will not have complete coverage of the LSST catalog.
+You may also wish to marginalize over all values beyond a limiting magnitude if the galaxy was observed, but not detected in that band.
+This might occur, for example, in the low wavelength bands of Lyman dropout galaxies observed by LSST.
+
 To visualize the impacts of error convolution and band marginalization, Figure \ref{fig:posteriors} shows a number of redshift posteriors for a single example galaxy.
@@
 \label{sec:fwd-model-conditional}
+
+% this figure is so high just to move it ahead one page in the document
+\begin{figure*}[t]
+    \script{photo-z/plot_ensemble_posteriors.py}
+    \begin{centering}
+        \includegraphics{figures/ensemble_posteriors.pdf}
+        \caption{
+            The ensemble of posteriors for three example galaxies.
+            Flows 1-4 label the individual posteriors produced by each of the flows that make up the ensemble.
+            The dashed black line is the mean of these individual posteriors and is the value used by the ensemble.
+            The vertical gray line labeled ``Truth'' denotes the true redshift of the galaxy.
+            Note these galaxies were specifically chosen for their broad, multimodal posteriors.
+            The posteriors of most galaxies are sharp and unimodal.
+        }
+        \label{fig:ensemble-posteriors}
+    \end{centering}
+\end{figure*}
 
@@
 \label{sec:photo-z}
-
-\begin{figure*}[t]
-    \script{photo-z/plot_ensemble_posteriors.py}
-    \begin{centering}
-        \includegraphics{figures/ensemble_posteriors.pdf}
-        \caption{
-            The ensemble of posteriors for three example galaxies.
-            Flows 1-4 label the individual posteriors produced by each of the flows that make up the ensemble.
-            The dashed black line is the mean of these individual posteriors and is the value used by the ensemble.
-            The vertical gray line labeled ``Truth'' denotes the true redshift of the galaxy.
-            Note these galaxies were specifically chosen for their broad, multimodal posteriors.
-            The posteriors of most galaxies are sharp and unimodal.
-        }
-        \label{fig:ensemble-posteriors}
-    \end{centering}
-\end{figure*}
 
@@
 
+\begin{figure}[t]
+    \script{photo-z/plot_marginalized_posterior.py}
+    \begin{centering}
+        \includegraphics{figures/posterior_marginalized.pdf}
+        \caption{
+            Redshift posteriors for an example galaxy.
+            The vertical red line indicates the true redshift.
+            The black posterior is estimated using the information from all 6 LSST bands (and includes convolution of the photometric errors).
+            The orange posterior has had the $u$ band marginalized, which opens up a second high-redshift mode.
+        }
+        \label{fig:posterior_marginalized}
+    \end{centering}
+\end{figure}
+
 For each flow in the ensemble, we estimate $p(z, m)$ by marginalizing over the photometric errors:
@@
 Conversely, this demonstrates how the additon of $u$ band information rules out the higher redshift solution for this galaxy.
-
-\begin{figure}[t]
-    \script{photo-z/plot_marginalized_posterior.py}
-    \begin{centering}
-        \includegraphics{figures/posterior_marginalized.pdf}
-        \caption{
-            Redshift posteriors for an example galaxy.
-            The vertical red line indicates the true redshift.
-            The black posterior is estimated using the information from all 6 LSST bands (and includes convolution of the photometric errors).
-            The orange posterior has had the $u$ band marginalized, which opens up a second high-redshift mode.
-        }
-        \label{fig:posterior_marginalized}
-    \end{centering}
-\end{figure}
 
@@
 \begin{acknowledgements}
+    Thanks to François~Lanusse for some early advice on normalizing flows and for reviewing the paper.
     J.~F.~Crenshaw, B.~J.~Kalmbach, and A.~J.~Connolly acknowledge support from the DiRAC Institute in the Department of Astronomy at the University of Washington.
@@
     A.~J.~Connolly: discussion during development. \\
-    F.~Lanusse: wrote early normalizing flow code that evolved into PZFlow. \\
     A.~I.~Malz: consulted on design and testing of PZFlow. \\
@@
 \label{app:training-details}
-
-In this section we list some technical details of training the normalizing flows.
-Every flow is trained via minimizing the negative log-likelihood
-\begin{align}
-    \mathcal{L} = - \, \mathbb{E}[ \, \log p(x) \, ],
-\end{align}
-where the expectation is performed over galaxies in the training set and $p(x)$ is defined in Equation \ref{eq:px}.
-
-For the main flow in Section \ref{sec:galaxy-catalog}, we trained for 150 epochs.
-We used the Adam optimizer \citep{adam}, starting with a learning rate of $10^{-3}$.
-We decreased the learning rate by a factor of 10 every 50 epochs.
-Training took 7 minutes on a Tesla P100 12GB GPU.
-The training loss for this flow is in the left panel of Figure \ref{fig:galaxy-flow-losses}.
-
-For the conditional flow in \ref{sec:galaxy-catalog}, we trained for 450 epochs.
-Again, we used Adam with an initial learning rate of $10^{-3}$.
-We decreased the learning rate by a factor of 10 every 150 epochs.
-The training loss for this flow is in the right panel of Figure \ref{fig:galaxy-flow-losses}.
-
-For each of the flows that make up the flow ensemble in Section \ref{sec:photo-z}, we trained for 150 epochs using the Adam optimizer.
-We started each with a learning rate of $10^{-4}$, and decreased the learning rate by a factor of 10 every 50 epochs.
-The training loss for the ensemble is in Figure \ref{fig:ensemble-losses}.
-You can see that each of the flows achieves a different minimum loss.
-Apparently, each has found a different potential solution in the neural network's parameter space.
 
@@
 \end{figure}
+
+In this section we list some technical details of training the normalizing flows.
+Every flow is trained via minimizing the negative log-likelihood
+\begin{align}
+    \mathcal{L} = - \, \mathbb{E}[ \, \log p(x) \, ],
+\end{align}
+where the expectation is performed over galaxies in the training set and $p(x)$ is defined in Equation \ref{eq:px}.
+
+For the main flow in Section \ref{sec:galaxy-catalog}, we trained for 150 epochs.
+We used the Adam optimizer \citep{adam}, starting with a learning rate of $10^{-3}$.
+We decreased the learning rate by a factor of 10 every 50 epochs.
+Training took 7 minutes on a Tesla P100 12GB GPU.
+The training loss for this flow is in the left panel of Figure \ref{fig:galaxy-flow-losses}.
+
+For the conditional flow in \ref{sec:galaxy-catalog}, we trained for 450 epochs.
+Again, we used Adam with an initial learning rate of $10^{-3}$.
+We decreased the learning rate by a factor of 10 every 150 epochs.
+The training loss for this flow is in the right panel of Figure \ref{fig:galaxy-flow-losses}.
+
+For each of the flows that make up the flow ensemble in Section \ref{sec:photo-z}, we trained for 150 epochs using the Adam optimizer.
+We started each with a learning rate of $10^{-4}$, and decreased the learning rate by a factor of 10 every 50 epochs.
+The training loss for the ensemble is in Figure \ref{fig:ensemble-losses}.
+You can see that each of the flows achieves a different minimum loss.
+Apparently, each has found a different potential solution in the neural network's parameter space.
 

### 2023-04-12 src/tex/ms.tex
Ack Martine

@@
 
-\author{Alex~I.~Malz}
-\affiliation{\note{Alex's affiliation}}
+\author[0000-0002-8676-1622]{Alex~I.~Malz}
+\affiliation{McWilliams Center for Cosmology, Department of Physics, Carnegie Mellon University}
 
@@
 
-Astronomical data represent realizations of complex probability distributions.
+Testing some new text here.
+
+We deleted a sentence here.
 A common goal in research is to infer the underlying distribution from a limited set of noisy data, and use this distribution to estimate posterior distributions over galaxy and population parameters.
-An important example is photometric redshift (photo-z) estimation, where galaxy redshift posteriors are estimated from galaxy photometry, using a model informed by a training set of spectroscopic galaxy data \citep{newman2022}.
+An important example AND ARE ADDING some new text HERE in the middle of this sentence is photometric redshift (photo-z) estimation, where galaxy redshift posteriors are estimated from galaxy photometry, using a model informed by a training set of spectroscopic galaxy data \citep{newman2022}.
 These redshift posteriors are then used to estimate posterior distributions for cosmological parameters \citep{descSRD}.
@@
 \begin{acknowledgements}
-    Thanks to François~Lanusse for some early advice on normalizing flows and for reviewing the paper.
+    Thanks to François~Lanusse for some early advice on normalizing flows and for reviewing the paper, and to Martine Lokken for testing PZFlow.
     J.~F.~Crenshaw, B.~J.~Kalmbach, and A.~J.~Connolly acknowledge support from the DiRAC Institute in the Department of Astronomy at the University of Washington.
@@
     Z.~Yan acknowledges support from the Max Planck Society and the Alexander von Humboldt Foundation in the framework of the Max Planck-Humboldt Research Award endowed by the German Federal Ministry of Education and Research.
+    AIM acknowledges support during the course of this work from the Max Planck Society and the Alexander von Humboldt Foundation in the framework of the Max Planck-Humboldt Research Award endowed by the Federal Ministry of Education and Research.
 

### 2023-04-22 src/tex/ms.tex
Some text revisions.

@@
 % Define document class
-\documentclass[twocolumn,twocolappendix]{aastex631}
+\documentclass[twocolumn,twocolappendix,linenumbers]{aastex631}
 \usepackage{showyourwork}
@@
     We introduce PZFlow, a Python package for statistical modeling of tabular data using normalizing flows.
-    We use PZFlow to model photometric galaxy catalogs including redshifts, LSST photometry, size, and ellipticity, and generate a synthetic catalog.
+    We use PZFlow to model photometric galaxy catalogs including redshifts, photometry, size, and shape, and generate a synthetic catalog.
     In this catalog, each galaxy has a \emph{true} redshift posterior.
@@
 
-Testing some new text here.
-
-We deleted a sentence here.
+Astronomical data represent realizations of complex probability distributions.
 A common goal in research is to infer the underlying distribution from a limited set of noisy data, and use this distribution to estimate posterior distributions over galaxy and population parameters.
-An important example AND ARE ADDING some new text HERE in the middle of this sentence is photometric redshift (photo-z) estimation, where galaxy redshift posteriors are estimated from galaxy photometry, using a model informed by a training set of spectroscopic galaxy data \citep{newman2022}.
+An important example is photometric redshift (photo-z) estimation, where galaxy redshift posteriors are estimated from galaxy photometry, using a model informed by a training set of spectroscopic galaxy data \citep{newman2022}.
 These redshift posteriors are then used to estimate posterior distributions for cosmological parameters \citep{descSRD}.
@@
 
-Inferring the underlying probability distribution from a high-dimensional data set is a difficult problem, and many machine learning tools have been developed for this task.
-Two popular deep learning examples are Generative Adversarial Networks (GANs; \citealt{goodfellow2014}) and Variational Autoencoders (VAEs; \citealt{kingma2014}).
-Both methods excel at forward modeling complex data sets, but neither allow exact likelihood inference.
+Inferring the underlying probability distribution, or likelihood, from a high-dimensional data set is a difficult problem.
+Many machine learning tools have been developed for this task, including Generative Adversarial Networks (GANs; \citealt{goodfellow2014}) and Variational Autoencoders (VAEs; \citealt{kingma2014}).
+Both are neural networks that excel at forward modeling complex data sets, but neither allow exact likelihood calculation for the simulated data.
+This means that questions such as ``under my model, what is the posterior distribution for redshift given the simulated photometry'' have only approximate answers.
 
-Normalizing flows, on the other hand, are a deep learning tool that excel at forward modeling, while also allowing exact likelihood inference.
-They operate by learning an invertible transformation of the data distribution into a simpler, tractable latent distribution (such as a normal distribution, hence the name \emph{normalizing} flow).
-This allows sampling and likelihood inference to be performed within the context of the latent distribution, with the normalizing flow acting as a translator between the latent samples and likelihoods, and their corresponding values in the context of the data distribution.
+Normalizing flows, on the other hand, are a deep learning tool that excel at forward modeling, while also allowing exact, analytic likelihood calculation.
+In other words, they allow you to analytically calculate likelihoods and posteriors with respect to the distribution from which simulated data is drawn.
+Normalizing flows operate by learning an invertible transformation of the data distribution into a simpler, tractable distribution, known as the latent distribution.
+A common choice for latent distribution is a normal distribution, hence the name \emph{normalizing} flow.
+This allows sampling and likelihood calculation to be performed within the context of the simple latent distribution, with the normalizing flow acting as a translator between the latent samples and likelihoods, and their corresponding values in the context of the complicated data distribution.
 
 Normalizing flows have gained popularity as tools for efficient and flexible sampling for parameter inference (e.g., \citealt{dai2022,dacunha2022,hassan2022}).
-In this paper, we focus on photo-z's and using normalizing flows to generate synthetic catalogs.
-Since normalizing flows provide exact likelihood inference, the properties of objects in these catalogs have \emph{true} posteriors.
-When calculating these posteriors, you can convolve error distributions, and marginalize over chosen properties.
-Catalogs with \emph{true} posteriors are useful for the testing and validation of analysis pipelines that estimate posteriors, such as the photo-z estimators used in much of astrophysics and cosmology.
+In this paper, we focus on photo-z's and forward modeling photometric galaxy catalogs.
+Since normalizing flows allow exact, analytic probability calculation, the properties of objects in these catalogs have \emph{true} posteriors.
+This means that questions such as ``under my model, what is the posterior distribution for redshift given the simulated photometry'' have exact answers.
+Catalogs with true posteriors are useful for the testing and validation of analysis pipelines that estimate posteriors, such as the photo-z estimators used in much of astrophysics and cosmology.
 Previous evaluations of these estimators have focused on comparing point estimates to true values (e.g., \citealt{hildebrandt2010,sanchez2014,graham2018}), or evaluating ensembles of posteriors \citePZp.
-Normalizing flow catalogs with \emph{true} posteriors open up a new, more comprehensive avenue for evaluation of these estimators by enabling posterior-to-posterior comparison.
+Normalizing flow catalogs with true posteriors open up a new, more comprehensive avenue for evaluation of these estimators by enabling direct posterior-to-posterior comparison.
 
@@
 
-If we further assume that $f$ is invertible, we can sample from the data distribution by appling $f^{-1}$ to samples from the latent distribution:
+If we further assume that $f$ is invertible, we can sample from the data distribution by applying $f^{-1}$ to samples from the latent distribution\footnote{Here, $~$ means ``is drawn from.''}:
 \begin{align}
@@
 
+A bijection is an invertible map between two sets.
+In a normalizing flow, the bijection maps the data distribution onto the latent distribution for likelihood calculation, and the inverse of the bijection maps samples from the latent distribution back to the data distribution.
 The bijection of a normalizing flow must be powerful enough to model complex relationships in data, while remaining invertible and simultaneously possessing an efficiently computable Jacobian determinant.
@@
 Some bijections are specialized to be particularly efficient at either density estimation or sampling, but for many science cases, we wish to do both.
-For this reason, we will focus on Rational-Quadratic Rolling Spline Couplings (RQ-RSCs), which achieve state-of-the-art performance, while being efficient with both tasks \citep{durkan2019}.
+For this reason, we will focus on Rational-Quadratic Rolling Spline Couplings (RQ-RSCs), bijections which achieve state-of-the-art performance, while being efficient with both tasks \citep{durkan2019}.
 
@@
         \draw[->] (2) -- (3);
-
         % red boxes
@@
         \draw[->] (5) -- (6);
-
         % coupling function
         \draw[->] (1) -- (5) node [midway, draw, circle, fill=white, minimum size=8mm, fill=gray!20] (7) {$m$};
-
     \end{tikzpicture}
@@
 
-RQ-RSCs are based on coupling layers \citep{dinh2015, dinh2017}.
+RQ-RSCs are bijections that are composed of coupling layers \citep{dinh2015, dinh2017}.
 A coupling layer partitions the data, $x \in \R^D$, into two sets, $x_{1:d}$ and $x_{d+1:D}$.
@@
 \end{align}
-and in particular, the Jacobian determinant is
+where $I_d$ is the $d \times d$ identity matrix.
+In particular, the Jacobian determinant is
 \begin{align}
@@
 \end{align}
-Notice that neither inverting a coupling layer nor calculating the Jacobian determinant requires inverting or taking derivatives of the coupling function $m$, which can thus be arbitrarily complex.
+Notice that neither inverting a coupling layer $g$, nor calculating the Jacobian determinant, requires inverting or taking derivatives of the coupling function $m$, which can thus be arbitrarily complex.
 
@@
 
-There are a variety of different coupling laws one can use.
-One particularly expressive choice is a Rational-Quadratic Neural Spline Coupling (RQ-NSC; \citealt{durkan2019}).
-As the name suggests, the \emph{coupling law} for RQ-NSCs is a set of rational-quadratic splines and the \emph{coupling function} is a neural network.
-RQ-NSCs define a spline $g_i: [-B, B] \subset \R \to [-B, B]$ for each dimension $i$ of $x_{d+1:D}$, where $g_i$ is a piecewise combination of $K$ segments, and each segment is a rational-quadratic function.
-The positions and derivatives of the knots that parameterize the splines are calculated from $x_{1:d}$ by a neural network.
+In a general coupling layer $g$, there are a variety of coupling laws $m$ one can use.
+RQ-RSC's use Rational-Quadratic Neural Spline Coupling \citep{durkan2019}.
+As the name  suggests, the coupling law $g$ is a set of rational-quadratic splines.
+In particular, $g_i: [-B, B] \to [-B, B]$ for each dimension $i$ of $x_{d+1:D}$, where $g_i$ is a piecewise combination of $K$ segments, and each segment is a rational-quadratic function.
+The positions and derivatives of the knots that parameterize the splines are calculated using the coupling function $m$, which is a dense neural network applied to $x_{1:d}$.
 
-RQ-NSCs are state-of-the-art for efficient sampling and density estimation \citep{kobyzev2020}, and are flexible enough to model complex distributions with multiple discontinuities and hundreds of modes.
+The result is a bijection that achieves state-of-the-art performance and efficiency for forward modeling and density estimation \citep{kobyzev2020}, and are flexible enough to model complex distributions with multiple discontinuities and hundreds of modes.
 In addition, they are easily adaptable for flows with periodic topology (Section \ref{sec:periodic}).
@@
 
-In this work, we introduce a simple extension of RQ-NSCs, which we name Rational-Quadratic Rolling Spline Couplings (RQ-RSCs).
-An RQ-RSC models $D$ dimensional data by stacking $D$ RQ-NSC layers (with $d=D-1$), each followed by a Rolling layer.
-A Rolling layer shifts the dimensions of $x$ by one place:
+In this work, we stack Rational-Quadratic Neural Spline Couplings, with Rolling Layers between each -- a configuration we name Rational-Quadratic Rolling Spline Couplings (RQ-RSCs).
+Rolling Layers shift the dimensions of $x$ by one place:
 \begin{align}
@@
 \end{align}
-In other words, RQ-RSCs individually transform each of the $D$ dimensions of $x$ as a function of the other $D-1$ dimensions.
+By constructing a stack with $D$ coupling layers, RQ-RSCs individually transform each of the $D$ dimensions of $x$ as a function of the other $D-1$ dimensions.
 In the limit of high spline resolution (i.e.
 $K \to \infty$), RQ-RSCs can model any differentiable, monotonic function on $[-B, B]^D$ and can thus model arbitrarily complex distributions in this region.
-In practice, we find very good performance for diverse data sets with $K=16$.
+In practice, we find very good performance for diverse data sets with $K \approx 16$.
 
 Note you can specify a different value of $K$ for each of the $D$ spline layers in order to individually control the resolution of each dimension.
+Lowering $K$ typically results in a smoother distribution, while increasing $K$ increases the complexity the normalizing flow can capture, while also increasing computational and memory cost.
 
@@
         \draw[->] (5) -- (6);
-
         % coupling function
         \draw[->] (1) -- (5) node [midway, draw, circle, fill=white, minimum size=8mm, fill=gray!20] (7) {$m$};
-
         % conditional variables
@@
         \draw[-] (8.south east) -- (7.north west);
-
     \end{tikzpicture}
@@
 Positions on the sphere are specified by two angles\footnote{
-We use the physicists' convention where $\theta$ and $\phi$ are the zenith and azimuthal angles, respectively.
+We use the convention where $\theta$ and $\phi$ are the zenith and azimuthal angles, respectively.
 },
@@
 The only constraint we must impose is that samples of $\cos\theta$ must lie in the range $[-1, 1]$.
-Fortunately, RQ-NSCs are bounded, mapping a range in $u$ to the same range in $x$.
-Thus, if we pick a latent distribution with compact support in $[-1, 1]$, samples of $\cos\theta$ are guaranteed to lie in the same range, as long as we set the range of the RQ-NSC $B$ = 1.
+Fortunately, RQ-RSCs are bounded, mapping a range in $u$ to the same range in $x$.
+Thus, if we pick a latent distribution with compact support in $[-1, 1]$, samples of $\cos\theta$ are guaranteed to lie in the same range, as long as we set the range of the RQ-RSC $B$ = 1.
 
 Next we will address transformations of $\phi$.
-For $f$ to be a valid diffeomorphism on the circle, $S^1$, it is sufficient that $f$ obey the following constraints:
+For $f$ to be a differentiable bijection on the circle, $S^1$, it is sufficient that $f$ obey the following constraints:
 \begin{align}
@@
 The first two constraints ensure continuity of $f$ by designating $\phi=0$ as a fixed point, and the third constraint ensures continuity of $\nabla f$ at that fixed point.
-While the designation of $\phi=0$ as a fixed point is an unnecessary restriction on $f$, any diffeomorphism on the circle has at least one fixed point up to a phase change, and so this restriction does not actually restrict the expressiveness of $f$.
+While the designation of $\phi=0$ as a fixed point is an unnecessary restriction on $f$, any differentiable bijection on the circle has at least one fixed point up to a phase change, and so this restriction does not actually restrict the expressiveness of $f$.
 The fourth restriction ensures monotonicity, which guarantees invertibility.
@@
         \caption{
-            Distribution of the CosmoDC2 test set compared to the distribution learned by PZFlow.
+            Distribution of true redshifts and noiseless photometry from the CosmoDC2 test set, compared to a sample drawn from the distribution learned by PZFlow.
             The close overlap of every pair-wise distribution demonstrates that PZFlow has learned the distribution in CosmoDC2 with high fidelity.
@@
         \caption{
-            Comparing the $r-i$ vs redshift distribution for galaxy samples from CosmoDC2 (left) and from the normalizing flow (right).
+            Comparing the noiseless $r-i$ vs redshift distribution for galaxy samples from CosmoDC2 (left) and from the normalizing flow (right).
             The high-redshift galaxies in CosmoDC2 lie along discrete tracks in color space due to the discrete number of galaxy SED templates used in the simulation.
@@
 To create a generative model of galaxy redshifts and photometry, we use the true redshifts and $ugrizy$ magnitudes from the CosmoDC2 simulation \citep{dc2, cosmodc2} of the LSST Dark Energy Science Collaboration (DESC).
-We selected all galaxies from CosmoDC2 with at most one band beyond the LSST 10-year 5-sigma point source depths.
+We selected all galaxies from CosmoDC2 with at most one band with a signal-to-noise ratio (SNR) less than 10, using the forecasted photometric errors from LSST year 10.
 Of these, we randomly selected $10^6$ galaxies and split them intro training and test sets consisting of 80\% and 20\% of the galaxies, respectively.
@@
 
-Since we have direct access to the probability distribution from which the photometry and redshifts are drawn, we can calculate the true redshift posterior for each galaxy: $p(z|m)$ where $m$ is the vector of galaxy magnitudes.
+Since we have direct access to the probability distribution from which the photometry and redshifts are drawn, we can analytically calculate the true redshift posterior for each galaxy: $p(z|m)$ where $m$ is the vector of galaxy magnitudes.
 We note that this is not an estimate, like what would be returned by a photo-z estimator, but rather the truth, obtained from the model that generated the photometry and redshifts in the first place.
@@
 When forward modeling in Section \ref{sec:galaxy-catalog}, we wanted a realistic model that captured the relevant correlations between galaxy photometry, redshift, shape, and size.
-However, when estimating redshifts, we do not simply want a realistic model, but rather a model that matches our galaxy sample as closely as possible.
+However, when estimating redshifts, we do not simply want a realistic model, but rather a model that matches our specific galaxy sample as closely as possible.
 
@@
 \subsection{Estimating posteriors}
-
-\begin{figure}[t]
-    \script{photo-z/plot_marginalized_posterior.py}
-    \begin{centering}
-        \includegraphics{figures/posterior_marginalized.pdf}
-        \caption{
-            Redshift posteriors for an example galaxy.
-            The vertical red line indicates the true redshift.
-            The black posterior is estimated using the information from all 6 LSST bands (and includes convolution of the photometric errors).
-            The orange posterior has had the $u$ band marginalized, which opens up a second high-redshift mode.
-        }
-        \label{fig:posterior_marginalized}
-    \end{centering}
-\end{figure}
 

### 2023-04-22 src/scripts/forward_model/train_conditional_galaxy_flow.py

@@ -44,13 +44,13 @@ flow = Flow(
 
 # train for three rounds of 150 epochs
 # after each round, decrease the learning rate by a factor of 10
-opt = optax.adam(1e-3)
+opt = optax.adam(1e-5)
 losses = flow.train(train_set, epochs=150, optimizer=opt, seed=0, verbose=True)
 
-opt = optax.adam(1e-4)
+opt = optax.adam(1e-6)
 losses += flow.train(train_set, epochs=150, optimizer=opt, seed=1, verbose=True)
 
-opt = optax.adam(1e-5)
+opt = optax.adam(1e-7)
 losses += flow.train(train_set, epochs=150, optimizer=opt, seed=2, verbose=True)
 
 # save some info with the model

### 2023-04-22 src/scripts/forward_model/train_main_galaxy_flow.py

@@ -55,13 +55,13 @@ flow = Flow(train_set.columns, bijector=bijector)
 
 # train for three rounds of 50 epochs
 # after each round, decrease the learning rate by a factor of 10
-opt = optax.adam(1e-3)
+opt = optax.adam(1e-5)
 losses = flow.train(data, epochs=50, optimizer=opt, seed=0, verbose=True)
 
-opt = optax.adam(1e-4)
+opt = optax.adam(1e-6)
 losses += flow.train(data, epochs=50, optimizer=opt, seed=1, verbose=True)
 
-opt = optax.adam(1e-5)
+opt = optax.adam(1e-7)
 losses += flow.train(data, epochs=50, optimizer=opt, seed=2, verbose=True)
 
 # save some info with the model

### 2023-04-22 src/scripts/photo-z/train_pz_ensemble.py

@@ -112,17 +112,17 @@ flowEns = FlowEnsemble(
 
 # we will also convolve the photometric errors during training
 
-opt = optax.adam(1e-4)
+opt = optax.adam(1e-6)
 losses1 = flowEns.train(
     train_set, convolve_errs=True, epochs=50, optimizer=opt, seed=0, verbose=True
 )
 
-opt = optax.adam(1e-5)
+opt = optax.adam(1e-7)
 losses2 = flowEns.train(
     train_set, convolve_errs=True, epochs=50, optimizer=opt, seed=1, verbose=True
 )
 
-opt = optax.adam(1e-6)
+opt = optax.adam(1e-8)
 losses3 = flowEns.train(
     train_set, convolve_errs=True, epochs=50, optimizer=opt, seed=2, verbose=True
 )

### 2023-11-22 src/tex/ms.tex
Lots of manuscript changes.

@@
 % Title
-\title{Statistical Modeling of Galaxy Catalogues with Normalizing Flows}
+\title{Probabilistic Forward Modeling of Galaxy Catalogs with Normalizing Flows}
 
@@
 \begin{abstract}
-    Normalizing flows are powerful tools for learning high-dimensional probability distributions from samples thereof, and have many applications in astronomy, including posterior estimation and forward modeling.
-    We introduce PZFlow, a Python package for statistical modeling of tabular data using normalizing flows.
-    We use PZFlow to model photometric galaxy catalogs including redshifts, photometry, size, and shape, and generate a synthetic catalog.
-    In this catalog, each galaxy has a \emph{true} redshift posterior.
-    We discuss the importance of these true posteriors for the comprehensive evaluation of redshift posteriors produced by photometric redshift (photo-z) estimators.
-    We also demonstrate the use of an ensemble of normalizing flows for density estimation, applied to photo-z estimation.
-    While we focus on photo-z estimation and validation, we emphasize that these methods are applicable to any galaxy properties, and any other tabular data.
+    Evaluating the accuracy and calibration of the redshift posteriors produced by photometric redshift (photo-z) estimators is vital for enabling precision cosmology and extragalactic astrophysics with modern wide-field photometric surveys.
+    Evaluating photo-z posteriors on a per-galaxy basis is difficult, however, as galaxies have a true redshift, but not a true redshift posterior.
+    We introduce PZFlow, a Python package for the probabilistic forward modeling of galaxy catalogs with normalizing flows.
+    For catalogs simulated with PZFlow, there is a natural notion of ``true'' redshift posteriors that can be used for photo-z validation.
+    We use PZFlow to simulate a photometric galaxy catalog where each galaxy has a redshift, noisy photometry, shape information, and a true redshift posterior.
+    We also demonstrate the use of an ensemble of normalizing flows for photo-z estimation.
+    We discuss how PZFlow will be used to validate the photo-z estimation pipeline of the Dark Energy Science Collaboration (DESC), and the wider applicability of PZFlow for statistical modeling of any tabular data.
 \end{abstract}
@@
 
-Astronomical data represent realizations of complex probability distributions.
-A common goal in research is to infer the underlying distribution from a limited set of noisy data, and use this distribution to estimate posterior distributions over galaxy and population parameters.
-An important example is photometric redshift (photo-z) estimation, where galaxy redshift posteriors are estimated from galaxy photometry, using a model informed by a training set of spectroscopic galaxy data \citep{newman2022}.
-These redshift posteriors are then used to estimate posterior distributions for cosmological parameters \citep{descSRD}.
+Photometric redshift (photo-z) estimation is necessary for the study of cosmology and galaxy evolution with the huge galaxy catalogs generated by modern wide-field photometric surveys like the Vera C.
+Rubin Observatory's Legacy Survey of Space and Time (LSST; \citealt{ivezic2019}).
+The colors of galaxies in these surveys are used to estimate posterior distributions over the possible redshift of each galaxy.
+These posteriors are often multimodal, reflecting degeneracies in redshift and galaxy spectral type.
+Guaranteeing the accuracy and calibration of these posteriors, and understanding their biases, is vital for enabling precision cosmology and studies of galaxy evolution \citep{newman2022,descSRD}.
 
-Knowledge of the probability distributions underlying our data is also valuable for simulation and forward modeling astronomical data sets.
-Simulating data sets with realistic statistical properties enables methodological development and the calibration of systematic uncertainties.
-Forward modeling also facilitates data augmentation (e.g.
-\citealt{lokken2022}) and the creation of large data challenges (e.g.
-\citealt{kessler2019, dc2, cosmodc2}), which are becoming more prevalent in the big data era of astronomy.
+There have been many studies evaluating the performance of photo-z estimation by comparing estimates to known true redshifts (e.g., \citealt{hildebrandt2010}, \citealt{sanchez2014}, \citealt{graham2018}), however to meet the needs of modern surveys, we must evaluate the accuracy of the full redshift posteriors generated for these galaxies.
+In other words, not only must we verify that photo-z estimators assign a high posterior probability to the true redshift, but also that the posteriors account for all the degeneracies inherent in photo-z estimation, and that the distributions are neither too narrow nor too broad.
+This kind of validation is difficult, however, because individual galaxies do not have a ``true'' redshift posterior to which you can compare the estimated posterior.
+\citePZt studied whether these questions could be answered using ensemble-level comparisons of photo-z posteriors and true redshifts, but found that these metrics could be fooled by a pathological photo-z estimator that ignored galaxy colors altogether.
+This raised the need for new methodologies that enable the evaluation of photo-z posterior estimates at the per-galaxy level.
 
-Inferring the underlying probability distribution, or likelihood, from a high-dimensional data set is a difficult problem.
-Many machine learning tools have been developed for this task, including Generative Adversarial Networks (GANs; \citealt{goodfellow2014}) and Variational Autoencoders (VAEs; \citealt{kingma2014}).
-Both are neural networks that excel at forward modeling complex data sets, but neither allow exact likelihood calculation for the simulated data.
-This means that questions such as ``under my model, what is the posterior distribution for redshift given the simulated photometry'' have only approximate answers.
+One way to enable posterior validation at the per-galaxy level is to simulate catalogs by sampling galaxies from a probabilistic model for which you have direct access to the underlying probability distribution.
+Then, for each galaxy, you can calculate the posterior over redshift with respect to the same model that generated the galaxy.
+This provides a ``true'' redshift posterior for each simulated galaxy, to which you can compare the posteriors from photo-z estimators.
 
-Normalizing flows, on the other hand, are a deep learning tool that excel at forward modeling, while also allowing exact, analytic likelihood calculation.
-In other words, they allow you to analytically calculate likelihoods and posteriors with respect to the distribution from which simulated data is drawn.
-Normalizing flows operate by learning an invertible transformation of the data distribution into a simpler, tractable distribution, known as the latent distribution.
-A common choice for latent distribution is a normal distribution, hence the name \emph{normalizing} flow.
-This allows sampling and likelihood calculation to be performed within the context of the simple latent distribution, with the normalizing flow acting as a translator between the latent samples and likelihoods, and their corresponding values in the context of the complicated data distribution.
+Normalizing flows are a generative model of this variety.
+Normalizing flows model complex, high-dimensional probability distributions using deep neural networks that learn an invertible mapping between the complicated data distribution and a more simple distribution, known as the latent distribution.
+This allows you to sample from and calculate probabilities with respect to the latent distribution, and use the normalizing flow to translate these values back to the space of the original data distribution.
+A common choice for the latent distribution is the Normal distribution, hence the name \emph{normalizing} flow.
+Unlike other deep generative models like Generative Adversarial Networks (GANs; \citealt{goodfellow2014}) and Variational Autoencoders (VAEs; \citealt{kingma2014}), normalizing flows provide a deterministic mapping between the data and latent spaces.
+It is this feature of normalizing flows that enables us to provide an exact answer to the question ``under the generative model, what is the posterior distribution for redshift given the simulated photometry''.
+This allows us to define a ``true'' redshift posterior for each galaxy in our simulated catalog, making normalizing flows a very powerful tool for validating photo-z inference.
 
-Normalizing flows have gained popularity as tools for efficient and flexible sampling for parameter inference (e.g., \citealt{dai2022,dacunha2022,hassan2022}).
-In this paper, we focus on photo-z's and forward modeling photometric galaxy catalogs.
-Since normalizing flows allow exact, analytic probability calculation, the properties of objects in these catalogs have \emph{true} posteriors.
-This means that questions such as ``under my model, what is the posterior distribution for redshift given the simulated photometry'' have exact answers.
-Catalogs with true posteriors are useful for the testing and validation of analysis pipelines that estimate posteriors, such as the photo-z estimators used in much of astrophysics and cosmology.
-Previous evaluations of these estimators have focused on comparing point estimates to true values (e.g., \citealt{hildebrandt2010,sanchez2014,graham2018}), or evaluating ensembles of posteriors \citePZp.
-Normalizing flow catalogs with true posteriors open up a new, more comprehensive avenue for evaluation of these estimators by enabling direct posterior-to-posterior comparison.
+In this paper, we introduce PZFlow, a normalizing flow package for Python that is designed to facilitate forward modeling galaxy catalogs with true posteriors for galaxy properties.
+In addition, we demonstrate PZFlow as a photo-z estimator.
+With relatively little tuning required by the user, PZFlow can provide a generative model for any tabular data, including continuous and discrete variables, and variables with Euclidean or periodic topology (e.g.
+the celestial sphere).
+While calculating posteriors, PZFlow can convolve error distributions and marginalize over missing values.
 
-To facilitate the statistical modeling of galaxy catalogs and other astronomical data sets, we have developed PZFlow, a normalizing flow package for Python.
-With relatively little tuning required by the user, PZFlow can provide a generative model for any tabular data, including continuous and discrete variables, and variables with Euclidean or periodic topology.
-In addition to generative modeling, PZFlow can calculate posteriors over any columns in your data set, and can convolve errors and marginalize over missing columns while training the model or calculating posteriors for samples.
-
-In this paper, we provide the background on normalizing flows (Section \ref{sec:nf}) required to understand PZFlow (Section \ref{sec:pzflow}).
-We then use PZFlow to simulate a galaxy catalog (Section \ref{sec:galaxy-catalog}), where each object has photometry, size, ellipticity, redshift, and a true redshift posterior.
-We also demonstrate using PZFlow as a density estimator, via the example of photo-z estimation (Section \ref{sec:photo-z}).
+In Section \ref{sec:nf} we provide the background on normalizing flows required to understand PZFlow, which we describe in Section \ref{sec:pzflow}.
+In Section \ref{sec:fwd-model}, we demonstrate using PZFlow to simulate a galaxy catalog where each object has a redshift, photometry in the six LSST bands, a true photo-z posterior, a size, and an ellipticity.
+In Section \ref{sec:photo-z}, we demonstrate using PZFlow as a density estimator by estimating photo-z's for our simulated catalog.
 We conclude in Section \ref{sec:conclusion}.
@@
 
-If we further assume that $f$ is invertible, we can sample from the data distribution by applying $f^{-1}$ to samples from the latent distribution\footnote{Here, $~$ means ``is drawn from.''}:
+If we further assume that $f$ is invertible, we can sample from the data distribution by applying $f^{-1}$ to samples from the latent distribution\footnote{Here, $\sim$ means ``is drawn from.''}:
 \begin{align}
@@
 A bijection is an invertible map between two sets.
-In a normalizing flow, the bijection maps the data distribution onto the latent distribution for likelihood calculation, and the inverse of the bijection maps samples from the latent distribution back to the data distribution.
+In a normalizing flow, the bijection maps the data distribution onto the latent distribution for probability calculation, and the inverse of the bijection maps samples from the latent distribution back to the data distribution.
 The bijection of a normalizing flow must be powerful enough to model complex relationships in data, while remaining invertible and simultaneously possessing an efficiently computable Jacobian determinant.
@@
 We give this generalization the name \emph{latent-adaptive flows} (LAFs; inspired by the Tail Adaptive Flows of \citealt{jaini2020}).
-Our experiments indicate that learnable latent distributions can improve training loss, but require more care in training.
+Learnable latent distributions can improve training loss, but require more care in training.
 
@@
 
-PZFlow is a Python package for building normalizing flows, with a focus on easy high-performance modeling of high-dimensional tabular data.
+PZFlow is a Python package for building normalizing flows, with a focus on features useful for forward modeling and density estimation for tabular data.
 Data is handled in Pandas DataFrames \citep{pandas}, while the normalizing flows are implemented in Jax \citep{jax}, which allows for efficient, parallelizable, GPU-enabled calculations for very large data sets.
-The code is easily installable from the Python Package Index\footnote{\url{https://pypi.org/project/pzflow/}} (PyPI) and is hosted on Github,\footnote{\url{https://github.com/jfcrenshaw/pzflow}}.
+The code is easily installable from the Python Package Index\footnote{\url{https://pypi.org/project/pzflow/}} (PyPI) and is hosted on Github\footnote{\url{https://github.com/jfcrenshaw/pzflow}}.
 The documentation\footnote{\url{https://jfcrenshaw.github.io/pzflow/}} includes tutorial notebooks demonstrating the features mentioned in this paper on different example problems.
 
-The rest of this paper will demonstrate using PZFlow for the statistical modeling of galaxy catalogs.
+The rest of this papaper will focus on forward modeling a photometric galaxy catalog, and photo-z inference.
 Section \ref{sec:galaxy-catalog} uses PZFlow to forward model a galaxy catalog, including photometry, spectroscopic redshifts (spec-z's), \emph{true} photo-z posteriors, ellipticities, and sizes.
@@
 
-In addition to the examples in this paper, PZFlow has already being used in various other projects:
+In addition to the examples in this paper, PZFlow has already being used in various other DESC projects:
 \begin{itemize}
@@
 In this section, we use PZFlow to forward model a photometric galaxy catalog for the Vera Rubin Observatory's Legacy Survey of Space and Time (LSST; \citealt{ivezic2019}).
-The advantage of using a catalog generated from a normalizing flow is that we have direct access to the exact distribution from which the data is drawn, enabling us to calculate true values for derived statistical products, such as the \emph{true} photo-z redshift posterior for each galaxy.
+The advantage of using a catalog generated from a normalizing flow is that we have direct access to the probability distribution from which the data is drawn, enabling us to calculate true values for derived statistical products, such as the \emph{true} photo-z redshift posterior for each galaxy.
 
@@
         \caption{
-            Distribution of true redshifts and noiseless photometry from the CosmoDC2 test set, compared to a sample drawn from the distribution learned by PZFlow.
+            Distribution of true redshifts and noisy photometry from the CosmoDC2 test set, compared to a sample drawn from the distribution learned by PZFlow.
             The close overlap of every pair-wise distribution demonstrates that PZFlow has learned the distribution in CosmoDC2 with high fidelity.
@@
         \caption{
-            Comparing the noiseless $r-i$ vs redshift distribution for galaxy samples from CosmoDC2 (left) and from the normalizing flow (right).
-            The high-redshift galaxies in CosmoDC2 lie along discrete tracks in color space due to the discrete number of galaxy SED templates used in the simulation.
-            PZFlow smooths over these discrete tracks, resulting in a color distribution that is smooth to high redshifts.
+            Comparing the $r-i$ vs redshift distribution for galaxy samples from CosmoDC2 without photometric noise (left), CosmoDC2 with photometric noise (middle), and from the normalizing flow (right).
+            High-redshift galaxies in CosmoDC2 lie along discrete tracks in redshift-color space.
+            Adding photometric noise ameliorates but does not totally remove the problem.
+            PZFlow produces a catalog with a smooth redshift-color distribution.
         }
@@
 To create a generative model of galaxy redshifts and photometry, we use the true redshifts and $ugrizy$ magnitudes from the CosmoDC2 simulation \citep{dc2, cosmodc2} of the LSST Dark Energy Science Collaboration (DESC).
-We selected all galaxies from CosmoDC2 with at most one band with a signal-to-noise ratio (SNR) less than 10, using the forecasted photometric errors from LSST year 10.
-Of these, we randomly selected $10^6$ galaxies and split them intro training and test sets consisting of 80\% and 20\% of the galaxies, respectively.
+We add photometric errors to the true $ugrizy$ magnitudes using the 10-year-depth LSST extended-source error model of our PhotErr package (see Appendix \ref{app:error-model}), and selected galaxies with a signal-to-noise ratio (SNR) greater than 5 in the $i$ band.
+Of these, we randomly selected $10^6$ galaxies and split them into training and test sets consisting of 80\% and 20\% of the galaxies, respectively.
 
@@
 In addition, Figure \ref{fig:smooth-color-dist} compares the distribution of galaxy $r-i$ vs redshift.
-The CosmoDC2 simulation is known to exhibit discrete tracks in this space at high redshift, due to the discrete number of SED templates used during simulation.
-These tracks are visible in the left panel.
-The right panel shows that PZFlow smooths over this discreteness, resulting in a color distribution that is smooth up to high redshifts.
-
+The CosmoDC2 simulation is known to exhibit discrete tracks in this space at high redshift, due to the way galaxies were assigned to a discrete set of SED templates during simulation.
+These tracks are easily visible in the left panel.
+Adding photometric errors largely smooths the distribution, but close inspection reveals that discrete tracks are still present in the data, especially at $z > 1$.
+The right panel shows the distribution produced by PZFlow, which has a smooth color distribution, even at high redshift.
 We note that these results were obtained without any extensive hyperparameter search, and that very similar (slightly worse results) are obtained without the \texttt{ColorTransform} bijection, demonstrating the flexibility of the method to adapt to unseen data sets.
 
-With this normalizing flow, we have an efficient CosmoDC2 emulator that produces a smooth distribution of realistic galaxies up to high-redshifts.
-We use this emulator to generate a catalog with $10^4$ galaxies.
-We add photometric errors using the LSST error model of our PhotErr package (see Appendix \ref{app:error-model}).
+With this normalizing flow, we have an efficient, probabilistic CosmoDC2 emulator that models a smooth color-redshift distribution up to redshift 3.
+We generate a catalog by sampling $10^4$ galaxies from the flow, each with noisy photometry and a true redshift.
 Importantly, since we have access to the probability distribution from which the galaxies were generated, we can calculate \emph{true} redshift posteriors for each galaxy.
@@
         \caption{
-            True redshift posteriors for a galaxy representing different amounts of information.
-            The black posterior is calculated with the true magnitudes;
-            the blue posterior is calculated after adding photometric errors;
-            the orange posterior is calculated after adding photometric errors, but with the errors convolved during posterior calculation;
-            the green posteriors is the same as the orange, except with the $u$ band marginalized over.
+            Example of redshift posteriors for a galaxy simulated with PZFlow.
+            The true redshift of the galaxy is marked by the vertical dashed red line, and the true redshift posterior for the galaxy is drawn in black.
+            Calculating the posterior while marginalizing over the $u$ band magnitude yields the posterior in blue.
+            Note that the $u$ band marginalization is only approximate, but increasing the resolution of the grid of $u$ band values causes the resultant posterior to converge.
         }
@@
 
-Since we have direct access to the probability distribution from which the photometry and redshifts are drawn, we can analytically calculate the true redshift posterior for each galaxy: $p(z|m)$ where $m$ is the vector of galaxy magnitudes.
-We note that this is not an estimate, like what would be returned by a photo-z estimator, but rather the truth, obtained from the model that generated the photometry and redshifts in the first place.
+Since we have direct access to the probability distribution from which the photometry and redshifts are drawn, we can analytically calculate the true redshift posterior for each galaxy: $p(z|\mathbf{\hat{m}})$ where $\mathbf{\hat{m}}$ is the vector of noisy galaxy magnitudes.
+We note this is not an estimate, like would be returned by a photo-z estimator, but rather the truth, obtained from the model that generated the photometry and redshifts in the first place.
 
-In addition to calculating the true posterior, $p(z|m)$, we can calculate a true posterior that is consistent with the photometric errors:
+When calculating these posteriors, for each galaxy we can also marginalize over the magnitudes in any missing bands, $n$:
 \begin{align}
-    p(z|m, \sigma_m) = \int p(z|\hat{m}) p(\hat{m}|m, \sigma_m) d\hat{m},
-    \label{eq:err-conv}
+    p(z|\mathbf{\hat{m}}) = \frac{1}{p(\mathbf{\hat{m}})} \int p(z, n, \mathbf{\hat{m}}) dn,
 \end{align}
-where $\sigma_m$ is the vector of photometric errors returned by the error model, and $\hat{m}$ are possible noisy observations of the true magnitudes, $m$.
-We can evaluate this integral numerically by sampling $\hat{m} \sim p(\hat{m}|m, \sigma_m)$, evaluating $p(z, \hat{m})$ on a redshift grid, and normalizing to obtain $p(z|\hat{m})$.
-Averaging $p(z|\hat{m})$ over the samples $\hat{m}$ yields $p(z|m, \sigma_m)$.
-This is the ``best case scenario'' redshift posterior that photo-z estimators hope to estimate.
+which can be calculated by evaluating $p(z, n, \mathbf{\hat{m}})$ on a grid of $z$ and possible values of $n$, summing over $n$ to yield $p(z, \mathbf{\hat{m}})$, and normalizing with respect to redshift to yield $p(z|\mathbf{\hat{m}})$.
+You may wish to marginalize over all values of $n$ if the galaxy was not observed in that band.
+This may occur, for example, when simulated a joint Euclid-LSST catalog \citep{euclid}, as not all galaxies will have photometry from both.
+You may also wish to marginalize over all values beyond a limiting magnitude to simulate a galaxy that was observed but not detected in the corresponding band.
+This might occur, for example, in the low wavelength bands of Lyman-dropout galaxies observed by LSST.
 
-We can also marginalize over any missing bands, $n$:
-\begin{align}
-    p(z|\hat{m}) = \frac{1}{p(\hat{m})} \int p(z, n, \hat{m}) dn,
-\end{align}
-which can be calculated by evaluating $p(z, n, \hat{m})$ on a grid of $z$ and $n$, summing over $n$ to yield $p(z, \hat{m})$, and normalizing with respect to redshift to yield $p(z|\hat{m})$.
-You can once again average over $\hat{m}$ samples to convolve the photometric errors.
-You may wish to marginalize over all values of $n$ if the galaxy was not observed in that band.
-This might occur, for example, if you include photometry from the Euclid Space Telescope \citep{euclid}, which will not have complete coverage of the LSST catalog.
-You may also wish to marginalize over all values beyond a limiting magnitude if the galaxy was observed, but not detected in that band.
-This might occur, for example, in the low wavelength bands of Lyman dropout galaxies observed by LSST.
+Note that this marginalization is only approximate, is weakens our ability to refer to these as ``true'' redshift posteriors.
+However, increasing the resolution of the grid used for $u$ band marginalization causes the resultant posteriors to converge.
+Thus, we believe it is still appropriate to treat these marginalized posteriors as the truth for the purpose of photo-z validation.
 
-To visualize the impacts of error convolution and band marginalization, Figure \ref{fig:posteriors} shows a number of redshift posteriors for a single example galaxy.
-The black posterior is calculated using the true galaxy magnitudes, while the blue posteriors is calculated after adding photometric errors.
-Calculating the posterior using the noisy photometry results in a biased posterior.
-The orange posterior has had the photometric errors convolved as in Equation \ref{eq:err-conv}.
-Convolving the errors broadens the posterior so that the true redshift lies within the support of the posterior.
-This broadening reflects the increased uncertainty due to the photometric errors.
-Finally, the green posterior has had the $u$ band marginalized.
-This posterior favors higher redshifts, but still comfortably agrees with the true redshift.
-This demonstrates how the $u$ band helps constrain the redshift, and the loss of this information leads to greater uncertainty.
+Redshift posteriors for an example galaxy are displayed in Figure \ref{fig:posteriors}.
+The black posterior is calculated using the full set of galaxy magnitudes.
+The true redshift, marked with by the vertical red line, nearly coincides with the mode of this posterior.
+The blue posterior has been calculated while marginalizing over the $u$ band.
+Throwing away the information in the $u$ band slightly broadens the posterior and shifts it toward higher redshifts.
 
@@
 This is important, as modern cosmology analyses are beginning to increasingly rely on full redshift posteriors \citep{mandelbaum2008,newman2022}.
-\citePZt showed that popular metrics for evaluating photo-z estimators using ensembles of photo-z posteriors can be misleading, and are not well suited to our science.
-PZFlow catalogs with \emph{true} redshift posteriors provide a path forward by enabling the evaluation of photo-z estimators on a per-posterior basis.
+\citePZt showed that popular metrics for evaluating photo-z estimators using ensembles of photo-z posteriors can be misleading, and are not well suited to the needs of precision cosmology.
+PZFlow catalogs with true redshift posteriors provide a path forward by enabling the evaluation of photo-z estimators on a per-posterior basis.
 
@@
             The vertical gray line labeled ``Truth'' denotes the true redshift of the galaxy.
-            Note these galaxies were specifically chosen for their broad, multimodal posteriors.
-            The posteriors of most galaxies are sharp and unimodal.
+            Averaging the posteriors from each flow in the ensemble approximately marginalizes over the neural network parameters, and smooths over the small-scale variations found in the posterior from any individual flow.
         }
@@
 Therefore, we will build a second flow that models these additional values conditioned on the galaxy redshift and magnitudes.
+In other words, we are modeling the full joint distribution via the decomposition
+\begin{align}
+    p(z, \mathbf{\hat{m}}, s, e) = p(z, \mathbf{\hat{m}}) \cdot p(s, e | z, \mathbf{\hat{m}}),
+    \label{eq:chain-rule}
+\end{align}
+where $s$ is the size (the half-light radius in arcseconds) and $e$ is the ellipticity.
+The first distribution on the right hand side of Equation \ref{eq:chain-rule} is modeled by our original flow, and the second distribution will be modeled using the new conditional flow.
 Note that while we have only chosen to model these additional two properties, any other values you desire can be similarly modeled.
@@
 \end{align}
-The RQ-RSC acts on the two dimensional space of size and ellipticity, but also takes the galaxy redshift and magnitudes as inputs (see Figure \ref{fig:conditional-coupling}).
+The RQ-RSC acts on the two dimensional space of size and ellipticity, but also takes the galaxy redshift and magnitudes as inputs (see the conditional variables in green in Figure \ref{fig:conditional-coupling}).
 The redshifts and magnitudes are transformed to have zero mean and unit variance before being input to the neural network\footnote{These variables are standard scaled instead of mapped onto the domain [-5, 5], because the neural network that parameterizes the splines has no limit on inputs, unlike the splines themselves, which are limited to the range [-5, 5].} that parameterizes the splines.
@@
 
-After training the flow (see Appendix \ref{app:training-details}), we sample a size and ellipticity for each galaxy in the PZFlow catalog created in the previous section (conditioned on the true magnitudes), and plot the distribution of these features against the distribution in the test set (Figure \ref{fig:conditional-corner}).
+After training the flow (see Appendix \ref{app:training-details}), we sample a size and ellipticity for each galaxy in the PZFlow catalog created in the previous section (conditioned on the redshift and noisy magnitudes), and plot the distribution of these features against the distribution in the test set (Figure \ref{fig:conditional-corner}).
 Once again, we see the normalizing flow does a good job of emulating the CosmoDC2 galaxy distribution.
 
-The final simulated catalog consists of $10^4$ galaxies, each with a redshift, $ugrizy$ magnitudes including photometric errors, a true photo-z posterior consistent with the photometric errors, a size, and an ellipticity.
+The final simulated catalog consists of $10^4$ galaxies, each with a redshift, noisy $ugrizy$ magnitudes, a true photo-z posterior, a size, and an ellipticity.
+We use the magnitudes, size, and ellipticity to estimate the photometric errors using the 10-year-depth LSST extended-source error model of PhotErr.
 This small catalog was generated for visualization purposes, but the normalizing flows can be used to generate catalogs of arbitrarily large size.
+In particular, one can generate catalogs with a millions of galaxies, or can augment existing catalogs, in only a few minutes, substantially faster than is required by large-scale simulations like CosmoDC2.
 
@@
 This makes them useful tools for estimating posterior distributions for galaxy properties, conditioned on observed features of the galaxy.
-A common example of this in cosmology is photometric redshift estimation, in which you estimate the redshift of a galaxy using its magnitude in several photometric bands.
-In this section, we use PZFlow as a photo-z estimator to demonstrate using normalizing flows for density estimation.
+In this section, we apply this to photo-z estimation for the simulated catalog from the previous section.
 
@@
 By averaging over the individual posteriors, we select for features that are common between models, while smoothing over features that are present in only a single model.
+The first example galaxy in Figure~\ref{fig:ensemble-posteriors} is at $z < 1$, and all flows in the ensemble return essentially the same narrow redshift posterior.
+This is typical for low-redshift galaxies whose photo-z's are relatively well constrained by LSST photometry.
+
+The other two example galaxies, however, are in the $1.5 < z < 2.6$ redshift range, where the Balmer Break (at $\sim 4000$~\AA) has redshifted out of LSST's wavelength coverage, while the Lyman Limit (at $912$~\AA) has not yet redshifted into LSST's wavelength coverage.
+As a result, these posteriors are much broader and less well constrained.
+For these two galaxies, the flows in the ensemble return posteriors with different small-scale variations, and the best-estimate redshift (i.e.
+the mode of the redshift posterior) varies by as much as 0.5 for each set of posteriors.
+Marginalizing over the individual posteriors smooths over these variations.
 We can also treat the ensemble of posteriors as a distribution over possible posteriors, which will allow for more consistent error calibration in cosmological analyses \citep{zhang2023}.
-
-We can calculate posteriors for galaxies with missing magnitudes by marginalizing over the missing magnitudes, as described in Section \ref{sec:true-posteriors}.
-An example is shown in Figure \ref{fig:posterior_marginalized}, which compares a posterior calculated using every LSST band to a posterior with the $u$ band marginalized.
-Sacrificing the information in the $u$ band increases the uncertainty of the redshift inference.
-In this instance, a second potential higher-redshift mode is created.
-Conversely, this demonstrates how the additon of $u$ band information rules out the higher redshift solution for this galaxy.
 
@@
             The bias, scatter, and outlier fraction of the photo-z point estimates as a function of true galaxy redshift.
-            The dashed black lines represent the requirements for LSST cosmology as stated in the LSST DESC SRD \citep{descSRD}.
+            The dashed black lines represent the upper limits for LSST cosmology as stated in the LSST DESC SRD \citep{descSRD}.
             These lines are to provide a sense of scale for these metrics.
@@
 There are the common photo-z ``wings'', indicating redshifts where important spectral features are transitioning between neighboring photometric bands.
-There is also a small population of high-redshift objects that were mistakenly identified as very low redshift objects.
 This point estimate plot is comparable to other high-performance machine learning photo-z estimators when provided with representative training sets \citep{sanchez2014}.
@@
 Another common metric is the probability integral transform (PIT) (see e.g. \citePZa, \citealt{dey2022}).
-The PIT metric is used to determine if posteriors are well calibrated, i.e. if true values fall within X\% confidence intervals X\% of the time.
+The PIT metric is used to determine if posteriors are well calibrated.
+For example, if the photo-z posteriors produced by an estimator are well calibrated, you would expect the true redshifts of the galaxies to fall within the photo-z 50\% confidence intervals 50\% of the time.
 The PIT histogram for our estimator is shown in Figure \ref{fig:pit-histogram}.
-Ideally, this histogram would be uniform and match the dashed horizontal line.
+Ideally, this histogram would be uniform and match the dashed horizontal line, which represents perfect calibration.
 The fact that the histogram bulges at the center indicates that our estimator is too conservative -- i.e. the posteriors it produces are too broad.
@@
 This results in overly conservative posterior predictions.
-The low values at the edges of the PIT histogram indicate the relative rarity of catastrophic outliers, which is also reflected in the far right panel of Figure \ref{fig:binned-metrics}, where you can see that our estimator meets the requirement on the outlier fraction at all redshifts.
+
+The low values at the edges of the PIT histogram indicate the relative rarity of catastrophic outliers, which is also reflected in the far right panel of Figure \ref{fig:binned-metrics}, where you can see our estimator meets the requirement on the outlier fraction at all redshifts.
 There is also a slight rightward tilt.
 This indicates a small negative bias, which reflects the intrinsic prior towards smaller redshifts, as this is where the majority of galaxies in the training set lie.
-This negative bias is also visible in the far left panel of Figure \ref{fig:binned-metrics}.
+This negative bias is visible for high-redshift galaxies in the far left panel of Figure \ref{fig:binned-metrics}.
 
@@
 The methods introduced in this paper enable the creation of galaxy catalogs for which each galaxy has a \emph{true} redshift posterior, which will enable more comprehensive evaluation of photo-z estimators.
-An full evaluation of photo-z estimators on a posterior-by-posterior basis is a major goal of the LSST DESC.
+Full evaluation of photo-z estimators on a posterior-by-posterior basis is a major goal of the LSST DESC, and will be the focus of forthcoming work.
 
@@
             Training losses for the four flows in the flow ensemble.
-            We have zoomed in to the bottom of the loss curve so you can see that each of the flows converges to a different minimum loss.
+            We have zoomed in to the bottom of the loss curve so you can see that each of the flows converges to a slightly different minimum loss.
         }
@@
 
-For the main flow in Section \ref{sec:galaxy-catalog}, we trained for 150 epochs.
-We used the Adam optimizer \citep{adam}, starting with a learning rate of $10^{-3}$.
-We decreased the learning rate by a factor of 10 every 50 epochs.
+For the main flow in Section \ref{sec:galaxy-catalog}, we trained for 200 epochs.
+We used the Adam optimizer \citep{adam}, starting with a learning rate of $10^{-5}$.
+We decreased the learning rate by a factor of 10 after the 100th and 150th epochs.
 Training took 7 minutes on a Tesla P100 12GB GPU.
@@
 
-For the conditional flow in \ref{sec:galaxy-catalog}, we trained for 450 epochs.
-Again, we used Adam with an initial learning rate of $10^{-3}$.
-We decreased the learning rate by a factor of 10 every 150 epochs.
+For the conditional flow in \ref{sec:galaxy-catalog}, we trained for 600 epochs.
+Again, we used Adam with an initial learning rate of $10^{-5}$.
+We decreased the learning rate by a factor of 10 every 200 epochs.
 The training loss for this flow is in the right panel of Figure \ref{fig:galaxy-flow-losses}.
 
-For each of the flows that make up the flow ensemble in Section \ref{sec:photo-z}, we trained for 150 epochs using the Adam optimizer.
-We started each with a learning rate of $10^{-4}$, and decreased the learning rate by a factor of 10 every 50 epochs.
+For each of the flows that make up the flow ensemble in Section \ref{sec:photo-z}, we trained for 200 epochs using the Adam optimizer.
+We started each with a learning rate of $10^{-4}$, which we decreased by a factor of 10 after the 100th and 150th epochs.
 The training loss for the ensemble is in Figure \ref{fig:ensemble-losses}.
-You can see that each of the flows achieves a different minimum loss.
-Apparently, each has found a different potential solution in the neural network's parameter space.
+Each flow achieved nearly the same training loss.
 

### 2023-11-22 src/scripts/photo-z/plot_ensemble_losses.py

@@ -16,7 +16,7 @@ fig, ax = plt.subplots(figsize=(3, 3), constrained_layout=True)
 for i, loss in enumerate(losses.values()):
     ax.plot(loss, alpha=0.5, label=f"Flow {i+1}")
 ax.legend()
-ax.set(xlabel="Epochs", ylabel="Loss")
+ax.set(xlabel="Epochs", ylabel="Loss", ylim=(-2.5, -1))
 
 # save the figure
 fig.savefig(paths.figures / "ensemble_losses.pdf")

### 2023-11-28 src/tex/ms.tex
More text revisions.

@@
 % Define document class
-\documentclass[twocolumn,twocolappendix,linenumbers]{aastex631}
+\documentclass[twocolumn,twocolappendix]{aastex631}
 \usepackage{showyourwork}
@@
 A common choice for the latent distribution is the Normal distribution, hence the name \emph{normalizing} flow.
+
 Unlike other deep generative models like Generative Adversarial Networks (GANs; \citealt{goodfellow2014}) and Variational Autoencoders (VAEs; \citealt{kingma2014}), normalizing flows provide a deterministic mapping between the data and latent spaces.
@@
 
-In Section \ref{sec:nf} we provide the background on normalizing flows required to understand PZFlow, which we describe in Section \ref{sec:pzflow}.
-In Section \ref{sec:fwd-model}, we demonstrate using PZFlow to simulate a galaxy catalog where each object has a redshift, photometry in the six LSST bands, a true photo-z posterior, a size, and an ellipticity.
-In Section \ref{sec:photo-z}, we demonstrate using PZFlow as a density estimator by estimating photo-z's for our simulated catalog.
-We conclude in Section \ref{sec:conclusion}.
+In Section~\ref{sec:nf} we provide the background on normalizing flows required to understand PZFlow, which we describe in Section~\ref{sec:pzflow}.
+In Section~\ref{sec:galaxy-catalog}, we demonstrate using PZFlow to simulate a galaxy catalog where each object has a redshift, photometry in the six LSST bands, a true photo-z posterior, a size, and an ellipticity.
+In Section~\ref{sec:photo-z}, we demonstrate using PZFlow as a density estimator by estimating photo-z's for our simulated catalog.
+We conclude in Section~\ref{sec:conclusion}.
 
@@
 
-Figure \ref{fig:two-moons} shows an example of a normalizing flow that transforms the scikit-learn \citep{sklearn} two moons distribution into a uniform distribution.
+Figure~\ref{fig:two-moons} shows an example of a normalizing flow that transforms the scikit-learn \citep{sklearn} two moons distribution into a uniform distribution.
 The data points are colored by quadrant to visualize their image under $f$.
@@
 The following sections discuss how to build a normalizing flow to model data with various features.
-Section \ref{sec:bijections} discusses the bijection $f$ and introduces the building blocks from which our bijections will be built;
-Section \ref{sec:latent} discusses how to choose an appropriate latent distribution for your data;
-Section \ref{sec:conditional} describes how to build a flow that models a conditional distribution;
-Section \ref{sec:periodic} explains how to model data with periodic topology;
-finally Section \ref{sec:discrete} explains how to model data with discrete variables.
+Section~\ref{sec:bijections} discusses the bijection $f$ and introduces the building blocks from which our bijections will be built;
+Section~\ref{sec:latent} discusses how to choose an appropriate latent distribution for your data;
+Section~\ref{sec:conditional} describes how to build a flow that models a conditional distribution;
+Section~\ref{sec:periodic} explains how to model data with periodic topology;
+finally Section~\ref{sec:discrete} explains how to model data with discrete variables.
 
@@
 where $g : \R^{D-d} \times \R^d \to \R^{D-d}$ is an invertible \emph{coupling law}, and $m$ is a \emph{coupling function} defined on $\R^d$.
-This is illustrated in Figure \ref{fig:coupling}.
+This is illustrated in Figure~\ref{fig:coupling}.
 The advantage of this structure is that the Jacobian is triangular,
@@
 The result is a bijection that achieves state-of-the-art performance and efficiency for forward modeling and density estimation \citep{kobyzev2020}, and are flexible enough to model complex distributions with multiple discontinuities and hundreds of modes.
-In addition, they are easily adaptable for flows with periodic topology (Section \ref{sec:periodic}).
+In addition, they are easily adaptable for flows with periodic topology (Section~\ref{sec:periodic}).
 For more details, see \citet{durkan2019}.
@@
 
-For an example of building an application-specific data processing bijection, see the \emph{Color Transform} bijection defined in Section \ref{sec:fwd-model}, which maps galaxy magnitudes to galaxy colors.
-See section \ref{sec:discrete} for data processing bijections that enable modeling of discrete data.
+For an example of building an application-specific data processing bijection, see the \emph{Color Transform} bijection defined in Section~\ref{sec:fwd-model}, which maps galaxy magnitudes to galaxy colors.
+See section~\ref{sec:discrete} for data processing bijections that enable modeling of discrete data.
 
@@
 you only need to make the replacement $f(x) \to f(x;z)$, where $z$ is a vector of conditions \citep{winkler2019}.
-This is illustrated in Figure \ref{fig:conditional-coupling}, which is a modification of Figure \ref{fig:coupling} to include the input of conditional variables to the coupling function $m$.
+This is illustrated in Figure~\ref{fig:conditional-coupling}, which is a modification of Figure~\ref{fig:coupling} to include the input of conditional variables to the coupling function $m$.
 In practice, since $m$ is usually a neural network, this amounts to just appending the conditions $z$ to the inputs of the neural network.
@@
 to match an identity transform for inputs outside of the range $[-\pi, \pi]$.
-By choosing a latent distribution with compact support in the range $[-\pi, \pi]$, we ensure that no samples will lie outside the range of the splines, and so we can relax the boundary condition of Equation \ref{eq:df=1} in favor of the boundary condition in Equation \ref{eq:df=df}.
+By choosing a latent distribution with compact support in the range $[-\pi, \pi]$, we ensure that no samples will lie outside the range of the splines, and so we can relax the boundary condition of Equation~\ref{eq:df=1} in favor of the boundary condition in Equation~\ref{eq:df=df}.
 Spline transforms with this relaxed boundary condition are named \emph{Circular Splines} by \citet{rezende2020}.
@@
 The circular spline construction above is easily generalized to n-spheres and n-tori: $S^n \to [-1, 1]^{n-1} \times S^1$ and $T^n \to (S^1)^n$ (see \citealt{rezende2020} for more details).
-We can model the joint distribution of periodic and non-periodic variables with RQ-RSCs simply by choosing appropriate bounds $B$ for each dimension, and by swapping boundary condition \ref{eq:df=1} for condition \ref{eq:df=df} for any periodic dimensions.
+We can model the joint distribution of periodic and non-periodic variables with RQ-RSCs simply by choosing appropriate bounds $B$ for each dimension, and by swapping boundary condition~\ref{eq:df=1} for condition~\ref{eq:df=df} for any periodic dimensions.
 
@@
 
-The rest of this papaper will focus on forward modeling a photometric galaxy catalog, and photo-z inference.
-Section \ref{sec:galaxy-catalog} uses PZFlow to forward model a galaxy catalog, including photometry, spectroscopic redshifts (spec-z's), \emph{true} photo-z posteriors, ellipticities, and sizes.
-Section \ref{sec:photo-z} uses PZFlow for photo-z estimation, demonstrating the power of PZFlow as a density estimator, including numerous useful features for photo-z estimation.
+The rest of this paper will focus on forward modeling a photometric galaxy catalog, and photo-z inference.
+Section~\ref{sec:galaxy-catalog} uses PZFlow to forward model a galaxy catalog, including photometry, spectroscopic redshifts (spec-z's), true photo-z posteriors, ellipticities, and sizes.
+Section~\ref{sec:photo-z} uses PZFlow for photo-z estimation, demonstrating the power of PZFlow as a density estimator, including numerous useful features for photo-z estimation.
 
@@
 In this section, we use PZFlow to forward model a photometric galaxy catalog for the Vera Rubin Observatory's Legacy Survey of Space and Time (LSST; \citealt{ivezic2019}).
-The advantage of using a catalog generated from a normalizing flow is that we have direct access to the probability distribution from which the data is drawn, enabling us to calculate true values for derived statistical products, such as the \emph{true} photo-z redshift posterior for each galaxy.
+The advantage of using a catalog generated from a normalizing flow is that we have direct access to the probability distribution from which the data is drawn, enabling us to calculate true values for derived statistical products, such as the true photo-z redshift posterior for each galaxy.
 
-In Section \ref{sec:fwd-model} we construct a normalizing flow to model the galaxy redshifts and photometry and generate a new simulated catalog.
-In Section \ref{sec:true-posteriors}, we calculate true redshift posteriors for the new catalog.
-In Section \ref{sec:fwd-model-conditional} we build a conditional flow to add additional galaxy properties to the catalog.
+In Section~\ref{sec:fwd-model} we construct a normalizing flow to model the galaxy redshifts and photometry and generate a new simulated catalog.
+In Section~\ref{sec:true-posteriors}, we calculate true redshift posteriors for the new catalog.
+In Section~\ref{sec:fwd-model-conditional} we build a conditional flow to add additional galaxy properties to the catalog.
 
@@
             High-redshift galaxies in CosmoDC2 lie along discrete tracks in redshift-color space.
-            Adding photometric noise ameliorates but does not totally remove the problem.
+            Adding photometric noise somewhat smooths but does not totally remove these tracks.
             PZFlow produces a catalog with a smooth redshift-color distribution.
@@
 To create a generative model of galaxy redshifts and photometry, we use the true redshifts and $ugrizy$ magnitudes from the CosmoDC2 simulation \citep{dc2, cosmodc2} of the LSST Dark Energy Science Collaboration (DESC).
-We add photometric errors to the true $ugrizy$ magnitudes using the 10-year-depth LSST extended-source error model of our PhotErr package (see Appendix \ref{app:error-model}), and selected galaxies with a signal-to-noise ratio (SNR) greater than 5 in the $i$ band.
+We add photometric errors to the true $ugrizy$ magnitudes using the 10-year-depth LSST extended-source error model of our PhotErr package (see Appendix~\ref{app:error-model}), and selected galaxies with a signal-to-noise ratio (SNR) greater than 5 in the $i$ band.
 Of these, we randomly selected $10^6$ galaxies and split them into training and test sets consisting of 80\% and 20\% of the galaxies, respectively.
+We then train a normalizing flow to learn the distribution $p(z, \mathbf{\hat{m}})$, where $z$ is the true redshift, and $\mathbf{\hat{m}}$ is the vector of noisy magnitudes in the LSST bands.
 
@@
 
-The next layer, Shift Bounds, is the data processing bijection defined in Section \ref{sec:data-processing}, which maps the range of the data onto the support of the RQ-RSC.
+The next layer, Shift Bounds, is the data processing bijection defined in Section~\ref{sec:data-processing}, which maps the range of the data onto the support of the RQ-RSC.
 Note that since Shift Bounds is on the ``other side'' of the Color Transform, we need to map the ranges of the colors $u-g$, $g-r$, etc.
@@
 
-The final layer is an RQ-RSC, described in detail in Section \ref{sec:rq-rsc}.
+The final layer is an RQ-RSC, described in detail in Section~\ref{sec:rq-rsc}.
 This layer performs the heavy lifting of transforming the data distribution into the uniform latent distribution.
@@
 We use $K=16$ spline knots.
+This number was chosen to be large enough to capture the complexity of the data, but small enough so that the flow smooths over the discrete tracks in Figure~\ref{fig:smooth-color-dist} (more on this below).
 
-After training the flow (see Appendix \ref{app:training-details}), we assess the results by drawing $10^4$ galaxies from the trained flow, and plotting their distribution against  $10^4$ galaxies from the test set (Figure \ref{fig:main-corner}).
+After training the flow (see Appendix~\ref{app:training-details}), we assess the results by drawing $10^4$ galaxies from the trained flow, and plotting their distribution against  $10^4$ galaxies from the test set (Figure~\ref{fig:main-corner}).
 We see the normalizing flow has done an excellent job of reproducing the distribution of galaxies in CosmoDC2, without any unusual artifacts or outliers.
-In addition, Figure \ref{fig:smooth-color-dist} compares the distribution of galaxy $r-i$ vs redshift.
-The CosmoDC2 simulation is known to exhibit discrete tracks in this space at high redshift, due to the way galaxies were assigned to a discrete set of SED templates during simulation.
+In addition, Figure~\ref{fig:smooth-color-dist} compares the distribution of galaxy $r-i$ vs redshift.
+High-redshift galaxies in the CosmoDC2 simulation lie on discrete tracks in this space due to the way galaxies were assigned to a discrete set of SED templates during simulation.
 These tracks are easily visible in the left panel.
-Adding photometric errors largely smooths the distribution, but close inspection reveals that discrete tracks are still present in the data, especially at $z > 1$.
+Adding photometric errors somewhat smooths the distribution, but close inspection reveals there is still granularity in the distribution of high-redshift galaxies.
 The right panel shows the distribution produced by PZFlow, which has a smooth color distribution, even at high redshift.
@@
 We generate a catalog by sampling $10^4$ galaxies from the flow, each with noisy photometry and a true redshift.
-Importantly, since we have access to the probability distribution from which the galaxies were generated, we can calculate \emph{true} redshift posteriors for each galaxy.
+Importantly, since we have access to the probability distribution from which the galaxies were generated, we can calculate true redshift posteriors for each galaxy.
 This is the subject of the next section.
@@
 
-When calculating these posteriors, for each galaxy we can also marginalize over the magnitudes in any missing bands, $n$:
+When calculating these posteriors, for each galaxy we can also marginalize over the magnitudes in any missing bands.
+Imagine, for a galaxy, that we partition the vector of magnitudes $\mathbf{\hat{m}}$ into an observed set $\mathbf{\hat{m}}_0$ and a missing set $\mathbf{\hat{m}}_\text{x}$.
+We can marginalize over the missing magnitudes when calculating the posterior
 \begin{align}
-    p(z|\mathbf{\hat{m}}) = \frac{1}{p(\mathbf{\hat{m}})} \int p(z, n, \mathbf{\hat{m}}) dn,
+    p(z|\mathbf{\hat{m}}_0) = \frac{1}{p(\mathbf{\hat{m}}_0)} \int p(z, \mathbf{\hat{m}}_0, \mathbf{\hat{m}}_\text{x}) d\mathbf{\hat{m}}_\text{x},
 \end{align}
-which can be calculated by evaluating $p(z, n, \mathbf{\hat{m}})$ on a grid of $z$ and possible values of $n$, summing over $n$ to yield $p(z, \mathbf{\hat{m}})$, and normalizing with respect to redshift to yield $p(z|\mathbf{\hat{m}})$.
-You may wish to marginalize over all values of $n$ if the galaxy was not observed in that band.
+which can be calculated by evaluating $p(z, \mathbf{\hat{m}})$ on a grid of $z$ and possible values of $\mathbf{\hat{m}}_\text{x}$, summing over $\mathbf{\hat{m}}_\text{x}$ to yield $p(z, \mathbf{\hat{m}}_0)$, and normalizing with respect to redshift to yield $p(z|\mathbf{\hat{m}}_0)$.
+PZFlow possess a flexible method for performing this marginalization: the grid for each band in $\mathbf{\hat{m}}_\text{x}$ can be a function of other galaxy properties (e.g.
+the observed magnitudes, $\mathbf{\hat{m}}_0$).
+
+You may wish to marginalize over all values of $\mathbf{\hat{m}}_\text{x}$ if the galaxy was not observed in those bands.
 This may occur, for example, when simulated a joint Euclid-LSST catalog \citep{euclid}, as not all galaxies will have photometry from both.
-You may also wish to marginalize over all values beyond a limiting magnitude to simulate a galaxy that was observed but not detected in the corresponding band.
+You may also wish to marginalize over all values beyond the limiting magnitudes to simulate a galaxy that was observed but not detected in the corresponding bands.
 This might occur, for example, in the low wavelength bands of Lyman-dropout galaxies observed by LSST.
 
-Note that this marginalization is only approximate, is weakens our ability to refer to these as ``true'' redshift posteriors.
+Note this marginalization is only approximate, and therefore weakens our ability to refer to these as ``true'' redshift posteriors.
 However, increasing the resolution of the grid used for $u$ band marginalization causes the resultant posteriors to converge.
@@
 
-Redshift posteriors for an example galaxy are displayed in Figure \ref{fig:posteriors}.
+Redshift posteriors for an example galaxy are displayed in Figure~\ref{fig:posteriors}.
 The black posterior is calculated using the full set of galaxy magnitudes.
-The true redshift, marked with by the vertical red line, nearly coincides with the mode of this posterior.
+The true redshift, marked by the vertical red line, nearly coincides with the mode of this posterior.
 The blue posterior has been calculated while marginalizing over the $u$ band.
@@
 
-Calculating these posteriors enables direct comparison of true redshift posteriors (consistent with photometric errors and missing bands) with the redshift posteriors estimated by photo-z estimators.
+Calculating these posteriors enables direct comparison of true redshift posteriors with the redshift posteriors estimated by photo-z estimators.
 This is important, as modern cosmology analyses are beginning to increasingly rely on full redshift posteriors \citep{mandelbaum2008,newman2022}.
@@
             Averaging the posteriors from each flow in the ensemble approximately marginalizes over the neural network parameters, and smooths over the small-scale variations found in the posterior from any individual flow.
+            The first panel is a galaxy with a narrow and unimodal redshift posterior, while the next two panels demonstrate broad, multimodal posteriors, which is typical of galaxies in the range $1.5 < z < 2.6$.
         }
@@
 where $s$ is the size (the half-light radius in arcseconds) and $e$ is the ellipticity.
-The first distribution on the right hand side of Equation \ref{eq:chain-rule} is modeled by our original flow, and the second distribution will be modeled using the new conditional flow.
-Note that while we have only chosen to model these additional two properties, any other values you desire can be similarly modeled.
+The first distribution on the right hand side of Equation~\ref{eq:chain-rule} is modeled by our original flow, and the second distribution will be modeled using the new conditional flow.
+While we have only chosen to model these additional two properties, any other values you desire can be similarly modeled.
 
@@
 \end{align}
-The RQ-RSC acts on the two dimensional space of size and ellipticity, but also takes the galaxy redshift and magnitudes as inputs (see the conditional variables in green in Figure \ref{fig:conditional-coupling}).
+The RQ-RSC acts on the two dimensional space of size and ellipticity, but also takes the galaxy redshift and magnitudes as inputs (see the conditional variables in green in Figure~\ref{fig:conditional-coupling}).
 The redshifts and magnitudes are transformed to have zero mean and unit variance before being input to the neural network\footnote{These variables are standard scaled instead of mapped onto the domain [-5, 5], because the neural network that parameterizes the splines has no limit on inputs, unlike the splines themselves, which are limited to the range [-5, 5].} that parameterizes the splines.
@@
 
-After training the flow (see Appendix \ref{app:training-details}), we sample a size and ellipticity for each galaxy in the PZFlow catalog created in the previous section (conditioned on the redshift and noisy magnitudes), and plot the distribution of these features against the distribution in the test set (Figure \ref{fig:conditional-corner}).
+After training the flow (see Appendix~\ref{app:training-details}), we sample a size and ellipticity for each galaxy in the PZFlow catalog created in the previous section (conditioned on the redshift and noisy magnitudes), and plot the distribution of these features against the distribution in the test set (Figure~\ref{fig:conditional-corner}).
 Once again, we see the normalizing flow does a good job of emulating the CosmoDC2 galaxy distribution.
+We note that, if desired, the TARP test of \citet{lemos2023} can provide a quantitative test of the fidelity of the conditional flow.
 
@@
 This small catalog was generated for visualization purposes, but the normalizing flows can be used to generate catalogs of arbitrarily large size.
-In particular, one can generate catalogs with a millions of galaxies, or can augment existing catalogs, in only a few minutes, substantially faster than is required by large-scale simulations like CosmoDC2.
+In particular, in only a few minutes, one can generate new catalogs or augment existing catalogs with millions of galaxies.
+This is substantially faster than re-running large scale simulations like CosmoDC2.
 
@@
 This makes them useful tools for estimating posterior distributions for galaxy properties, conditioned on observed features of the galaxy.
-In this section, we apply this to photo-z estimation for the simulated catalog from the previous section.
+In this section, we demonstrate this by applying PZFlow to photo-z estimation for the simulated catalog from the previous section.
 
@@
 
-When forward modeling in Section \ref{sec:galaxy-catalog}, we wanted a realistic model that captured the relevant correlations between galaxy photometry, redshift, shape, and size.
+When forward modeling in Section~\ref{sec:galaxy-catalog}, we wanted a realistic model that captured the relevant correlations between galaxy photometry, redshift, shape, and size.
 However, when estimating redshifts, we do not simply want a realistic model, but rather a model that matches our specific galaxy sample as closely as possible.
@@
 
-We train an ensemble of 4 normalizing flows, each with the same architecture and training schedule as the regular flow described in Section \ref{sec:galaxy-catalog}.
+We train an ensemble of 4 normalizing flows, each with the same architecture and training schedule as the regular flow described in Section~\ref{sec:galaxy-catalog}.
 With PZFlow, this is as simple as swapping \texttt{FlowEnsemble} for \texttt{Flow} in the code.
 
-For the training set, we use 100,000 galaxies from the catalog created in Section \ref{sec:galaxy-catalog}.
+For the training set, we use 100,000 galaxies from the catalog created in Section~\ref{sec:galaxy-catalog}.
 Each galaxy in the training set has a true redshift and observed magnitudes in the $ugrizy$ bands, with corresponding photometric errors.
@@
 \begin{align}
-    m \sim p(\hat{m}, \sigma_m),
+    \mathbf{m} \sim p(\mathbf{\hat{m}}, \mathbf{\sigma_m}),
 \end{align}
-where $\hat{m}$ are the observed magnitudes with photometric errors $\sigma_m$, and $p(\hat{m}, \sigma_m)$ is a Gaussian in flux space.
-This allows our ensemble of flows to approximate the distribution $p(z, m)$.
-For more details on training the ensemble, see Appendix \ref{app:training-details}.
+where $\mathbf{\hat{m}}$ are the observed magnitudes with photometric errors $\mathbf{\sigma_m}$, and $p(\mathbf{\hat{m}}, \mathbf{\sigma_m})$ is a Gaussian in flux space.
+This allows our ensemble of flows to approximate the distribution $p(z, \mathbf{m})$, where \mathbf{m} is the vector of true magnitudes for the galaxy.
+For more details on training the ensemble, see Appendix~\ref{app:training-details}.
 
@@
 
-For each flow in the ensemble, we estimate $p(z, m)$ by marginalizing over the photometric errors:
+For each flow in the ensemble, we estimate $p(z, \mathbf{m})$ by marginalizing over the photometric errors:
 \begin{align}
-    p(z| \hat{m}, \sigma_m) \propto \int dm \, p(z, m) \, p(m| \hat{m}, \sigma_m),
+    p(z| \mathbf{\hat{m}}, \mathbf{\sigma_m}) \propto \int p(z, \mathbf{m}) \, p(\mathbf{m}| \mathbf{\hat{m}}, \mathbf{\sigma_m}) \, d\mathbf{m},
 \end{align}
-which is estimated by sampling $m \sim p(\hat{m}, \sigma_m)$ and averaging $p(z, m)$ over these samples.
-We then average the $p(z, m)$ from each flow, and normalize with respect to the redshift grid.
+which is estimated by sampling $\mathbf{m} \sim p(\mathbf{\hat{m}}, \mathbf{\sigma_m})$ and averaging $p(z, \mathbf{m})$ over these samples.
+We then average the $p(z, \mathbf{m})$ from each flow, and normalize with respect to the redshift grid.
 This provides a redshift posterior for each galaxy.
 
-Posteriors for three galaxies can be seen in Figure \ref{fig:ensemble-posteriors}.
+Posteriors for three galaxies can be seen in Figure~\ref{fig:ensemble-posteriors}.
 Each flow produces a PDF which may contain slightly different features in each case.
@@
 
-Figure \ref{fig:point-estimates} compares the photo-z point estimates to the true redshifts.
+Figure~\ref{fig:point-estimates} compares the photo-z point estimates to the true redshifts.
 The point estimates for most galaxies lie along the diagonal, indicating strong performance.
@@
 
-Figure \ref{fig:binned-metrics} shows the photo-z point estimate metrics from the LSST DESC Science Requirements Document (SRD; \citealt{descSRD}) as a function of true redshift.
+Figure~\ref{fig:binned-metrics} shows the photo-z point estimate metrics from the LSST DESC Science Requirements Document (SRD; \citealt{descSRD}) as a function of true redshift.
 The \emph{bias} is defined as the median of $\Delta z$; the \emph{scatter} is defined as $\text{IQR} / 1.349$, where IQR is the interquartile range of $\Delta z$; the \emph{outlier fraction} is defined as the fraction of galaxies for which $\Delta z$ is greater than three times the scatter.
@@
 The PIT metric is used to determine if posteriors are well calibrated.
-For example, if the photo-z posteriors produced by an estimator are well calibrated, you would expect the true redshifts of the galaxies to fall within the photo-z 50\% confidence intervals 50\% of the time.
-The PIT histogram for our estimator is shown in Figure \ref{fig:pit-histogram}.
+For example, if the photo-z posteriors produced by an estimator are well calibrated, you would expect the true redshifts of the galaxies to fall within 50\% confidence intervals 50\% of the time.
+The PIT histogram for our estimator is shown in Figure~\ref{fig:pit-histogram}.
 Ideally, this histogram would be uniform and match the dashed horizontal line, which represents perfect calibration.
-The fact that the histogram bulges at the center indicates that our estimator is too conservative -- i.e. the posteriors it produces are too broad.
+The fact that the histogram bulges at the center indicates our estimator is too conservative -- i.e. the posteriors it produces are too broad.
 This can be explained by the fact that normalizing flows exhibit mode covering behavior (the opposite of the mode collapse seen in GANs; \citealt{salimans2016}).
@@
 
-The low values at the edges of the PIT histogram indicate the relative rarity of catastrophic outliers, which is also reflected in the far right panel of Figure \ref{fig:binned-metrics}, where you can see our estimator meets the requirement on the outlier fraction at all redshifts.
+The low values at the edges of the PIT histogram indicate the relative rarity of catastrophic outliers, which is also reflected in the far right panel of Figure~\ref{fig:binned-metrics}, where you can see our estimator meets the requirement on the outlier fraction at all redshifts.
 There is also a slight rightward tilt.
 This indicates a small negative bias, which reflects the intrinsic prior towards smaller redshifts, as this is where the majority of galaxies in the training set lie.
-This negative bias is visible for high-redshift galaxies in the far left panel of Figure \ref{fig:binned-metrics}.
+This negative bias is visible for high-redshift galaxies in the far left panel of Figure~\ref{fig:binned-metrics}.
 
 The previous metrics analyze photo-z performance for point estimates, which are insufficient for modern cosmology \citep{newman2022}, and for ensembles of posteriors, which is often misleading and not a good indicator of performance for science applications \citePZp.
-The methods introduced in this paper enable the creation of galaxy catalogs for which each galaxy has a \emph{true} redshift posterior, which will enable more comprehensive evaluation of photo-z estimators.
+The methods introduced in this paper enable the creation of galaxy catalogs for which each galaxy has a true redshift posterior, which will enable more comprehensive evaluation of photo-z estimators.
 Full evaluation of photo-z estimators on a posterior-by-posterior basis is a major goal of the LSST DESC, and will be the focus of forthcoming work.
 
-\section{Conclusion \& Summary}
+\section{Conclusion}
 \label{sec:conclusion}
 
-In this paper we introduced PZFlow, a normalizing flow package in Python, designed for the statistical modeling of tabular astronomical data.
-We used PZFlow to forward model galaxy catalogs that include photometry, redshifts, sizes, and ellipticities.
-In addition, each galaxy in our catalog has a \emph{true} redshift posterior, which can be convolved with measurement errors.
-These true posteriors allow a direct evaluation of the posteriors produced by photo-z estimators.
-A similar comparison can be made to posterior estimates for any other galaxy properties.
+In this paper we introduced PZFlow, a Python package for probabilistic forward modeling of galaxy catalogs, and demonstrated how it will be used to assist the photo-z calibration efforts of the LSST DESC.
+In particular, galaxies generated from a PZFlow model have a natural notion of a true photo-z posterior, to which the redshift posteriors estimated by photo-z algorithms can be directly compared.
+This enables a more comprehensive evaluation of the posteriors produced by photo-z estimators and allows us to avoid the traps of ensemble-only metrics that were identified by \citePZt.
+Validating the full posteriors produced by photo-z estimators is vital for enabling unbiased cosmology inference with next generation surveys like the LSST \citep{newman2022}.
+Synthetic catalogs from PZFlow, together with new metrics of posterior calibration (e.g.
+the tests of local conditional calibration of \citealt{dey2021b, dey2022}) will be used in future data challenges to optimize and to quantify the error rate and biases of the DESC photo-z estimation pipeline.
 
-Direct evaluation of photo-z posteriors is vital for future cosmology analyses which must use all of the information incorporated in the full redshift posteriors \citep{newman2022}.
-This is because only the full redshift posteriors allow one to account for degeneracies in color-redshift space, which will otherwise bias cosmological inference.
-Previous evaluations of photo-z performance have focused on point estimates and metrics for ensembles of posteriors, but \citePZt demonstrated these metrics are misleading and inadequate for modern cosmology.
-This work enables future analysis of photo-z estimators on a per-posterior basis, which is a major goal of the LSST DESC.
+In addition to forward modeling, PZFlow is a powerful tool for density estimation applied to tabular data.
+We demonstrated this by applying PZFlow to the task of photo-z estimation.
+PZFlow achieves high accuracy with very little fine tuning and very few modeling assumptions.
+However, as PZFlow is trained via likelihood maximization of the training set, it exhibits mode-covering behavior --- i.e., in order to not miss any modes in the data, PZFlow tends to be conservative and produce overly broad posteriors.
+Tradeoffs of this variety, however, are inherent in any choice of machine learning model \citePZp.
 
-In addition to forward modeling galaxy catalogs, we demonstrated PZFlow's utility as a density estimator that can be applied to photo-z estimation and other statistical analyses.
-PZFlow achieves high accuracy with relatively little fine tuning and with very few modeling assumptions.
-This makes PZFlow a powerful tool for the statistical analysis of tabular astronomical data.
+While we have developed PZFlow to address the calibration needs of DESC photo-z validation, and have focused on those applications in this paper, we emphasize that PZFlow is a powerful and flexible tool for statistical modeling of any tabular data.
 
@@
 \begin{acknowledgements}
-    Thanks to François~Lanusse for some early advice on normalizing flows and for reviewing the paper, and to Martine Lokken for testing PZFlow.
+    This paper has undergone internal review in the LSST Dark Energy Science Collaboration.
+    We thank the internal reviewers: Martine Lokken, Boris Leistedt, and Benjamin Remy.
+    We also thank François Lanusse for some early advice on normalizing flows and to Martine Lokken for testing PZFlow.
+    This work was funded by the U.S. Department of Energy, Office of Science, under Award DE-SC0011665.
     J.~F.~Crenshaw, B.~J.~Kalmbach, and A.~J.~Connolly acknowledge support from the DiRAC Institute in the Department of Astronomy at the University of Washington.
@@
         \caption{
-            Training losses for the normalizing flows.
+            Training losses for the normalizing flows used to simulate the galaxy catalog.
             Left: losses for the regular flow.
-            After epochs 50 and 100, you can see a drop in the loss due to the decrease in the learning rate.
             Right: losses for the conditional flow.
-            After epochs 150 and 300, you can see a drop in the loss due to the decrease in the learning rate.
         }
@@
 \end{align}
-where the expectation is performed over galaxies in the training set and $p(x)$ is defined in Equation \ref{eq:px}.
+where the expectation is performed over galaxies in the training set and $p(x)$ is defined in Equation~\ref{eq:px}.
 
-For the main flow in Section \ref{sec:galaxy-catalog}, we trained for 200 epochs.
+For the main flow in Section~\ref{sec:galaxy-catalog}, we trained for 200 epochs.
 We used the Adam optimizer \citep{adam}, starting with a learning rate of $10^{-5}$.
@@
 Training took 7 minutes on a Tesla P100 12GB GPU.
-The training loss for this flow is in the left panel of Figure \ref{fig:galaxy-flow-losses}.
+The training loss for this flow is in the left panel of Figure~\ref{fig:galaxy-flow-losses}.
 
-For the conditional flow in \ref{sec:galaxy-catalog}, we trained for 600 epochs.
+For the conditional flow in~\ref{sec:galaxy-catalog}, we trained for 600 epochs.
 Again, we used Adam with an initial learning rate of $10^{-5}$.
 We decreased the learning rate by a factor of 10 every 200 epochs.
-The training loss for this flow is in the right panel of Figure \ref{fig:galaxy-flow-losses}.
+The training loss for this flow is in the right panel of Figure~\ref{fig:galaxy-flow-losses}.
 
-For each of the flows that make up the flow ensemble in Section \ref{sec:photo-z}, we trained for 200 epochs using the Adam optimizer.
+For each of the flows that make up the flow ensemble in Section~\ref{sec:photo-z}, we trained for 200 epochs using the Adam optimizer.
 We started each with a learning rate of $10^{-4}$, which we decreased by a factor of 10 after the 100th and 150th epochs.
-The training loss for the ensemble is in Figure \ref{fig:ensemble-losses}.
+The training loss for the ensemble is in Figure~\ref{fig:ensemble-losses}.
 Each flow achieved nearly the same training loss.
@@
 This latter approximation is made by \citet{ivezic2019}, and errors are assumed to be Gaussian in magnitude space.
-In contrast, we use the exact form of Equation \ref{eq:err}, and model errors as Gaussian in flux space.
+In contrast, we use the exact form of Equation~\ref{eq:err}, and model errors as Gaussian in flux space.
 Note that after the photometric errors are applied, the error is re-calculated from the ``observed'' flux, and this new error is reported as the estimated photometric error.

### 2023-11-28 src/tex/ms.tex
Fixed latex typo

@@
 where $\mathbf{\hat{m}}$ are the observed magnitudes with photometric errors $\mathbf{\sigma_m}$, and $p(\mathbf{\hat{m}}, \mathbf{\sigma_m})$ is a Gaussian in flux space.
-This allows our ensemble of flows to approximate the distribution $p(z, \mathbf{m})$, where \mathbf{m} is the vector of true magnitudes for the galaxy.
+This allows our ensemble of flows to approximate the distribution $p(z, \mathbf{m})$, where $\mathbf{m}$ is the vector of true magnitudes for the galaxy.
 For more details on training the ensemble, see Appendix~\ref{app:training-details}.

### 2024-02-17 src/tex/ms.tex
Moved of rev2

@@
     Evaluating the accuracy and calibration of the redshift posteriors produced by photometric redshift (photo-z) estimators is vital for enabling precision cosmology and extragalactic astrophysics with modern wide-field photometric surveys.
-    Evaluating photo-z posteriors on a per-galaxy basis is difficult, however, as galaxies have a true redshift, but not a true redshift posterior.
+    Evaluating photo-z posteriors on a per-galaxy basis is difficult, however, as real galaxies have a true redshift but not a true redshift posterior.
     We introduce PZFlow, a Python package for the probabilistic forward modeling of galaxy catalogs with normalizing flows.
@@
 One way to enable posterior validation at the per-galaxy level is to simulate catalogs by sampling galaxies from a probabilistic model for which you have direct access to the underlying probability distribution.
-Then, for each galaxy, you can calculate the posterior over redshift with respect to the same model that generated the galaxy.
+Then, for each galaxy, you can calculate the posterior over redshift \textbf{directly from} the same model that generated the galaxy.
 This provides a ``true'' redshift posterior for each simulated galaxy, to which you can compare the posteriors from photo-z estimators.
@@
 By constructing a stack with $D$ coupling layers, RQ-RSCs individually transform each of the $D$ dimensions of $x$ as a function of the other $D-1$ dimensions.
+\textbf{This allows the network to learn complex relationships between every subset of the dimensions.}
 In the limit of high spline resolution (i.e.
@@
 This distribution is compact, and by varying $\alpha$ and $\beta$ this distribution can take on a wide variety of shapes with different means, skews, and kurtoses, allowing the inductive bias of the prior to adapt to structure in the data during training.
-However, as RQ-RSCs are defined on the domain $[-B, B]$, it is more convenient to use a modified Beta distribution, which we name the \emph{Centered Beta distribution}:
+\textbf{However, the Beta distribution is defined on the domain $[0, 1]$, while} RQ-RSCs are defined on $[-B, B]$.
+It is therefore more convenient to use a modified Beta distribution, which we name the \emph{Centered Beta distribution}:
 \begin{align}
@@
         \caption{
-            Comparing the $r-i$ vs redshift distribution for galaxy samples from CosmoDC2 without photometric noise (left), CosmoDC2 with photometric noise (middle), and from the normalizing flow (right).
+            Comparison of the $r-i$ vs redshift distribution for galaxy samples from CosmoDC2 without photometric noise (left), CosmoDC2 with photometric noise (middle), and from the normalizing flow (right).
             High-redshift galaxies in CosmoDC2 lie along discrete tracks in redshift-color space.
@@
 
-Since we have direct access to the probability distribution from which the photometry and redshifts are drawn, we can analytically calculate the true redshift posterior for each galaxy: $p(z|\mathbf{\hat{m}})$ where $\mathbf{\hat{m}}$ is the vector of noisy galaxy magnitudes.
-We note this is not an estimate, like would be returned by a photo-z estimator, but rather the truth, obtained from the model that generated the photometry and redshifts in the first place.
+Since we have direct access to the probability distribution from which the photometry and redshifts are drawn, \textbf{using Equation~\ref{eq:px},} we can analytically calculate the true redshift posterior for each galaxy: $p(z|\mathbf{\hat{m}})$ where $\mathbf{\hat{m}}$ is the vector of noisy galaxy magnitudes.
+We note this is not an estimate, like would be returned by a photo-z estimator, but rather the \textbf{simulated} truth, obtained from the model that generated the photometry and redshifts in the first place.
 
@@
 You may wish to marginalize over all values of $\mathbf{\hat{m}}_\text{x}$ if the galaxy was not observed in those bands.
-This may occur, for example, when simulated a joint Euclid-LSST catalog \citep{euclid}, as not all galaxies will have photometry from both.
+This may occur, for example, when simulating a joint Euclid-LSST catalog \citep{euclid}, as not all galaxies will have photometry from both.
 You may also wish to marginalize over all values beyond the limiting magnitudes to simulate a galaxy that was observed but not detected in the corresponding bands.
@@
 For the training set, we use 100,000 galaxies from the catalog created in Section~\ref{sec:galaxy-catalog}.
-Each galaxy in the training set has a true redshift and observed magnitudes in the $ugrizy$ bands, with corresponding photometric errors.
+Each galaxy in the training set has a true redshift and observed \textbf{noisy} magnitudes in the $ugrizy$ bands, with corresponding photometric errors.
 To account for the photometric error, at the start of each training epoch, we resample the training set from the photometric error distributions.
@@
 
-For each flow in the ensemble, we estimate $p(z, \mathbf{m})$ by marginalizing over the photometric errors:
+For each flow in the ensemble, we estimate \textbf{the redshift posterior} by marginalizing over the photometric errors:
 \begin{align}
@@
             The bias, scatter, and outlier fraction of the photo-z point estimates as a function of true galaxy redshift.
-            The dashed black lines represent the upper limits for LSST cosmology as stated in the LSST DESC SRD \citep{descSRD}.
-            These lines are to provide a sense of scale for these metrics.
-            You can see that PZFlow meets the bias and scatter requirements up to redshift $\sim$ 1.5, while meeting the outlier fraction requirements for all redshifts.
+            \textbf{
+            The dashed black lines represent the limits for LSST cosmology as stated in the LSST DESC SRD \citep{descSRD}.
+            The bias must be between these dashed lines, while the scatter and outlier fraction must be below the dashed lines.
+            }
+            You can see that PZFlow meets the bias and scatter requirements up to redshift $\sim$ 1.5 \textbf{(the scatter is high at low redshifts due to the relative small number of low-redshift galaxies in our training set)}.
+            PZFlow meets the outlier requirements for all redshifts.
             We note that individual redshifts do not actually need to meet the bias requirement as long as the bias can be well calibrated via some other source, e.g. galaxy clustering.
@@
 
-Like many photo-z estimators, PZFlow performs well to a redshift of approximately 1.5.
+Like many photo-z estimators, PZFlow performs well to a redshift of approximately 1.5 \textbf{(the scatter is high at low redshifts due to the relatively small number of low-redshift galaxies in our training set)}.
 At higher redshifts, our estimator does not meet the bias and scatter requirements, because there is very little training data in this redshift range.
@@
 
-Another common metric is the probability integral transform (PIT) (see e.g. \citePZa, \citealt{dey2022}).
-The PIT metric is used to determine if posteriors are well calibrated.
-For example, if the photo-z posteriors produced by an estimator are well calibrated, you would expect the true redshifts of the galaxies to fall within 50\% confidence intervals 50\% of the time.
+Another common metric is the probability integral transform (PIT) (see e.g. \citePZa, \citealt{dey2022}), \textbf{which is a histogram of the cumulative density function (CDF) of each posterior.
+I.e., given an estimated posterior  $p(z| \mathbf{\hat{m}}, \mathbf{\sigma_m})$ and the true redshift $z_\text{true}$, the CDF is
+\begin{align}
+    \mathrm{CDF} = \int_0^{z_\text{true}} p(z| \mathbf{\hat{m}}, \mathbf{\sigma_m}) \, dz.
+\end{align}
+For perfectly calibrated posteriors, the CDF distribution (the PIT histogram) is uniform between $0$ and $1$.
+This is because, for example, if the photo-z posteriors produced by an estimator are well calibrated you would expect the true redshifts of the galaxies to fall within the 50\% confidence intervals 50\% of the time.}
+
 The PIT histogram for our estimator is shown in Figure~\ref{fig:pit-histogram}.
@@
 This negative bias is visible for high-redshift galaxies in the far left panel of Figure~\ref{fig:binned-metrics}.
+\textbf{Calibrating these posteriors, either via altering the training loss or post-processing the posteriors, is beyond the scope of this paper.
+However this calibration could be achieved, for example, using the methods of \citet{dey2022}.}
 
@@
 In particular, galaxies generated from a PZFlow model have a natural notion of a true photo-z posterior, to which the redshift posteriors estimated by photo-z algorithms can be directly compared.
-This enables a more comprehensive evaluation of the posteriors produced by photo-z estimators and allows us to avoid the traps of ensemble-only metrics that were identified by \citePZt.
+This enables a more comprehensive evaluation of the posteriors produced by photo-z estimators that \textbf{we expect will avoid the traps} of ensemble-only metrics that were identified by \citePZt.
 Validating the full posteriors produced by photo-z estimators is vital for enabling unbiased cosmology inference with next generation surveys like the LSST \citep{newman2022}.
@@
 However, as PZFlow is trained via likelihood maximization of the training set, it exhibits mode-covering behavior --- i.e., in order to not miss any modes in the data, PZFlow tends to be conservative and produce overly broad posteriors.
-Tradeoffs of this variety, however, are inherent in any choice of machine learning model \citePZp.
+Increasing the amount of training data will likely alleviate these issues, but tradeoffs of this variety are inherent in any choice of machine learning model \citePZp.
 

### 2024-04-11 src/tex/ms.tex
Added Moskowitz paper and removed bolding from review process.

@@
 One way to enable posterior validation at the per-galaxy level is to simulate catalogs by sampling galaxies from a probabilistic model for which you have direct access to the underlying probability distribution.
-Then, for each galaxy, you can calculate the posterior over redshift \textbf{directly from} the same model that generated the galaxy.
-This provides a ``true'' redshift posterior for each simulated galaxy, to which you can compare the posteriors from photo-z estimators.
+Then, for each galaxy, you can calculate the posterior over redshift directly from the same model that generated the galaxy.
+This provides a ``true'' redshift posterior for each simulated galaxy to which you can compare the posteriors from photo-z estimators, enabling the validation of individual photo-z posterior estimates.
 
@@
 By constructing a stack with $D$ coupling layers, RQ-RSCs individually transform each of the $D$ dimensions of $x$ as a function of the other $D-1$ dimensions.
-\textbf{This allows the network to learn complex relationships between every subset of the dimensions.}
+This allows the network to learn complex relationships between every subset of the dimensions.
 In the limit of high spline resolution (i.e.
@@
 This distribution is compact, and by varying $\alpha$ and $\beta$ this distribution can take on a wide variety of shapes with different means, skews, and kurtoses, allowing the inductive bias of the prior to adapt to structure in the data during training.
-\textbf{However, the Beta distribution is defined on the domain $[0, 1]$, while} RQ-RSCs are defined on $[-B, B]$.
+However, the Beta distribution is defined on the domain $[0, 1]$, while RQ-RSCs are defined on $[-B, B]$.
 It is therefore more convenient to use a modified Beta distribution, which we name the \emph{Centered Beta distribution}:
@@
     \item \citet{stylianou2022} used PZFlow to forward model galaxy data with true redshift posteriors in order to evaluate the impact of survey incompleteness and spec-z errors on photo-z estimation;
-    \item \citet{lokken2022} used PZFlow to smooth high-redshift artifacts in simulations of host galaxies for supernovae and other transients.
+    \item \citet{lokken2022} used PZFlow to smooth high-redshift artifacts in simulations of host galaxies for supernovae and other transients;
+    \item \citet{moskowitz2024} used PZFlow to explore data augmentation for photo-z spectroscopic training sets.
 \end{itemize}
@@
 
-Since we have direct access to the probability distribution from which the photometry and redshifts are drawn, \textbf{using Equation~\ref{eq:px},} we can analytically calculate the true redshift posterior for each galaxy: $p(z|\mathbf{\hat{m}})$ where $\mathbf{\hat{m}}$ is the vector of noisy galaxy magnitudes.
-We note this is not an estimate, like would be returned by a photo-z estimator, but rather the \textbf{simulated} truth, obtained from the model that generated the photometry and redshifts in the first place.
+Since we have direct access to the probability distribution from which the photometry and redshifts are drawn, using Equation~\ref{eq:px}, we can analytically calculate the true redshift posterior for each galaxy: $p(z|\mathbf{\hat{m}})$ where $\mathbf{\hat{m}}$ is the vector of noisy galaxy magnitudes.
+We note this is not an estimate, like would be returned by a photo-z estimator, but rather the simulated truth, obtained from the model that generated the photometry and redshifts in the first place.
+Of course the resolution of the posterior is limited to the choice of redshift grid.
 
@@
 For the training set, we use 100,000 galaxies from the catalog created in Section~\ref{sec:galaxy-catalog}.
-Each galaxy in the training set has a true redshift and observed \textbf{noisy} magnitudes in the $ugrizy$ bands, with corresponding photometric errors.
+Each galaxy in the training set has a true redshift and observed noisy magnitudes in the $ugrizy$ bands, with corresponding photometric errors.
 To account for the photometric error, at the start of each training epoch, we resample the training set from the photometric error distributions.
@@
 
-For each flow in the ensemble, we estimate \textbf{the redshift posterior} by marginalizing over the photometric errors:
+After training, we use each flow in the ensemble to estimate the redshift posterior by marginalizing over the photometric errors:
 \begin{align}
@@
             The bias, scatter, and outlier fraction of the photo-z point estimates as a function of true galaxy redshift.
-            \textbf{
             The dashed black lines represent the limits for LSST cosmology as stated in the LSST DESC SRD \citep{descSRD}.
             The bias must be between these dashed lines, while the scatter and outlier fraction must be below the dashed lines.
-            }
-            You can see that PZFlow meets the bias and scatter requirements up to redshift $\sim$ 1.5 \textbf{(the scatter is high at low redshifts due to the relative small number of low-redshift galaxies in our training set)}.
+            You can see that PZFlow meets the bias and scatter requirements up to redshift $\sim$ 1.5 (the scatter is high at low redshifts due to the relative small number of low-redshift galaxies in our training set).
             PZFlow meets the outlier requirements for all redshifts.
@@
 
-Like many photo-z estimators, PZFlow performs well to a redshift of approximately 1.5 \textbf{(the scatter is high at low redshifts due to the relatively small number of low-redshift galaxies in our training set)}.
+Like many photo-z estimators, PZFlow performs well to a redshift of approximately 1.5 (the scatter is high at low redshifts due to the relatively small number of low-redshift galaxies in our training set).
 At higher redshifts, our estimator does not meet the bias and scatter requirements, because there is very little training data in this redshift range.
@@
 
-Another common metric is the probability integral transform (PIT) (see e.g. \citePZa, \citealt{dey2022}), \textbf{which is a histogram of the cumulative density function (CDF) of each posterior.
+Another common metric is the probability integral transform (PIT) (see e.g. \citePZa, \citealt{dey2022}), which is a histogram of the cumulative density function (CDF) of each posterior.
 I.e., given an estimated posterior  $p(z| \mathbf{\hat{m}}, \mathbf{\sigma_m})$ and the true redshift $z_\text{true}$, the CDF is
@@
 For perfectly calibrated posteriors, the CDF distribution (the PIT histogram) is uniform between $0$ and $1$.
-This is because, for example, if the photo-z posteriors produced by an estimator are well calibrated you would expect the true redshifts of the galaxies to fall within the 50\% confidence intervals 50\% of the time.}
+This is because, for example, if the photo-z posteriors produced by an estimator are well calibrated you would expect the true redshifts of the galaxies to fall within the 50\% confidence intervals 50\% of the time.
 
@@
 This negative bias is visible for high-redshift galaxies in the far left panel of Figure~\ref{fig:binned-metrics}.
-\textbf{Calibrating these posteriors, either via altering the training loss or post-processing the posteriors, is beyond the scope of this paper.
-However this calibration could be achieved, for example, using the methods of \citet{dey2022}.}
+Calibrating these posteriors, either via altering the training loss or post-processing the posteriors, is beyond the scope of this paper.
+However this calibration could be achieved, for example, using the methods of \citet{dey2022}.
 
@@
 In particular, galaxies generated from a PZFlow model have a natural notion of a true photo-z posterior, to which the redshift posteriors estimated by photo-z algorithms can be directly compared.
-This enables a more comprehensive evaluation of the posteriors produced by photo-z estimators that \textbf{we expect will avoid the traps} of ensemble-only metrics that were identified by \citePZt.
+This enables a more comprehensive evaluation of the posteriors produced by photo-z estimators that we expect will avoid the traps of ensemble-only metrics that were identified by \citePZt.
 Validating the full posteriors produced by photo-z estimators is vital for enabling unbiased cosmology inference with next generation surveys like the LSST \citep{newman2022}.
@@
 }
+
+\newpage
 
